---
title: "Figure2_RPCAs"
author: "Jacoby Baker"
date: "2026-03-23"
output: html_document
---


# Set up directories and variables
```{r}
# setwd('scripts/')
proj_prefix <- "AVITI_MiSeq_comparison"
proj_name <- "AVITI VS MiSeq Comparison"

data_directory <- paste0('../data/filtered_seq_data/')
rarefaction_data <- paste0('../data/rarefaction_data/')


getwd()
fig2_directory <- paste0("../figures/Fig2/")
# diversity_directory <- paste0("../figures/diversity/")
rarefaction_directory <- paste0("../figures/rarefaction/")
deicode_directory <- paste0('../data/DEICODE/18S/')
# barchart_directory <- paste0("../figures/bar_charts/")

marker = "18S"
```



# Load libraries
```{r}

#load libraries
library(qiime2R)
library(readr) #read csv files
library(lubridate) #for date modifications
library(vegan)

library(dplyr)
library(ggplot2)
library(ggthemes)
library(magrittr) # for piping commands
library(RColorBrewer) #colors for plotting
library("viridis") # for color fill 


library(ggalt)


library(tidyverse)
library(phyloseq)

# library(taxa)
library(dplyr)
# library(metacoder)
# library(ggpubr) # for ggarrange
# library(reshape2) # for presence/absence fig
# 
# library("ape") # to make phylo tree in phyloseq obj
# library(ggtree) # for phylogenetic tree
# library(ggtreeExtra) # for aligning graph layers to outerside of ggtree
# 
# library(ggnewscale) # multiple fill color scales

library(rstatix) # alpha diversity stats kruskal-wallis
```




# Load in Data
```{r}

#ASV table
otu.c <- read_csv(paste0(data_directory, proj_prefix, "_", marker, "_asv_filtered.csv"))
otu.c

#taxa table
tax.c <- read_csv(paste0(data_directory, proj_prefix, "_", marker, "_taxa_filtered.csv"))
tax.c

#metadata table
samp.c <- read_csv(paste0(data_directory, proj_prefix, "_", marker, "_meta_filtered.csv")) %>%
  # reorder the variable sample_source 
  mutate(sample_source = fct_relevel(sample_source,  "CN23 MARS" , "MBTS 17024", "MBTS 20524", "MBTS 22624", "MBTS 25424", "CN24F 150m exp"))
samp.c



#OTU table long format with percent total reads
potu.c <- otu.c %>%
  tidyr::pivot_longer( -ASV, names_to ='sample_name',values_to = 'reads' ) %>%
  ungroup() %>%
  group_by(sample_name) %>%
  mutate(per_tot = reads / sum(reads) *100) %>% 
  arrange(-reads)
head(potu.c)


```



## set lowest taxonomic annotation level for ASVs
```{r}

  species_label <- tax.c  %>%
  # assign  algae
  mutate(Phylum = case_when(Class == "Raphidophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Phaeophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Pelagophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Dictyochophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Chrysophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Synurophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Bolidophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Eustigmatophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  # assign oomycotes
  mutate(Phylum = case_when(Class == "Oomycota" ~ "Oomycota", T ~ Phylum))  %>%
  mutate(Phylum = case_when(Class == "Oomycetes" ~ "Oomycota", T ~ Phylum))%>%
  # assign microeuk
  mutate(Phylum = case_when(Class == "Bigyra" ~ "Bigyra", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Choanoflagellata" ~ "Choanoflagellata", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Cryptophyceae" ~ "Cryptista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Order == "Telonemida" ~ "Cryptista", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Haptophyta" ~ "Haptista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Centroplasthelida" ~ "Haptista", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Ichthyosporea" ~ "Choanozoa", T ~ Phylum))  %>% 
  
  mutate(Class = case_when(Phylum == "Imbricatea" ~ "Imbricatea", T ~ Class)) %>%
  mutate(Phylum = case_when(Class == "Imbricatea" ~ "Cercozoa", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Polycystinea" ~ "Retaria", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Acantharea" ~ "Retaria", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Breviatea" ~ "Sulcozoa", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Aphelidea" ~ "Aphelida", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Filasterea" ~ "Choanozoa", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Phylum %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Phylum)) %>% 
  mutate(Class = case_when(Class %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Class)) %>% 
  mutate(Order = case_when(Order %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Order)) %>% 
  mutate(Family = case_when(Family %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Family)) %>% 
  mutate(Genus = case_when(Genus %in% c("unknown", "unassigned", "no_hit", "g_") ~ "unknown", T ~ Genus)) %>% 
  mutate(Species = case_when(Species %in% c("unknown", "unassigned", "no_hit", "s_") ~ "unknown", T ~ Species))

  species_label


```




# PHYLOSEQ  
define function
```{r}
vegan_otu <- function(physeq) {
  OTU <- otu_table(physeq)
  if (taxa_are_rows(OTU)) {
    OTU <- t(OTU)
  }
  return(as(OTU, "matrix"))
}
```


# ##### Non rarefied data #####
## ## Figure 2 A - All samples, not rarefied, all reads

### create Phyloseq object 
```{r}

  
  subset = c("all_reads")
  Sys.setenv(subset = subset)
  
  metadata_updated <- samp.c
  

  
species_label <- tax.c %>% column_to_rownames(var = "ASV")
  
  rownames(metadata_updated) <- NULL
  metadata_updated <- column_to_rownames(metadata_updated,"sample_name")
  
  
  phylo <- merge_phyloseq(otu_table(as.matrix(read.csv(paste0(data_directory, proj_prefix, "_", marker, "_asv_filtered.csv"), row.names = 1,check.names = FALSE)),taxa_are_rows = TRUE),
                          tax_table(as.matrix(species_label)),
                          sample_data(metadata_updated))
  
  phylo

    sample_data(phylo)
  
  phylo_obj <- phylo

  phylo_obj_env <- prune_samples(sample_data(phylo)$sample_type %in% c("environmental"), phylo) 
  
  
  
  
  
  sample_data(phylo_obj)
  sample_data(phylo_obj_env)
  # sample_data(phylo_obj_cb)
  
  # Create Tables -----------------------------------------------------------
  
  #create vegan tables
  vegan_matrix_env=vegan_otu(phylo_obj_env)
  otu_tab_env=otu_table(phylo_obj_env)
  tax_tab_env= tax_table(phylo_obj_env)
  samp_dat_env <- sample_data(phylo_obj_env)
  
  # Rarefy ------------------------------------------------------------------
  #print(sample_sums(phylo_obj_env))
  print(min(sample_sums(phylo_obj_env)))
  print(max(sample_sums(phylo_obj_env)))
  print(mean(sample_sums(phylo_obj_env)))
  print(phylo_obj_env)
  # # Remove samples with <5000 reads (like DEICODE)
  # oBiom_r_env = prune_samples(sample_sums(phylo_obj_env)>=5000, phylo_obj_env)
  # #set random seed for reproducibility later
  # print(min(sample_sums(oBiom_r_env)))
  # oBiom_r_env = rarefy_even_depth(oBiom_r_env, sample.size = min(sample_sums(oBiom_r_env)), rngseed = 678, replace = FALSE, trimOTUs = TRUE, verbose = TRUE)
  # print(oBiom_r_env)
  # assign(paste0("oBiom_r_env_",marker),oBiom_r_env)
  # 
  # #create matrixes
  # rvegan_matrix_env=vegan_otu(oBiom_r_env)
  # rotu_tab_env=otu_table(oBiom_r_env)
  # rtax_tab_env= tax_table(oBiom_r_env)
  # rsamp_dat_env <- sample_data(oBiom_r_env)
  # 
  # #save rarefied data
  # #save to csv file
  # file = paste(rarefaction_data,marker,'_env_Non-Rarefied_asv_table_',subset,'.csv', sep="")
  # write.csv(otu_tab_env, file)
  # file = paste(rarefaction_data,marker,'_env_Non-Rarefied_taxa_table_',subset,'.csv', sep="")
  # write.csv(tax_tab_env, file)
  # # file = paste(rarefaction_data,marker,'_env_Non-Rarefied_meta_table_',subset,'.csv', sep="")
  # # write.csv(samp_dat_env, file)
  # 
  # file = paste(rarefaction_data,marker,'_env_Rarefied_asv_table_',subset,'.csv', sep="")
  # write.csv(rotu_tab_env, file)
  # file = paste(rarefaction_data,marker,'_env_Rarefied_taxa_table_',subset,'.csv', sep="")
  # write.csv(rtax_tab_env, file)
  # # file = paste(rarefaction_data,marker,'_env_Rarefied_meta_table_',subset,'.csv', sep="")
  # # write.csv(rsamp_dat_env, file)
  
  
  ### DEICODE PCA 
  # environmental samples
  phylo_obj_env_asv_table <- as.data.frame(otu_table(phylo_obj_env))
  phylo_obj_env_asv_table <- rownames_to_column(phylo_obj_env_asv_table,"#OTUID")
  write.table(phylo_obj_env_asv_table,paste0(deicode_directory,"/phylo_obj_env_",marker,"_asv_table_for_biom_",subset,".txt"),sep = "\t",row.names = FALSE,quote = FALSE)
  phylo_obj_env_tax_table <- as.data.frame(as(tax_table(phylo_obj_env),"matrix"))
  phylo_obj_env_tax_table <- rownames_to_column(phylo_obj_env_tax_table,"#OTUID")
  write.table(phylo_obj_env_tax_table,paste0(deicode_directory,"/phylo_obj_env_",marker,"_tax_table_for_biom_",subset,".txt"),sep = "\t",row.names = FALSE,quote = FALSE)
  phylo_obj_env_metadata <- as.data.frame(as.matrix(sample_data(phylo_obj_env)))
  phylo_obj_env_metadata <- rownames_to_column(phylo_obj_env_metadata,"#SampleID")
  write.table(phylo_obj_env_metadata,paste0(deicode_directory,"/phylo_obj_env_",marker,"_metadata_for_biom_",subset,".txt"),sep = "\t",row.names = FALSE,quote = FALSE)
  

  
```


### DEICODE
#### deicode PCA use a loop to jump into each of the folders for the three markers for all samples
```{bash engine.opts='-l'}
# move into the directory where you are going to run the DEICODE script
cd ../data/DEICODE
pwd


# Activate the conda qiime2 enviroment - check version
conda activate qiime2-2022.11

for folder in 18S  
do
  for data in phylo_obj_env

  do
  cd $folder
  #!/usr/bin/env bash

  # This bash script can be run with bash deicode_template_072820.sh
  # All of the input files (asv_table_for_deicode.txt, tax_table_for_deicode.txt, and metadata_for_deicode.txt) must be in the   same directory as this script.

  ### Description ###
  #
  # You will need qiime2 and DEICODE installed on your machine in order for it to work.
  # You will have four outputs: 1) distance.qza, 2) ordination.qza, 3) biplot.qza, and 4_ qurro_plot_deicoderank.qzv
  # Of these outputs, the most import one is the "ordination.qza" file - this is what you will use in R
  # biplot.qza can be visualized at view.qiime2.org, and is an interactive way to view your PCA.
  # qurro_plot_deicoderank.qzv will show you the loadings of your taxa on the different PCs.

  #Make biom file
    biom convert -i ./${data}_${folder}_asv_table_for_biom_${subset}.txt -o table.from_txt_json_${data}_${subset}.biom --table-type="OTU table" --to-json
  #add metadata files to biom file - change the merged_tax_table_for_biomm.txt and _merged_for_biom.txt files
   biom add-metadata -i table.from_txt_json_${data}_${subset}.biom -o table.w_md_${data}_${subset}.biom --observation-metadata-fp ${data}_${folder}_tax_table_for_biom_${subset}.txt --sample-metadata-fp ${data}_${folder}_metadata_for_biom_${subset}.txt

  #import into Qiime2 - Change output file
    qiime tools import \
     --input-path table.w_md_${data}_${subset}.biom \
     --output-path ${data}_${folder}_master_${subset}.biom.qza \
     --type FeatureTable[Frequency]

  #run DEICODE - change input table
    qiime deicode rpca \
      --i-table ${data}_${folder}_master_${subset}.biom.qza \
      --p-n-components 3 \
      --p-min-feature-count 10 \
      --p-min-sample-count 5000 \
      --o-biplot ordination_${data}_${subset}.qza \
      --o-distance-matrix distance_${data}_${subset}.qza

  ## Create biplot - change metadata files
    qiime emperor biplot \
      --i-biplot ordination_${data}_${subset}.qza \
      --m-sample-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --m-feature-metadata-file ${data}_${folder}_tax_table_for_biom_${subset}.txt \
      --o-visualization biplot_${data}_${subset}.qzv \
      --p-number-of-features 8

  # Create Qurro Visualization - change metadata files
    qiime qurro loading-plot \
      --i-ranks ordination_${data}_${subset}.qza \
      --i-table ${data}_${folder}_master_${subset}.biom.qza \
      --m-sample-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --m-feature-metadata-file ${data}_${folder}_tax_table_for_biom_${subset}.txt \
      --o-visualization qurro_plot_deicoderank_${data}_${subset}.qzv

   # Run a permanova | permdisp | or anosim test between whatever metadata column you choose (argument --m-metadata-column [column of choosing])
     qiime diversity beta-group-significance \
      --i-distance-matrix distance_${data}_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --m-metadata-column sequencer \
      --p-method anosim \
      --o-visualization sequencer_significance_anosim_${data}_${subset}.qzv

    # Alpha diversity test
    #Shannon
     qiime diversity alpha \
       --i-table ${data}_${folder}_master_${subset}.biom.qza \
       --p-metric shannon \
       --o-alpha-diversity ${data}_${folder}_shannon_alpha_${subset}.qza

      qiime diversity alpha-group-significance \
      --i-alpha-diversity ${data}_${folder}_shannon_alpha_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --o-visualization ${data}_${folder}_shannon_alpha_${subset}.qzv
      
      #Chao1
      qiime diversity alpha \
       --i-table ${data}_${folder}_master_${subset}.biom.qza \
       --p-metric chao1 \
       --o-alpha-diversity ${data}_${folder}_chao1_alpha_${subset}.qza

      qiime diversity alpha-group-significance \
      --i-alpha-diversity ${data}_${folder}_chao1_alpha_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --o-visualization ${data}_${folder}_chao1_alpha_${subset}.qzv
      
      #Observed OTUs
      qiime diversity alpha \
       --i-table ${data}_${folder}_master_${subset}.biom.qza \
       --p-metric observed_features \
       --o-alpha-diversity ${data}_${folder}_observed_otus_alpha_${subset}.qza

      qiime diversity alpha-group-significance \
      --i-alpha-diversity ${data}_${folder}_observed_otus_alpha_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --o-visualization ${data}_${folder}_observed_otus_alpha_${subset}.qzv
   done

  cd ../
done

```




#### RPCA of all samples, non rarefied
```{r}

#Import Qiime2 Results
file = paste(deicode_directory,"/ordination_phylo_obj_env_",subset,".qza",sep="")
print(file)
pco<-read_qza(file)
pco$uuid
#look at data
head(pco$data$ProportionExplained)
pco$data$Vectors[1:5, 1:4]

#create proportion explained labels
label.PC1 <- paste("PC1: ", round(pco$data$ProportionExplained$PC1, 3)*100,"%")
label.PC1
label.PC2 <- paste("PC2: ", round(pco$data$ProportionExplained$PC2, 3)*100,"%")
label.PC2
label.PC3 <- paste("PC3: ", round(pco$data$ProportionExplained$PC3, 3)*100,"%")
label.PC3

#Join with sample data
pco$data$Vectors <- pco$data$Vectors %>% 
  rename(., "sample_name" = "SampleID" )

pcscores <- left_join(pco$data$Vectors, samp.c)


#format loading scores
loadings <- as.data.frame(pco$data$Species)
loadings$ASV <- loadings$FeatureID

#join on OTU, adding taxa info
loadings <- left_join(loadings, tax.c, by="ASV")

#export pcscores
file = paste(deicode_directory,"scores/pcscores_",marker,"_phylo_obj_env_Dada2_Qiime2_",subset,".csv",sep="")
print(file)
write.csv(pcscores, file)
file = paste(deicode_directory,"scores/loadings_",marker,"_phylo_obj_env_Dada2_Qiime2_",subset,".csv",sep="")
print(file)
write.csv(loadings, file)
pcscores[1:5, 1:9]  #long because of sample data
head(loadings)


# Plot RPCA ---------------------------------------------------------------

cols6 <- c("navyblue", "goldenrod3", "darkmagenta", "firebrick", "firebrick2", "darkolivegreen4", "violetred3")

# plot inverse of PC1 and PC2 values so the pattern matches more with the rarefied PCA (e.g., the 150m ESP samples in the same quadrant)

pcscores %>% mutate(depth_bin = case_when(depth >= 0 & depth <= 20 ~ "0-20"), 
                        depth_bin = case_when(depth == 60 ~ "60", T ~ depth_bin), 
                        depth_bin = case_when(depth == 100 ~ "100", T ~ depth_bin), 
                        depth_bin = case_when(depth == 150 ~ "150", T ~ depth_bin), 
                        depth_bin = case_when(depth >= 200 & depth <= 700 ~ "200-700", T ~ depth_bin),
                        depth_bin = case_when(sample_source == "CN24F 150m exp" ~ "150 - benchtop experiment", T ~ depth_bin),
                        depth_bin = fct_relevel(depth_bin, "0-20", "60", "100", "150", "150 - benchtop experiment", "200-700"),
                        source_depth_bin = paste(sample_source, depth_bin, sep = " - "),
                        source_depth_bin = reorder(source_depth_bin, depth)) %>% 
  ggplot(aes(-PC1,-PC2, fill=depth_bin)) +
  geom_line(aes(group = original_name, color = depth_bin), linetype = 2, alpha = 0.7) +
  geom_point(size=4, alpha=0.8, aes(shape = sequencer, color = depth_bin)) + 
  scale_shape_manual(values = c(21, 1)) +
    scale_fill_manual(values = cols6) +
    scale_color_manual(values = cols6, guide="none") +
  labs(x=label.PC1 , 
       y=label.PC2, 
       fill = "Depth Bin (m)",
       shape = "Sequencer",
       subtitle = "Comparison between sequencers\nLines connect sample replicates",
       title = "RPCA - non rarefied data, all reads"  )+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.text = element_text(size = 12, colour = "black"),
        legend.title = element_text(face = "bold", size=12),
        legend.key = element_rect(fill = FALSE, colour = FALSE),
        legend.key.size = unit(0.1,"line"),
        plot.title = element_text(size = 22),
        plot.subtitle = element_text(size = 18),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15)
  )+ guides(fill = guide_legend(override.aes = list(size=5, shape=c(21))),
            shape = guide_legend(override.aes = list(size = 5, shape = c(19, 1))))

# save as .png
filename = paste(fig2_directory, 'Figure2a_RPCA_by_depth_bin_inverse_',marker,"_",subset,'.png', sep='')
print('Plot of PC1PC2 colored by Year:')
filename
ggsave(filename,height = 9, width =10, units = 'in')

# save as .svg
filename = paste(fig2_directory, 'Figure2a_RPCA_by_depth_bin_inverse_',marker,"_",subset,'.svg', sep='')
print('Plot of PC1PC2 colored by Year:')
filename
ggsave(filename,height = 9, width =10, units = 'in')


```



# ##### Rarefied data #####
## ## Figure 2 B - all samples, rarefied data, all reads

### create Phyloseq object of all  samples
```{r}

  
  subset = c("Rarefied_data")
  Sys.setenv(subset = subset)
  
  metadata_updated <- samp.c %>% mutate(depth_bin = case_when(depth >= 0 & depth <= 20 ~ "0-20"), 
                        depth_bin = case_when(depth == 60 ~ "60", T ~ depth_bin), 
                        depth_bin = case_when(depth == 100 ~ "100", T ~ depth_bin), 
                        depth_bin = case_when(depth == 150 ~ "150", T ~ depth_bin), 
                        depth_bin = case_when(depth >= 200 & depth <= 700 ~ "200-700", T ~ depth_bin),
                        depth_bin = case_when(sample_source == "CN24F 150m exp" ~ "150 - benchtop experiment", T ~ depth_bin),
                        depth_bin = fct_relevel(depth_bin, "0-20", "60", "100", "150", "150 - benchtop experiment", "200-700"))
  

  
species_label <- tax.c %>% column_to_rownames(var = "ASV")
  
  rownames(metadata_updated) <- NULL
  metadata_updated <- column_to_rownames(metadata_updated,"sample_name")
  
  
  phylo <- merge_phyloseq(otu_table(as.matrix(read.csv(paste0(data_directory, proj_prefix, "_", marker, "_asv_filtered.csv"), row.names = 1,check.names = FALSE)),taxa_are_rows = TRUE),
                          tax_table(as.matrix(species_label)),
                          sample_data(metadata_updated))
  
  phylo

    sample_data(phylo)
  
  phylo_obj <- phylo

  phylo_obj_env <- prune_samples(sample_data(phylo)$sample_type %in% c("environmental"), phylo) 
  
  
  
  
  
  sample_data(phylo_obj)
  sample_data(phylo_obj_env)
  # sample_data(phylo_obj_cb)
  
  # Create Tables -----------------------------------------------------------
  
  #create vegan tables
  vegan_matrix_env=vegan_otu(phylo_obj_env)
  otu_tab_env=otu_table(phylo_obj_env)
  tax_tab_env= tax_table(phylo_obj_env)
  samp_dat_env <- sample_data(phylo_obj_env)
  
  # Rarefy ------------------------------------------------------------------
  #print(sample_sums(phylo_obj_env))
  print(min(sample_sums(phylo_obj_env)))
  print(max(sample_sums(phylo_obj_env)))
  print(mean(sample_sums(phylo_obj_env)))
  print(phylo_obj_env)
  # Remove samples with <500 reads (like DEICODE)
  oBiom_r_env = prune_samples(sample_sums(phylo_obj_env)>=5000, phylo_obj_env)
  #set random seed for reproducibility later
  print(min(sample_sums(oBiom_r_env)))
  oBiom_r_env = rarefy_even_depth(oBiom_r_env, sample.size = min(sample_sums(oBiom_r_env)), rngseed = 678, replace = FALSE, trimOTUs = TRUE, verbose = TRUE)
  print(oBiom_r_env)
  assign(paste0("oBiom_r_env_",marker),oBiom_r_env)
  
  #create matrixes
  rvegan_matrix_env=vegan_otu(oBiom_r_env)
  rotu_tab_env=otu_table(oBiom_r_env)
  rtax_tab_env= tax_table(oBiom_r_env)
  rsamp_dat_env <- rownames_to_column(as.data.frame(as.matrix(sample_data(oBiom_r_env))), var = "sample_name") #sample_data(oBiom_r_env)
  
  #save rarefied data
  # #save to csv file
  file = paste(rarefaction_data,marker,'_env_Non-Rarefied_asv_table_',subset,'.csv', sep="")
  write.csv(otu_tab_env, file)
  file = paste(rarefaction_data,marker,'_env_Non-Rarefied_taxa_table_',subset,'.csv', sep="")
  write.csv(tax_tab_env, file)
  # file = paste(rarefaction_data,marker,'_env_Non-Rarefied_meta_table_',subset,'.csv', sep="")
  # write.csv(samp_dat_env, file)

  file = paste(rarefaction_data,marker,'_env_Rarefied_asv_table_',subset,'.csv', sep="")
  write.csv(rotu_tab_env, file)
  file = paste(rarefaction_data,marker,'_env_Rarefied_taxa_table_',subset,'.csv', sep="")
  write.csv(rtax_tab_env, file)
  file = paste(rarefaction_data,marker,'_env_Rarefied_meta_table_',subset,'.csv', sep="")
  write.csv(rsamp_dat_env, file)
  
  
  ### DEICODE PCA 
  # rarefied
  phylo_obj_env_asv_table <- as.data.frame(otu_table(oBiom_r_env))
  phylo_obj_env_asv_table <- rownames_to_column(phylo_obj_env_asv_table,"#OTUID")
  write.table(phylo_obj_env_asv_table,paste0(deicode_directory,"/oBiom_r_env_",marker,"_asv_table_for_biom_",subset,".txt"),sep = "\t",row.names = FALSE,quote = FALSE)
 
  phylo_obj_env_tax_table <- as.data.frame(as(tax_table(oBiom_r_env),"matrix"))
  phylo_obj_env_tax_table <- rownames_to_column(phylo_obj_env_tax_table,"#OTUID")
  write.table(phylo_obj_env_tax_table,paste0(deicode_directory,"/oBiom_r_env_",marker,"_tax_table_for_biom_",subset,".txt"),sep = "\t",row.names = FALSE,quote = FALSE)
  
  phylo_obj_env_metadata <- as.data.frame(as.matrix(sample_data(oBiom_r_env)))
  phylo_obj_env_metadata <- rownames_to_column(phylo_obj_env_metadata,"#SampleID")
  write.table(phylo_obj_env_metadata,paste0(deicode_directory,"/oBiom_r_env_",marker,"_metadata_for_biom_",subset,".txt"),sep = "\t",row.names = FALSE,quote = FALSE)
  

  
```


### DEICODE
#### deicode PCA use a loop to jump into each of the folders for the three markers for all samples
```{bash engine.opts='-l'}
# move into the directory where you are going to run the DEICODE script
cd ../data/DEICODE
pwd


# Activate the conda qiime2 enviroment - check version
conda activate qiime2-2022.11

for folder in 18S  #for folder in 12S 18S COI
do
  #for data in phylo_obj phylo_obj_env phylo_obj_cb
  for data in oBiom_r_env

  do
  cd $folder
  #!/usr/bin/env bash

  # This bash script can be run with bash deicode_template_072820.sh
  # All of the input files (asv_table_for_deicode.txt, tax_table_for_deicode.txt, and metadata_for_deicode.txt) must be in the   same directory as this script.

  ### Description ###
  #
  # You will need qiime2 and DEICODE installed on your machine in order for it to work.
  # You will have four outputs: 1) distance.qza, 2) ordination.qza, 3) biplot.qza, and 4_ qurro_plot_deicoderank.qzv
  # Of these outputs, the most import one is the "ordination.qza" file - this is what you will use in R
  # biplot.qza can be visualized at view.qiime2.org, and is an interactive way to view your PCA.
  # qurro_plot_deicoderank.qzv will show you the loadings of your taxa on the different PCs.

  #Make biom file
    biom convert -i ./${data}_${folder}_asv_table_for_biom_${subset}.txt -o table.from_txt_json_${data}_${subset}.biom --table-type="OTU table" --to-json
  #add metadata files to biom file - change the merged_tax_table_for_biomm.txt and _merged_for_biom.txt files
   biom add-metadata -i table.from_txt_json_${data}_${subset}.biom -o table.w_md_${data}_${subset}.biom --observation-metadata-fp ${data}_${folder}_tax_table_for_biom_${subset}.txt --sample-metadata-fp ${data}_${folder}_metadata_for_biom_${subset}.txt

  #import into Qiime2 - Change output file
    qiime tools import \
     --input-path table.w_md_${data}_${subset}.biom \
     --output-path ${data}_${folder}_master_${subset}.biom.qza \
     --type FeatureTable[Frequency]

  #run DEICODE - change input table
    qiime deicode rpca \
      --i-table ${data}_${folder}_master_${subset}.biom.qza \
      --p-n-components 3 \
      --p-min-feature-count 10 \
      --p-min-sample-count 5000 \
      --o-biplot ordination_${data}_${subset}.qza \
      --o-distance-matrix distance_${data}_${subset}.qza

  ## Create biplot - change metadata files
    qiime emperor biplot \
      --i-biplot ordination_${data}_${subset}.qza \
      --m-sample-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --m-feature-metadata-file ${data}_${folder}_tax_table_for_biom_${subset}.txt \
      --o-visualization biplot_${data}_${subset}.qzv \
      --p-number-of-features 8

  # Create Qurro Visualization - change metadata files
    qiime qurro loading-plot \
      --i-ranks ordination_${data}_${subset}.qza \
      --i-table ${data}_${folder}_master_${subset}.biom.qza \
      --m-sample-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --m-feature-metadata-file ${data}_${folder}_tax_table_for_biom_${subset}.txt \
      --o-visualization qurro_plot_deicoderank_${data}_${subset}.qzv

   # Run a permanova | permdisp | or anosim test between whatever metadata column you choose (argument --m-metadata-column [column of choosing])
     qiime diversity beta-group-significance \
      --i-distance-matrix distance_${data}_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --m-metadata-column sequencer \
      --p-method anosim \
      --o-visualization sequencer_significance_anosim_${data}_${subset}.qzv

    # Alpha diversity test
    #Shannon
     qiime diversity alpha \
       --i-table ${data}_${folder}_master_${subset}.biom.qza \
       --p-metric shannon \
       --o-alpha-diversity ${data}_${folder}_shannon_alpha_${subset}.qza

      qiime diversity alpha-group-significance \
      --i-alpha-diversity ${data}_${folder}_shannon_alpha_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --o-visualization ${data}_${folder}_shannon_alpha_${subset}.qzv
      
      #Chao1
      qiime diversity alpha \
       --i-table ${data}_${folder}_master_${subset}.biom.qza \
       --p-metric chao1 \
       --o-alpha-diversity ${data}_${folder}_chao1_alpha_${subset}.qza

      qiime diversity alpha-group-significance \
      --i-alpha-diversity ${data}_${folder}_chao1_alpha_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --o-visualization ${data}_${folder}_chao1_alpha_${subset}.qzv
      
      #Observed OTUs
      qiime diversity alpha \
       --i-table ${data}_${folder}_master_${subset}.biom.qza \
       --p-metric observed_features \
       --o-alpha-diversity ${data}_${folder}_observed_otus_alpha_${subset}.qza

      qiime diversity alpha-group-significance \
      --i-alpha-diversity ${data}_${folder}_observed_otus_alpha_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --o-visualization ${data}_${folder}_observed_otus_alpha_${subset}.qzv
   done

  cd ../
done

```




#### RPCA of all samples, rarefied data
```{r}


#Import Qiime2 Results
file = paste(deicode_directory,"ordination_oBiom_r_env_",subset,".qza",sep="")
print(file)
pco<-read_qza(file)
pco$uuid
#look at data
head(pco$data$ProportionExplained)
pco$data$Vectors[1:5, 1:4]

#create proportion explained labels
label.PC1 <- paste("PC1: ", round(pco$data$ProportionExplained$PC1, 3)*100,"%")
label.PC1
label.PC2 <- paste("PC2: ", round(pco$data$ProportionExplained$PC2, 3)*100,"%")
label.PC2
label.PC3 <- paste("PC3: ", round(pco$data$ProportionExplained$PC3, 3)*100,"%")
label.PC3

#Join with sample data
pco$data$Vectors <- pco$data$Vectors %>% 
  rename(., "sample_name" = "SampleID" )

pcscores <- left_join(pco$data$Vectors, samp.c)


#format loading scores
loadings <- as.data.frame(pco$data$Species)
loadings$ASV <- loadings$FeatureID

#join on OTU, adding taxa info
loadings <- left_join(loadings, tax.c, by="ASV")

#export pcscores
file = paste(deicode_directory,"scores/pcscores_",marker,"_oBiom_r_env_Dada2_Qiime2_",subset,".csv",sep="")
print(file)
write.csv(pcscores, file)
file = paste(deicode_directory,"scores/loadings_",marker,"_oBiom_r_env_Dada2_Qiime2_",subset,".csv",sep="")
print(file)
write.csv(loadings, file)
pcscores[1:5, 1:9]  #long because of sample data
head(loadings)



# Plot RPCA ---------------------------------------------------------------



cols6 <- c("navyblue", "goldenrod3", "darkmagenta", "firebrick", "firebrick2", "darkolivegreen4", "violetred3")

pcscores %>% mutate(depth_bin = case_when(depth >= 0 & depth <= 20 ~ "0-20"), 
                        depth_bin = case_when(depth == 60 ~ "60", T ~ depth_bin), 
                        depth_bin = case_when(depth == 100 ~ "100", T ~ depth_bin), 
                        depth_bin = case_when(depth == 150 ~ "150", T ~ depth_bin), 
                        depth_bin = case_when(depth >= 200 & depth <= 700 ~ "200-700", T ~ depth_bin),
                        depth_bin = case_when(sample_source == "CN24F 150m exp" ~ "150 - benchtop experiment", T ~ depth_bin),
                        depth_bin = fct_relevel(depth_bin, "0-20", "60", "100", "150", "150 - benchtop experiment", "200-700"),
                        source_depth_bin = paste(sample_source, depth_bin, sep = " - "),
                        source_depth_bin = reorder(source_depth_bin, depth)) %>%  
  ggplot(aes(PC1,PC2, fill=depth_bin))+
  geom_line(aes(group = original_name, color = depth_bin), linetype = 2, alpha = 0.7) +
  geom_point(size=4, alpha=0.8, aes(shape = sequencer, color = depth_bin)) + 
  scale_shape_manual(values = c(21, 1)) +
    scale_fill_manual(values = cols6) +
    scale_color_manual(values = cols6, guide="none") +
  labs(x=label.PC1 , 
       y=label.PC2, 
       fill = "Depth Bin (m)",
       shape = "Sequencer",
       subtitle = "Comparison between sequencers\nLines connect sample replicates",
       title = "RPCA - rarefied data, all reads" )+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.text = element_text(size = 12, colour = "black"),
        legend.title = element_text(face = "bold", size=12),
        legend.key = element_rect(fill = FALSE, colour = FALSE),
        legend.key.size = unit(0.1,"line"),
        plot.title = element_text(size = 22),
        plot.subtitle = element_text(size = 18),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15),
  )+ guides(fill = guide_legend(override.aes = list(size=5, shape=21)),
            shape = guide_legend(override.aes = list(size = 5, shape = c(19, 1))))


# save as .png
filename = paste(fig2_directory, 'Figure2b_RPCA_by_depth_bin_',marker,"_",subset,'.png', sep='')
print('Plot of PC1PC2 colored by Year:')
filename
ggsave(filename,height = 9, width =10, units = 'in')

# save as .svg
filename = paste(fig2_directory, 'Figure2b_RPCA_by_depth_bin_',marker,"_",subset,'.svg', sep='')
print('Plot of PC1PC2 colored by Year:')
filename
ggsave(filename,height = 9, width =10, units = 'in')




```




## ## Figure 2 C - all samples, rarefied data, ASVs with 1 read in a sample removed

### create Phyloseq object of all  samples
```{r}

  
  subset = c("Rarefied_data_more_than_1_read")
  Sys.setenv(subset = subset)
  
#   metadata_updated <- samp.c %>% mutate(depth_bin = case_when(depth >= 0 & depth <= 20 ~ "0-20"), 
#                         depth_bin = case_when(depth == 60 ~ "60", T ~ depth_bin), 
#                         depth_bin = case_when(depth == 100 ~ "100", T ~ depth_bin), 
#                         depth_bin = case_when(depth == 150 ~ "150", T ~ depth_bin), 
#                         depth_bin = case_when(depth >= 200 & depth <= 700 ~ "200-700", T ~ depth_bin),
#                         depth_bin = case_when(sample_source == "CN24F 150m exp" ~ "150 - benchtop experiment", T ~ depth_bin),
#                         depth_bin = fct_relevel(depth_bin, "0-20", "60", "100", "150", "150 - benchtop experiment", "200-700"))
#   
# 
#   
# species_label <- tax.c %>% column_to_rownames(var = "ASV")
#   
#   rownames(metadata_updated) <- NULL
#   metadata_updated <- column_to_rownames(metadata_updated,"sample_name")
#   
#   more_than_1 <- as.matrix(read_csv(paste0(data_directory, proj_prefix, "_", marker, "_asv_filtered.csv")) %>% 
#     pivot_longer( -ASV, names_to ='sample_name',values_to = 'reads' ) %>%
#     filter(!reads <= 5) %>%
#     pivot_wider(names_from = "sample_name", values_from = "reads", values_fill = 0) %>%
#     column_to_rownames(var = "ASV"))
#   
#   # , row.names = 1,check.names = FALSE
#   
#   phylo <- merge_phyloseq(otu_table(more_than_1,taxa_are_rows = TRUE),
#                           tax_table(as.matrix(species_label)),
#                           sample_data(metadata_updated))
  
  
  # phylo <- merge_phyloseq(otu_table(as.matrix(read.csv(paste0(data_directory, proj_prefix, "_", marker, "_asv_filtered.csv"), row.names = 1,check.names = FALSE)),taxa_are_rows = TRUE),
  #                         tax_table(as.matrix(species_label)),
  #                         sample_data(metadata_updated))
  
  # phylo
  # 
  #   sample_data(phylo)
  # 
  # phylo_obj <- phylo
  # 
  # phylo_obj_env <- prune_samples(sample_data(phylo)$sample_type %in% c("environmental"), phylo) 
  # 
  # 
  # 
  # 
  # 
  # sample_data(phylo_obj)
  # sample_data(phylo_obj_env)
  # # sample_data(phylo_obj_cb)
  # 
  # # Create Tables -----------------------------------------------------------
  # 
  # #create vegan tables
  # vegan_matrix_env=vegan_otu(phylo_obj_env)
  # otu_tab_env=otu_table(phylo_obj_env)
  # tax_tab_env= tax_table(phylo_obj_env)
  # samp_dat_env <- sample_data(phylo_obj_env)
  # 
  # # Rarefy ------------------------------------------------------------------
  # #print(sample_sums(phylo_obj_env))
  # print(min(sample_sums(phylo_obj_env)))
  # print(max(sample_sums(phylo_obj_env)))
  # print(mean(sample_sums(phylo_obj_env)))
  # print(phylo_obj_env)
  # # Remove samples with <500 reads (like DEICODE)
  # oBiom_r_env = prune_samples(sample_sums(phylo_obj_env)>=5000, phylo_obj_env)
  # #set random seed for reproducibility later
  # print(min(sample_sums(oBiom_r_env)))
  # oBiom_r_env = rarefy_even_depth(oBiom_r_env, sample.size = min(sample_sums(oBiom_r_env)), rngseed = 678, replace = FALSE, trimOTUs = TRUE, verbose = TRUE)
  # print(oBiom_r_env)
  # assign(paste0("oBiom_r_env_",marker),oBiom_r_env)
  
  ##########
  ##########
  ##########
  ##########
  ### Take the rarefied data from earlier and now remove singleton ASVs from samples
  rotu_tab_env=otu_table(oBiom_r_env)
  
  as.data.frame(rotu_tab_env)

as.matrix(as.data.frame(rotu_tab_env) %>% 
            rownames_to_column(var = "ASV") %>%
    pivot_longer( -ASV, names_to ='sample_name',values_to = 'reads' ) %>%
    filter(!reads <= 1) %>%
    pivot_wider(names_from = "sample_name", values_from = "reads", values_fill = 0) %>%
    column_to_rownames(var = "ASV"))

phylo2 <- merge_phyloseq(otu_table(as.matrix(as.data.frame(rotu_tab_env) %>% 
            rownames_to_column(var = "ASV") %>%
    pivot_longer( -ASV, names_to ='sample_name',values_to = 'reads' ) %>%
    filter(reads > 1) %>%
    pivot_wider(names_from = "sample_name", values_from = "reads", values_fill = 0) %>%
    column_to_rownames(var = "ASV")),taxa_are_rows = TRUE),
                          tax_table(as.matrix(species_label)),
                          sample_data(metadata_updated))
phylo2
oBiom_r_env



oBiom_r_env <- phylo2


  
  #create matrixes
  rvegan_matrix_env=vegan_otu(oBiom_r_env)
  rotu_tab_env=otu_table(oBiom_r_env)
  rtax_tab_env= tax_table(oBiom_r_env)
  rsamp_dat_env <- rownames_to_column(as.data.frame(as.matrix(sample_data(oBiom_r_env))), var = "sample_name") #sample_data(oBiom_r_env)
  
  #save rarefied data
  # #save to csv file
  file = paste(rarefaction_data,marker,'_env_Non-Rarefied_asv_table_',subset,'.csv', sep="")
  write.csv(otu_tab_env, file)
  file = paste(rarefaction_data,marker,'_env_Non-Rarefied_taxa_table_',subset,'.csv', sep="")
  write.csv(tax_tab_env, file)
  # file = paste(rarefaction_data,marker,'_env_Non-Rarefied_meta_table_',subset,'.csv', sep="")
  # write.csv(samp_dat_env, file)

  file = paste(rarefaction_data,marker,'_env_Rarefied_asv_table_',subset,'.csv', sep="")
  write.csv(rotu_tab_env, file)
  file = paste(rarefaction_data,marker,'_env_Rarefied_taxa_table_',subset,'.csv', sep="")
  write.csv(rtax_tab_env, file)
  file = paste(rarefaction_data,marker,'_env_Rarefied_meta_table_',subset,'.csv', sep="")
  write.csv(rsamp_dat_env, file)
  
  
  ### DEICODE PCA 
  # rarefied
  phylo_obj_env_asv_table <- as.data.frame(otu_table(oBiom_r_env))
  phylo_obj_env_asv_table <- rownames_to_column(phylo_obj_env_asv_table,"#OTUID")
  write.table(phylo_obj_env_asv_table,paste0(deicode_directory,"/oBiom_r_env_",marker,"_asv_table_for_biom_",subset,".txt"),sep = "\t",row.names = FALSE,quote = FALSE)
 
  phylo_obj_env_tax_table <- as.data.frame(as(tax_table(oBiom_r_env),"matrix"))
  phylo_obj_env_tax_table <- rownames_to_column(phylo_obj_env_tax_table,"#OTUID")
  write.table(phylo_obj_env_tax_table,paste0(deicode_directory,"/oBiom_r_env_",marker,"_tax_table_for_biom_",subset,".txt"),sep = "\t",row.names = FALSE,quote = FALSE)
  
  phylo_obj_env_metadata <- as.data.frame(as.matrix(sample_data(oBiom_r_env)))
  phylo_obj_env_metadata <- rownames_to_column(phylo_obj_env_metadata,"#SampleID")
  write.table(phylo_obj_env_metadata,paste0(deicode_directory,"/oBiom_r_env_",marker,"_metadata_for_biom_",subset,".txt"),sep = "\t",row.names = FALSE,quote = FALSE)
  

  
```

### DEICODE
#### deicode PCA use a loop to jump into each of the folders for the three markers for all samples
```{bash engine.opts='-l'}
# move into the directory where you are going to run the DEICODE script
cd ../data/DEICODE
pwd


# Activate the conda qiime2 enviroment - check version
conda activate qiime2-2022.11

for folder in 18S  #for folder in 12S 18S COI
do
  #for data in phylo_obj phylo_obj_env phylo_obj_cb
  for data in oBiom_r_env

  do
  cd $folder
  #!/usr/bin/env bash

  # This bash script can be run with bash deicode_template_072820.sh
  # All of the input files (asv_table_for_deicode.txt, tax_table_for_deicode.txt, and metadata_for_deicode.txt) must be in the   same directory as this script.

  ### Description ###
  #
  # You will need qiime2 and DEICODE installed on your machine in order for it to work.
  # You will have four outputs: 1) distance.qza, 2) ordination.qza, 3) biplot.qza, and 4_ qurro_plot_deicoderank.qzv
  # Of these outputs, the most import one is the "ordination.qza" file - this is what you will use in R
  # biplot.qza can be visualized at view.qiime2.org, and is an interactive way to view your PCA.
  # qurro_plot_deicoderank.qzv will show you the loadings of your taxa on the different PCs.

  #Make biom file
    biom convert -i ./${data}_${folder}_asv_table_for_biom_${subset}.txt -o table.from_txt_json_${data}_${subset}.biom --table-type="OTU table" --to-json
  #add metadata files to biom file - change the merged_tax_table_for_biomm.txt and _merged_for_biom.txt files
   biom add-metadata -i table.from_txt_json_${data}_${subset}.biom -o table.w_md_${data}_${subset}.biom --observation-metadata-fp ${data}_${folder}_tax_table_for_biom_${subset}.txt --sample-metadata-fp ${data}_${folder}_metadata_for_biom_${subset}.txt

  #import into Qiime2 - Change output file
    qiime tools import \
     --input-path table.w_md_${data}_${subset}.biom \
     --output-path ${data}_${folder}_master_${subset}.biom.qza \
     --type FeatureTable[Frequency]

  #run DEICODE - change input table
    qiime deicode rpca \
      --i-table ${data}_${folder}_master_${subset}.biom.qza \
      --p-n-components 3 \
      --p-min-feature-count 10 \
      --p-min-sample-count 5000 \
      --o-biplot ordination_${data}_${subset}.qza \
      --o-distance-matrix distance_${data}_${subset}.qza

  ## Create biplot - change metadata files
    qiime emperor biplot \
      --i-biplot ordination_${data}_${subset}.qza \
      --m-sample-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --m-feature-metadata-file ${data}_${folder}_tax_table_for_biom_${subset}.txt \
      --o-visualization biplot_${data}_${subset}.qzv \
      --p-number-of-features 8

  # Create Qurro Visualization - change metadata files
    qiime qurro loading-plot \
      --i-ranks ordination_${data}_${subset}.qza \
      --i-table ${data}_${folder}_master_${subset}.biom.qza \
      --m-sample-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --m-feature-metadata-file ${data}_${folder}_tax_table_for_biom_${subset}.txt \
      --o-visualization qurro_plot_deicoderank_${data}_${subset}.qzv

   # Run a permanova | permdisp | or anosim test between whatever metadata column you choose (argument --m-metadata-column [column of choosing])
     qiime diversity beta-group-significance \
      --i-distance-matrix distance_${data}_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --m-metadata-column sequencer \
      --p-method anosim \
      --o-visualization sequencer_significance_anosim_${data}_${subset}.qzv

    # Alpha diversity test
    #Shannon
     qiime diversity alpha \
       --i-table ${data}_${folder}_master_${subset}.biom.qza \
       --p-metric shannon \
       --o-alpha-diversity ${data}_${folder}_shannon_alpha_${subset}.qza

      qiime diversity alpha-group-significance \
      --i-alpha-diversity ${data}_${folder}_shannon_alpha_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --o-visualization ${data}_${folder}_shannon_alpha_${subset}.qzv
      
      #Chao1
      qiime diversity alpha \
       --i-table ${data}_${folder}_master_${subset}.biom.qza \
       --p-metric chao1 \
       --o-alpha-diversity ${data}_${folder}_chao1_alpha_${subset}.qza

      qiime diversity alpha-group-significance \
      --i-alpha-diversity ${data}_${folder}_chao1_alpha_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --o-visualization ${data}_${folder}_chao1_alpha_${subset}.qzv
      
      #Observed OTUs
      qiime diversity alpha \
       --i-table ${data}_${folder}_master_${subset}.biom.qza \
       --p-metric observed_features \
       --o-alpha-diversity ${data}_${folder}_observed_otus_alpha_${subset}.qza

      qiime diversity alpha-group-significance \
      --i-alpha-diversity ${data}_${folder}_observed_otus_alpha_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --o-visualization ${data}_${folder}_observed_otus_alpha_${subset}.qzv
   done

  cd ../
done

```




#### RPCA of all samples, rarefied data, ASVs with 1 read in a sample removed
```{r}


#Import Qiime2 Results
file = paste(deicode_directory,"ordination_oBiom_r_env_",subset,".qza",sep="")
print(file)
pco<-read_qza(file)
pco$uuid
#look at data
head(pco$data$ProportionExplained)
pco$data$Vectors[1:5, 1:4]

#create proportion explained labels
label.PC1 <- paste("PC1: ", round(pco$data$ProportionExplained$PC1, 3)*100,"%")
label.PC1
label.PC2 <- paste("PC2: ", round(pco$data$ProportionExplained$PC2, 3)*100,"%")
label.PC2
label.PC3 <- paste("PC3: ", round(pco$data$ProportionExplained$PC3, 3)*100,"%")
label.PC3

#Join with sample data
pco$data$Vectors <- pco$data$Vectors %>% 
  rename(., "sample_name" = "SampleID" )

pcscores <- left_join(pco$data$Vectors, samp.c)


#format loading scores
loadings <- as.data.frame(pco$data$Species)
loadings$ASV <- loadings$FeatureID

#join on OTU, adding taxa info
loadings <- left_join(loadings, tax.c, by="ASV")

#export pcscores
file = paste(deicode_directory,"scores/pcscores_",marker,"_oBiom_r_env_Dada2_Qiime2_",subset,".csv",sep="")
print(file)
write.csv(pcscores, file)
file = paste(deicode_directory,"scores/loadings_",marker,"_oBiom_r_env_Dada2_Qiime2_",subset,".csv",sep="")
print(file)
write.csv(loadings, file)
pcscores[1:5, 1:9]  #long because of sample data
head(loadings)


# Plot RPCA ---------------------------------------------------------------



cols6 <- c("navyblue", "goldenrod3", "darkmagenta", "firebrick", "firebrick2", "darkolivegreen4", "violetred3")

pcscores %>% mutate(depth_bin = case_when(depth >= 0 & depth <= 20 ~ "0-20"), 
                        depth_bin = case_when(depth == 60 ~ "60", T ~ depth_bin), 
                        depth_bin = case_when(depth == 100 ~ "100", T ~ depth_bin), 
                        depth_bin = case_when(depth == 150 ~ "150", T ~ depth_bin), 
                        depth_bin = case_when(depth >= 200 & depth <= 700 ~ "200-700", T ~ depth_bin),
                        depth_bin = case_when(sample_source == "CN24F 150m exp" ~ "150 - benchtop experiment", T ~ depth_bin),
                        depth_bin = fct_relevel(depth_bin, "0-20", "60", "100", "150", "150 - benchtop experiment", "200-700"),
                        source_depth_bin = paste(sample_source, depth_bin, sep = " - "),
                        source_depth_bin = reorder(source_depth_bin, depth)) %>%  #merge( aggregate(cbind(mean.x=PC1, mean.y=PC2)~sequencer,pcscores,mean),by="sequencer") %>%
  ggplot(aes(PC1,PC2, fill=depth_bin))+
  geom_line(aes(group = original_name, color = depth_bin), linetype = 2, alpha = 0.7) +
  geom_point(size=4, alpha=0.8, aes(shape = sequencer, color = depth_bin)) + 
  # stat_ellipse(aes(group = depth_bin, color = depth_bin)) +
  # geom_point(aes(x=mean.x,y=mean.y),size=6)+
  # geom_segment(aes( x=mean.x, y=mean.y, xend=PC1, yend=PC2)) +
  # geom_segment(aes(x=mean.x, y=mean.y, xend=PC1, yend=PC2)) +
  scale_shape_manual(values = c(21, 1)) +
    scale_fill_manual(values = cols6) +
    scale_color_manual(values = cols6, guide="none") +
  # scale_color_tableau(palette = "Tableau 20") +
  # scale_color_viridis(option = "G", direction = 1, discrete = T, begin=.1, end = .9) +
  labs(x=label.PC1 , 
       y=label.PC2, 
       fill = "Depth Bin (m)",
       shape = "Sequencer",
       subtitle = "Comparison between sequencers\nLines connect sample replicates",
       title = "RPCA - rarefied data, ASVs with 1 read in a sample removed")+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.text = element_text(size = 12, colour = "black"),
        legend.title = element_text(face = "bold", size=12),
        legend.key = element_rect(fill = FALSE, colour = FALSE),
        legend.key.size = unit(0.1,"line"),
        plot.title = element_text(size = 22),
        plot.subtitle = element_text(size = 18),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15),
  )+ guides(fill = guide_legend(override.aes = list(size=5, shape=21)),
            shape = guide_legend(override.aes = list(size = 5, shape = c(19, 1))))


# save as .png
filename = paste(fig2_directory, 'Figure2c_RPCA_by_depth_bin_',marker,"_",subset,'.png', sep='')
print('Plot of PC1PC2 colored by Year:')
filename
ggsave(filename,height = 9, width =10, units = 'in')

# save as .svg
filename = paste(fig2_directory, 'Figure2c_RPCA_by_depth_bin_',marker,"_",subset,'.svg', sep='')
print('Plot of PC1PC2 colored by Year:')
filename
ggsave(filename,height = 9, width =10, units = 'in')


```



