---
title: "Figure3_sample_statistics_and_alpha_diversity"
author: "Jacoby Baker"
date: "2026-03-23"
output: html_document
---


# Set up directories and variables
```{r}
# setwd('scripts/')
proj_prefix <- "AVITI_MiSeq_comparison"
proj_name <- "AVITI VS MiSeq Comparison"

data_directory <- paste0('../data/filtered_seq_data/')
rarefaction_data <- paste0('../data/rarefaction_data/')
diversity_directory <- paste0("../figures/diversity/")
deicode_directory <- paste0('../data/DEICODE/18S/')

getwd()
fig3_directory <- paste0("../figures/Fig3/")

marker = "18S"
```



# Load libraries
```{r}

#load libraries
library(readr) #read csv files
# library(vegan3d)

library(dplyr)
library(ggplot2)
library(ggthemes)
library(magrittr)
library(RColorBrewer) #colors for plotting
# library("viridis") # for color fill 


library(tidyverse)
library(phyloseq)

library(ggpubr) # for ggboxplot

# library(taxa)
# library(metacoder)
# library(ggpubr) # for ggarrange
# library(reshape2) # for presence/absence fig


library(ggnewscale) # multiple fill color scales

library(cowplot) # for getting legend
# library(ggh4x) # for facet nested wrap

# library(ggbreak)

```


# # Non rarefied data

## Load in Data
```{r}

#ASV table
otu.c <- read_csv(paste0(data_directory, proj_prefix, "_", marker, "_asv_filtered.csv"))
otu.c

#taxa table
tax.c <- read_csv(paste0(data_directory, proj_prefix, "_", marker, "_taxa_filtered.csv"))
tax.c

#metadata table
samp.c <- read_csv(paste0(data_directory, proj_prefix, "_", marker, "_meta_filtered.csv")) %>%
  mutate(sample_source = fct_relevel(sample_source,  "CN23 MARS" , "MBTS 17024", "MBTS 20524", "MBTS 22624", "MBTS 25424", "CN24F 150m exp")) %>%
  mutate(short_name = "") %>%
  mutate(short_name = case_when(grepl("eDNA", sample_name) ~ substr(sample_name,1,nchar(sample_name)-9), T ~ short_name)) %>%
  mutate(short_name = case_when(grepl("CN24F150", sample_name) ~ substr(sample_name,14,nchar(sample_name)-3), T ~ short_name))
  
samp.c

seq.c <- read_csv(paste0(data_directory, proj_prefix, "_", marker, "_seq_filtered.csv"))
seq.c

#OTU table long format with percent total reads
potu.c <- otu.c %>%
  tidyr::pivot_longer( -ASV, names_to ='sample_name',values_to = 'reads' ) %>%
  ungroup() %>%
  group_by(sample_name) %>%
  mutate(per_tot = reads / sum(reads) *100) %>% #proportions(reads)*100
  arrange(-reads)
head(potu.c)


```



## set lowest taxonomic annotation level for ASVs
```{r}

  species_label <- tax.c  %>%
  # assign  algae
  mutate(Phylum = case_when(Class == "Raphidophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Phaeophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Pelagophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Dictyochophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Chrysophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Synurophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Bolidophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Eustigmatophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  # assign oomycotes
  mutate(Phylum = case_when(Class == "Oomycota" ~ "Oomycota", T ~ Phylum))  %>%
  mutate(Phylum = case_when(Class == "Oomycetes" ~ "Oomycota", T ~ Phylum))%>%
  # assign microeuk
  mutate(Phylum = case_when(Class == "Bigyra" ~ "Bigyra", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Choanoflagellata" ~ "Choanoflagellata", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Cryptophyceae" ~ "Cryptista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Order == "Telonemida" ~ "Cryptista", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Haptophyta" ~ "Haptista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Centroplasthelida" ~ "Haptista", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Ichthyosporea" ~ "Choanozoa", T ~ Phylum))  %>% 
  
  mutate(Class = case_when(Phylum == "Imbricatea" ~ "Imbricatea", T ~ Class)) %>%
  mutate(Phylum = case_when(Class == "Imbricatea" ~ "Cercozoa", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Polycystinea" ~ "Retaria", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Acantharea" ~ "Retaria", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Breviatea" ~ "Sulcozoa", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Aphelidea" ~ "Aphelida", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Filasterea" ~ "Choanozoa", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Phylum %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Phylum)) %>% 
  mutate(Class = case_when(Class %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Class)) %>% 
  mutate(Order = case_when(Order %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Order)) %>% 
  mutate(Family = case_when(Family %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Family)) %>% 
  mutate(Genus = case_when(Genus %in% c("unknown", "unassigned", "no_hit", "g_") ~ "unknown", T ~ Genus)) %>% 
  mutate(Species = case_when(Species %in% c("unknown", "unassigned", "no_hit", "s_") ~ "unknown", T ~ Species))


  species_label


```




## Compile data - 
```{r}

# environmental data
env_data <- inner_join(samp.c, potu.c, by = c("sample_name")) %>% 
  left_join(species_label)

env_data 


```



```{r}


# all_data <- env_data %>% 
#   select(sample_name, PlateID, SAMPLING_station, original_name, sequencer, depth_round, source_depth, ASV, reads, per_tot, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
#   filter(reads > 0)
# 
# all_data



```


## FIGURE 3 A - number of reads per sampler/sequencer - non rarefied data
### Number of reads per sample
```{r}

cols <- c("deepskyblue4", "darkolivegreen4")
pd <- position_dodge(0.4) # dodge position for aligning geom_line with geom_jitter
cols6 <- c("navyblue", "goldenrod3", "darkmagenta", "firebrick", "firebrick2", "darkolivegreen4", "violetred3")



# number of reads per sample - samples connected
reads_data <- env_data %>% select(sample_name, SAMPLING_station, depth, original_name,sample_source, ASV, reads, sequencer, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  mutate(depth_bin = case_when(depth >= 0 & depth <= 20 ~ "0-20"), 
                        depth_bin = case_when(depth == 60 ~ "60", T ~ depth_bin), 
                        depth_bin = case_when(depth == 100 ~ "100", T ~ depth_bin), 
                        depth_bin = case_when(depth == 150 ~ "150", T ~ depth_bin), 
                        depth_bin = case_when(depth >= 200 & depth <= 700 ~ "200-700", T ~ depth_bin),
                        depth_bin = case_when(sample_source == "CN24F 150m exp" ~ "150 - benchtop experiment", T ~ depth_bin),
                        depth_bin = fct_relevel(depth_bin, "0-20", "60", "100", "150", "150 - benchtop experiment", "200-700"),
                        source_depth_bin = paste(sample_source, depth_bin, sep = " - "),
         source_depth_bin = paste(sample_source, depth_bin, sep = " - "),
         source_depth_bin = reorder(source_depth_bin, depth),
         SAMPLING_station = case_when(grepl("c1", SAMPLING_station) ~ "C1", T ~ SAMPLING_station),
         SAMPLING_station = case_when(grepl("MOORING1|Mooring1|MOORING 1", SAMPLING_station) ~ "M1", T ~ SAMPLING_station),
         SAMPLING_station = case_when(grepl("MOORING2|Mooring2|MOORING 2", SAMPLING_station) ~ "M2", T ~ SAMPLING_station),
         SAMPLING_station = case_when(grepl("CN24F", sample_name) ~ "Benchtop ESP", T ~ SAMPLING_station),
         station_depth = paste(SAMPLING_station, depth, sep = " - "), 
         station_depth = reorder(station_depth, depth),
         station_depth_bin = paste(SAMPLING_station, depth_bin, sep = " - "),
         station_depth_bin = reorder(station_depth_bin, depth)) %>%
  filter(reads > 0) %>%
  mutate(count = 1) %>%
  group_by(sample_name) %>%
  mutate(sample_reads = sum(reads)) %>%
  ungroup() %>%
  distinct(sample_name, .keep_all = T) %>%
  mutate(sequencer = fct_relevel(sequencer, "MiSeq", "AVITI")) 

 fig3a_with_legend <- reads_data %>% ggplot(aes(x=sequencer, y = sample_reads, group = original_name)) +
  geom_line(position = pd, aes(color = depth_bin), alpha = .5, show.legend = F) +
  scale_color_manual(values = cols6) +
  new_scale_color() +
  geom_violin(aes(group = sequencer), alpha = 0, linewidth = 1) + #color = sequencer, 
  geom_point(position = pd, aes(fill = depth_bin), shape = 21, size = 3) +
  scale_fill_manual(values = cols6) +
  scale_y_log10(breaks = c(50000, 100000, 200000, 300000, 400000, 500000, 600000, 700000, 800000, 900000), labels = scales::label_comma()) +
  labs(
    subtitle = get_test_label(kruskal_test(sample_reads ~ sequencer, data = reads_data), detailed = TRUE),
    fill = "Depth Bin (m)",
    y = "Number of reads",
    x = "Sequencer",
    title = "Reads per sample - non rarefied"
    )+
    theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        plot.title = element_text(size = 24, face = "bold"), 
        plot.subtitle = element_text(size = 20), 
        plot.caption = element_blank(), 
        legend.text =  element_text(size = 20, colour = "black"), 
        legend.title =  element_text(face = "bold", size=23), 
        axis.text = element_text(size = 20),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24, face = "bold")
        ) +
  guides(fill=guide_legend(override.aes = list(size=7)))
  
fig3a_with_legend

# remove legend for figure panels
 fig3a_no_legend <- fig3a_with_legend +
    theme(
      legend.position = "none",
    )
  fig3a_no_legend


filename = paste(fig3_directory, 'Figure3a_scatter_plot_reads_per_sample_non_rarefied_',proj_prefix,'_',marker,'.png', sep='')
  print(marker)
  filename
  ggsave(filename,height = 10, width =8, units = 'in')
  
  filename = paste(fig3_directory, 'Figure3a_scatter_plot_reads_per_sample_non_rarefied_',proj_prefix,'_',marker,'.svg', sep='')
  print(marker)
  filename
  ggsave(filename,height = 10, width =8, units = 'in')

###########  
 
  # pull out legend and plot for figure panels
  fig3_legend <- get_legend(fig3a_with_legend)
  fig3_leg <- plot_grid(fig3_legend, rel_heights = c(2, 0.5))
  fig3_leg
  
  filename = paste(fig3_directory, 'Figure3_legend_',proj_prefix,'_',marker,'.png', sep='')
  print(marker)
  filename
  ggsave(filename,height = 6, width =8, units = 'in')
  
  filename = paste(fig3_directory, 'Figure3_legend_',proj_prefix,'_',marker,'.svg', sep='')
  print(marker)
  filename
  ggsave(filename,height = 6, width =8, units = 'in')

```



## FIGURE 3 B - number of ASVs per sampler/sequencer - non rarefied data
### Number of ASVs per sample
## Violin plots of unique ASVs and taxa per sequencer

```{r}

pd <- position_dodge(0.4) # dodge position for aligning geom_line with geom_jitter
cols6 <- c("navyblue", "goldenrod3", "darkmagenta", "firebrick", "firebrick2", "darkolivegreen4", "violetred3")





# ASV level
asv_data <- env_data %>% select(sample_name, SAMPLING_station, depth, original_name,sample_source, ASV, reads, sequencer, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>% mutate(depth_bin = case_when(depth >= 0 & depth <= 20 ~ "0-20"), 
                        depth_bin = case_when(depth == 60 ~ "60", T ~ depth_bin), 
                        depth_bin = case_when(depth == 100 ~ "100", T ~ depth_bin), 
                        depth_bin = case_when(depth == 150 ~ "150", T ~ depth_bin), 
                        depth_bin = case_when(depth >= 200 & depth <= 700 ~ "200-700", T ~ depth_bin),
                        depth_bin = case_when(sample_source == "CN24F 150m exp" ~ "150 - benchtop experiment", T ~ depth_bin),
                        depth_bin = fct_relevel(depth_bin, "0-20", "60", "100", "150", "150 - benchtop experiment", "200-700"),
                        source_depth_bin = paste(sample_source, depth_bin, sep = " - "),
                        source_depth_bin = reorder(source_depth_bin, depth),
         source_depth_bin = paste(sample_source, depth_bin, sep = " - "),
         source_depth_bin = reorder(source_depth_bin, depth),
         SAMPLING_station = case_when(grepl("c1", SAMPLING_station) ~ "C1", T ~ SAMPLING_station),
         SAMPLING_station = case_when(grepl("MOORING1|Mooring1|MOORING 1", SAMPLING_station) ~ "M1", T ~ SAMPLING_station),
         SAMPLING_station = case_when(grepl("MOORING2|Mooring2|MOORING 2", SAMPLING_station) ~ "M2", T ~ SAMPLING_station),
         SAMPLING_station = case_when(grepl("CN24F", sample_name) ~ "Benchtop ESP", T ~ SAMPLING_station),
         station_depth = paste(SAMPLING_station, depth, sep = " - "), 
         station_depth = reorder(station_depth, depth),
         station_depth_bin = paste(SAMPLING_station, depth_bin, sep = " - "),
         station_depth_bin = reorder(station_depth_bin, depth)) %>%
  filter(reads > 0) %>%
  mutate(count = 1) %>%
  group_by(sample_name) %>%
  mutate(ASV_count = sum(count)) %>%
  ungroup() %>%
  distinct(sample_name, .keep_all = T) %>%
  mutate(sequencer = fct_relevel(sequencer, "MiSeq", "AVITI")) 
  
  asv_data %>% ggplot(aes(x=sequencer, y = ASV_count, group = original_name)) +
  geom_line(position = pd, aes(color = depth_bin), alpha = .5, show.legend = F) +
  scale_color_manual(values = cols6) +
  new_scale_color() +
  geom_violin(aes(group = sequencer), alpha = 0, linewidth = 1) + 
  geom_point(position = pd, aes(fill = depth_bin), shape = 21, size = 3) +
  scale_fill_manual(values = cols6)  +
  labs(
    subtitle = get_test_label(kruskal_test(ASV_count ~ sequencer, data = asv_data), detailed = TRUE),
    color = "Depth Bin (m)",
    y = "Number of ASVs",
    x = "Sequencer",
    fill = "Depth Bin (m)",
    title = "ASVs per sample - non rarefied"
    )+
    theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        plot.title = element_text(size = 24, face = "bold"), 
        plot.subtitle = element_text(size = 20), 
        plot.caption = element_blank(), 
        legend.text =  element_text(size = 20, colour = "black"),
        legend.title =  element_text(face = "bold", size=23),
        legend.position = "none",
        axis.text = element_text(size = 20),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24, face = "bold")
        ) +
  guides(fill=guide_legend(override.aes = list(size=7)))


filename = paste(fig3_directory, 'Figure3b_scatter_plot_ASVs_per_sample_non_rarefied_',proj_prefix,'_',marker,'.png', sep='')
  print(marker)
  filename
  ggsave(filename,height = 10, width =8, units = 'in')
  
  filename = paste(fig3_directory, 'Figure3b_scatter_plot_ASVs_per_sample_non_rarefied_',proj_prefix,'_',marker,'.svg', sep='')
  print(marker)
  filename
  ggsave(filename,height = 10, width =8, units = 'in')
  

```


# #########
# # Rarefied data


## Load in Data
```{r}

#ASV table
otu.c.rarefied <- read_csv(paste0(rarefaction_data, marker, "_Rarefied_data_OTU_table.csv"))
otu.c.rarefied

#taxa table
tax.c.rarefied <- read_csv(paste0(rarefaction_data, marker, "_Rarefied_data_taxa_table.csv")) %>%
  # assign  algae
  mutate(Phylum = case_when(Class == "Raphidophyceae" ~ "Gyrista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Phaeophyceae" ~ "Gyrista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Pelagophyceae" ~ "Gyrista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Dictyochophyceae" ~ "Gyrista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Chrysophyceae" ~ "Gyrista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Synurophyceae" ~ "Gyrista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Bolidophyceae" ~ "Gyrista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Eustigmatophyceae" ~ "Gyrista", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Oomycota" ~ "Oomycota", T ~ Phylum))  %>%
  mutate(Phylum = case_when(Class == "Oomycetes" ~ "Oomycota", T ~ Phylum))%>%
  
  mutate(Phylum = case_when(Class == "Bigyra" ~ "Bigyra", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Choanoflagellata" ~ "Choanoflagellata", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Cryptophyceae" ~ "Cryptophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Order == "Telonemida" ~ "Cryptophyta", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Haptophyta" ~ "Haptista", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Ichthyosporea" ~ "Choanozoa", T ~ Phylum))  %>% 
  
  mutate(Class = case_when(Phylum == "Imbricatea" ~ "Imbricatea", T ~ Class)) %>%
  mutate(Phylum = case_when(Class == "Imbricatea" ~ "Cercozoa", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Phylum %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Phylum)) %>% 
  mutate(Class = case_when(Class %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Class)) %>% 
  mutate(Order = case_when(Order %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Order)) %>% 
  mutate(Family = case_when(Family %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Family)) %>% 
  mutate(Genus = case_when(Genus %in% c("unknown", "unassigned", "no_hit", "g_") ~ "unknown", T ~ Genus)) %>% 
  mutate(Species = case_when(Species %in% c("unknown", "unassigned", "no_hit", "s_") ~ "unknown", T ~ Species))
tax.c.rarefied

#metadata table
samp.c.rarefied <- read_csv(paste0(rarefaction_data, marker, "_Rarefied_data_metadata_table.csv"))  %>%
  mutate(sample_source = fct_relevel(sample_source,  "CN23 MARS" , "MBTS 17024", "MBTS 20524", "MBTS 22624", "MBTS 25424", "CN24F 150m exp")) %>%
  mutate(short_name = "") %>%
  mutate(short_name = case_when(grepl("eDNA", sample_name) ~ substr(sample_name,1,nchar(sample_name)-9), T ~ short_name)) %>%
  mutate(short_name = case_when(grepl("CN24F150", sample_name) ~ substr(sample_name,14,nchar(sample_name)-3), T ~ short_name))
samp.c.rarefied



#OTU table long format with percent total reads
potu.c.rarefied <- otu.c.rarefied %>%
  tidyr::pivot_longer( -ASV, names_to ='sample_name',values_to = 'reads' ) %>%
  ungroup() %>%
  group_by(sample_name) %>%
  mutate(per_tot = reads / sum(reads) *100) %>% #proportions(reads)*100
  arrange(-reads)
head(potu.c.rarefied)


```



## Compile data - 
```{r}

# environmental data
env_data_rarefied <- inner_join(samp.c.rarefied, potu.c.rarefied, by = c("sample_name")) %>% 
  left_join(tax.c.rarefied)

env_data_rarefied 


```


## FIGURE 3 C - number of ASVs per sample - rarefied data
```{r}

cols <- c("deepskyblue4", "darkolivegreen4")
pd <- position_dodge(0.4) # dodge position for aligning geom_line with geom_jitter
cols6 <- c("navyblue", "goldenrod3", "darkmagenta", "firebrick", "firebrick2", "darkolivegreen4", "violetred3")


# ASV level
asv_data <- env_data_rarefied %>% select(sample_name, SAMPLING_station, depth, original_name,sample_source, ASV, reads, sequencer, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  mutate(depth_bin = case_when(depth >= 0 & depth <= 20 ~ "0-20"), 
                        depth_bin = case_when(depth == 60 ~ "60", T ~ depth_bin), 
                        depth_bin = case_when(depth == 100 ~ "100", T ~ depth_bin), 
                        depth_bin = case_when(depth == 150 ~ "150", T ~ depth_bin), 
                        depth_bin = case_when(depth >= 200 & depth <= 700 ~ "200-700", T ~ depth_bin),
                        depth_bin = case_when(sample_source == "CN24F 150m exp" ~ "150 - benchtop experiment", T ~ depth_bin),
                        depth_bin = fct_relevel(depth_bin, "0-20", "60", "100", "150", "150 - benchtop experiment", "200-700"),
                        source_depth_bin = paste(sample_source, depth_bin, sep = " - "),
                        source_depth_bin = reorder(source_depth_bin, depth),
         source_depth_bin = paste(sample_source, depth_bin, sep = " - "),
         source_depth_bin = reorder(source_depth_bin, depth),
         SAMPLING_station = case_when(grepl("c1", SAMPLING_station) ~ "C1", T ~ SAMPLING_station),
         SAMPLING_station = case_when(grepl("MOORING1|Mooring1|MOORING 1", SAMPLING_station) ~ "M1", T ~ SAMPLING_station),
         SAMPLING_station = case_when(grepl("MOORING2|Mooring2|MOORING 2", SAMPLING_station) ~ "M2", T ~ SAMPLING_station),
         SAMPLING_station = case_when(grepl("CN24F", sample_name) ~ "Benchtop ESP", T ~ SAMPLING_station),
         station_depth = paste(SAMPLING_station, depth, sep = " - "), 
         station_depth = reorder(station_depth, depth),
         station_depth_bin = paste(SAMPLING_station, depth_bin, sep = " - "),
         station_depth_bin = reorder(station_depth_bin, depth)) %>%
  filter(reads > 0) %>%
  mutate(count = 1) %>%
  group_by(sample_name) %>%
  mutate(ASV_count = sum(count)) %>%
  ungroup() %>%
  distinct(sample_name, .keep_all = T) %>%
  mutate(sequencer = fct_relevel(sequencer, "MiSeq", "AVITI")) 

  asv_data %>% ggplot(aes(x=sequencer, y = ASV_count, group = original_name)) +
  geom_line(position = pd, aes(color = depth_bin), show.legend = F) +
  scale_color_manual(values = cols6) +
  new_scale_color() +
  geom_violin(aes(group = sequencer), alpha = 0, linewidth = 1) + 
  geom_point(position = pd, aes(fill = depth_bin), shape = 21, size = 3) +
  scale_fill_manual(values = cols6) +
  labs(
    subtitle = get_test_label(kruskal_test(ASV_count ~ sequencer, data = asv_data), detailed = TRUE),
    color = "Depth Bin (m)",
    y = "Number of ASVs per sample",
    x = "Sequencer",
    fill = "Depth Bin (m)",
    title = "ASVs per sample - rarefied"
    )+
    theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        plot.title = element_text(size = 24, face = "bold"), 
        plot.subtitle = element_text(size = 20), 
        plot.caption = element_blank(), 
        legend.text =  element_text(size = 20, colour = "black"),
        legend.title =  element_text(face = "bold", size=23),
        legend.position = "none",
        axis.text = element_text(size = 20),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24, face = "bold")
        ) +
  guides(fill=guide_legend(override.aes = list(size=7)))
  

filename = paste(fig3_directory, 'Figure3c_scatter_plot_ASVs_per_sample_rarefied_',proj_prefix,'_',marker,'.png', sep='')
  print(marker)
  filename
  ggsave(filename,height = 10, width =8, units = 'in')
  
  filename = paste(fig3_directory, 'Figure3c_scatter_plot_ASVs_per_sample_rarefied_',proj_prefix,'_',marker,'.svg', sep='')
  print(marker)
  filename
  ggsave(filename,height = 10, width =8, units = 'in')
  

```



# #ASVs with 1 read removed per sample
## Compile data - 
```{r}

subset <- "ASVs_with_more_than_1_read_per_sample"


potu.c.rarefied <- otu.c.rarefied %>%
  tidyr::pivot_longer( -ASV, names_to ='sample_name',values_to = 'reads' ) %>%
    filter(!reads <= 1) %>%
  ungroup() %>%
  group_by(sample_name) %>%
  mutate(per_tot = reads / sum(reads) *100) %>% #proportions(reads)*100
  arrange(-reads)
head(potu.c.rarefied)

   



# environmental data
env_data_rarefied <- inner_join(samp.c.rarefied, potu.c.rarefied, by = c("sample_name")) %>% 
  left_join(tax.c.rarefied)

env_data_rarefied 


```




## FIGURE 3 D - number of ASVs per sample 
```{r}

cols <- c("deepskyblue4", "darkolivegreen4")
pd <- position_dodge(0.4) # dodge position for aligning geom_line with geom_jitter
cols6 <- c("navyblue", "goldenrod3", "darkmagenta", "firebrick", "firebrick2", "darkolivegreen4", "violetred3")


# ASV level
asv_data <- env_data_rarefied %>% select(sample_name, SAMPLING_station, depth, original_name,sample_source, ASV, reads, sequencer, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  mutate(depth_bin = case_when(depth >= 0 & depth <= 20 ~ "0-20"), 
                        depth_bin = case_when(depth == 60 ~ "60", T ~ depth_bin), 
                        depth_bin = case_when(depth == 100 ~ "100", T ~ depth_bin), 
                        depth_bin = case_when(depth == 150 ~ "150", T ~ depth_bin), 
                        depth_bin = case_when(depth >= 200 & depth <= 700 ~ "200-700", T ~ depth_bin),
                        depth_bin = case_when(sample_source == "CN24F 150m exp" ~ "150 - benchtop experiment", T ~ depth_bin),
                        depth_bin = fct_relevel(depth_bin, "0-20", "60", "100", "150", "150 - benchtop experiment", "200-700"),
                        source_depth_bin = paste(sample_source, depth_bin, sep = " - "),
                        source_depth_bin = reorder(source_depth_bin, depth),
         source_depth_bin = paste(sample_source, depth_bin, sep = " - "),
         source_depth_bin = reorder(source_depth_bin, depth),
         SAMPLING_station = case_when(grepl("c1", SAMPLING_station) ~ "C1", T ~ SAMPLING_station),
         SAMPLING_station = case_when(grepl("MOORING1|Mooring1|MOORING 1", SAMPLING_station) ~ "M1", T ~ SAMPLING_station),
         SAMPLING_station = case_when(grepl("MOORING2|Mooring2|MOORING 2", SAMPLING_station) ~ "M2", T ~ SAMPLING_station),
         SAMPLING_station = case_when(grepl("CN24F", sample_name) ~ "Benchtop ESP", T ~ SAMPLING_station),
         station_depth = paste(SAMPLING_station, depth, sep = " - "), 
         station_depth = reorder(station_depth, depth),
         station_depth_bin = paste(SAMPLING_station, depth_bin, sep = " - "),
         station_depth_bin = reorder(station_depth_bin, depth)) %>%
  filter(reads > 0) %>%
  mutate(count = 1) %>%
  group_by(sample_name) %>%
  mutate(ASV_count = sum(count)) %>%
  ungroup() %>%
  distinct(sample_name, .keep_all = T) %>%
  mutate(sequencer = fct_relevel(sequencer, "MiSeq", "AVITI")) 


  asv_data %>% ggplot(aes(x=sequencer, y = ASV_count, group = original_name)) +
  geom_line(position = pd, aes(color = depth_bin), show.legend = F) +
  scale_color_manual(values = cols6) +
  new_scale_color() +
  geom_violin(aes(group = sequencer), alpha = 0, linewidth = 1) + 
  geom_point(position = pd, aes(fill = depth_bin), shape = 21, size = 3) +
  scale_fill_manual(values = cols6) +
  labs(
    subtitle = get_test_label(kruskal_test(ASV_count ~ sequencer, data = asv_data), detailed = TRUE),
    color = "Sequencer",
    y = "Number of ASVs",
    x = "Sequencer",
    fill = "Depth Bin (m)",
    title = "ASVs per sample - rarefied \nASVs with 1 read in a sample removed"
    )+
    theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        plot.title = element_text(size = 24, face = "bold"), 
        plot.subtitle = element_text(size = 20), 
        plot.caption = element_blank(), 
        legend.text =  element_text(size = 20, colour = "black"),
        legend.title =  element_text(face = "bold", size=23),
        legend.position = "none",
        axis.text = element_text(size = 20),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24, face = "bold")
        ) +
  guides(fill=guide_legend(override.aes = list(size=7)))
  

filename = paste(fig3_directory, 'Figure3d_scatter_plot_ASVs_per_sample_rarefied_ASVs_with_more_than_1_read_per_sample_',proj_prefix,'_',marker,'.png', sep='')
  print(marker)
  filename
  ggsave(filename,height = 10, width =8, units = 'in')
  
  filename = paste(fig3_directory, 'Figure3d_scatter_plot_ASVs_per_sample_rarefied_ASVs_with_more_than_1_read_per_sample_',proj_prefix,'_',marker,'.svg', sep='')
  print(marker)
  filename
  ggsave(filename,height = 10, width =8, units = 'in')
  

```


# #########
# diversity statistics on the rarefied data
## load in the data
```{r}

rotu_tab_env <- read_csv(paste(rarefaction_data,marker,'_env_Rarefied_asv_table_Rarefied_data.csv', sep=""))%>% column_to_rownames(var = "...1")
  
rtax_tab_env <- read_csv(paste(rarefaction_data,marker,'_env_Rarefied_taxa_table_Rarefied_data.csv', sep="")) %>% column_to_rownames(var = "...1")
  
rsamp_dat_env <- read_csv(paste(rarefaction_data,marker,'_env_Rarefied_meta_table_Rarefied_data.csv', sep="")) %>% column_to_rownames(var = "sample_name")
  
# combine into phyloseq object
 oBiom_r_env <- merge_phyloseq(otu_table(as.matrix(rotu_tab_env),taxa_are_rows = TRUE),
                          tax_table(as.matrix(rtax_tab_env)),
                          sample_data(rsamp_dat_env))
 
oBiom_r_env

```



## Chao1
##### sequencer - Chao1 diversity inex with kruskal-wallis and post hoc dunn test 
```{r}


sample_data(oBiom_r_env)
  
  p_oBiom <- plot_richness(oBiom_r_env,x="sequencer", measures=c("Chao1"), nrow=1, color='sequencer') +
    geom_violin(aes(x=sequencer, y=value)) +
  # geom_boxplot() +
    geom_jitter() +
    theme(text = element_text(size=16), strip.text.x = element_text(size = 16), axis.text = element_text(size = rel(1), colour = "black")) +
    xlab("sequencer") +
    labs(subtitle = "Chao1 Diversity") +
    ggtitle(paste("Community alpha diversity: ",marker,sep=""))+theme_bw() +scale_color_brewer(palette = "Dark2")

  p <- p_oBiom + 
    scale_fill_brewer(palette = "Spectral") + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  print(p)
  
set.seed(1234)
p_oBiom$data %>% sample_n_by(sequencer, size = 1)

p_oBiom$data <- p_oBiom$data 

p_oBiom$data %>% 
  group_by(sequencer) %>%
  get_summary_stats(value, type = "common")

ggboxplot(p_oBiom$data, x = "sequencer", y = "value")

res.kruskal <- p_oBiom$data %>% kruskal_test(value ~ sequencer)
res.kruskal

p_oBiom$data %>% kruskal_effsize(value ~ sequencer)

# Pairwise comparisons
pwc <- p_oBiom$data %>% 
  dunn_test(value ~ sequencer, p.adjust.method = "BH") 
pwc


cols6 <- c("navyblue", "goldenrod3", "darkmagenta", "firebrick", "firebrick2", "darkolivegreen4", "violetred3")

# 
# # Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "sequencer") 


pd <- position_dodge(0.2) # dodge position for aligning geom_line with geom_jitter


  ggplot(p_oBiom$data, aes(x = fct_rev(sequencer), y = value, group = original_name)) +
    geom_point(position = pd, aes(fill = depth_bin), shape = 21, size = 3) +
    scale_fill_manual(values = cols6) +
    geom_line(position = pd, aes(color = depth_bin), show.legend = "NA") +
    scale_color_manual(values = cols6) +
    new_scale_color() +
    geom_violin(aes(group = sequencer), alpha = 0, linewidth = 1) + 
    scale_color_manual(values = cols) +
  stat_pvalue_manual(pwc, hide.ns = TRUE, size = 3.88, bracket.nudge.y = 0, label = "p = {p.adj}") +
  labs(
    subtitle = get_test_label(res.kruskal, detailed = TRUE),
    caption = get_pwc_label(pwc),
    color = "sequencer",
    y = "Chao1 Diversity Index",
    x = "sequencer"
    )+
    ggtitle(paste("Chao1 alpha diversity - rarefied"))+
    theme_bw()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        plot.title = element_text(size = 24, face = "bold"), 
        plot.subtitle = element_text(size = 20), 
        plot.caption = element_blank(), 
        legend.text =  element_text(size = 20, colour = "black"),
        legend.title =  element_text(face = "bold", size=23),
        legend.position = "none",
        axis.text = element_text(size = 20),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24, face = "bold")
        ) +
  guides(fill=guide_legend(override.aes = list(size=7)))
  
filename = paste(fig3_directory, 'Figure3e_Chao1_alpha_diversity_',proj_prefix,'_',marker,'.png', sep='')
print(filename)
ggsave(filename, width = 8, height = 10, units = "in")

filename = paste(fig3_directory, 'Figure3e_Chao1_alpha_diversity_',proj_prefix,'_',marker,'.svg', sep='')
print(filename)
ggsave(filename, width = 8, height = 10, units = "in")
  
```



## Shannon
##### sequencer - Shannon diversity inex with kruskal-wallis and post hoc dunn test 
```{r}

sample_data(oBiom_r_env)
# season "Shannon", , "Simpson", "Chao1"
  
  p_oBiom <- plot_richness(oBiom_r_env,x="sequencer", measures=c("Shannon"), nrow=1, color='sequencer') +
    geom_violin(aes(x=sequencer, y=value)) +
  # geom_boxplot() +
    geom_jitter() +
    theme(text = element_text(size=16), strip.text.x = element_text(size = 16), axis.text = element_text(size = rel(1), colour = "black")) +
    xlab("sequencer") +
    labs(subtitle = "Shannon Diversity") +
    ggtitle(paste("Community alpha diversity: ",marker,sep=""))+theme_bw() +scale_color_brewer(palette = "Dark2")

  p <- p_oBiom + 
    scale_fill_brewer(palette = "Spectral") + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
  print(p)
  
set.seed(1234)
p_oBiom$data %>% sample_n_by(sequencer, size = 1)

p_oBiom$data <- p_oBiom$data 

p_oBiom$data %>% 
  group_by(sequencer) %>%
  get_summary_stats(value, type = "common")

ggboxplot(p_oBiom$data, x = "sequencer", y = "value")

res.kruskal <- p_oBiom$data %>% kruskal_test(value ~ sequencer)
res.kruskal

p_oBiom$data %>% kruskal_effsize(value ~ sequencer)

# Pairwise comparisons
pwc <- p_oBiom$data %>% 
  dunn_test(value ~ sequencer, p.adjust.method = "BH") 
pwc

pwc2 <- p_oBiom$data %>% 
  wilcox_test(value ~ sequencer, p.adjust.method = "BH")
pwc2

cols6 <- c("navyblue", "goldenrod3", "darkmagenta", "firebrick", "firebrick2", "darkolivegreen4", "violetred3")

# 
# # Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "sequencer") 



pd <- position_dodge(0.2) # dodge position for aligning geom_line with geom_jitter


  ggplot(p_oBiom$data, aes(x = fct_rev(sequencer), y = value, group = original_name)) +
    geom_point(position = pd, aes(fill = depth_bin), shape = 21, size = 3) +
    scale_fill_manual(values = cols6) +
    geom_line(position = pd, aes(color = depth_bin), show.legend = "NA") +
    scale_color_manual(values = cols6) +
    new_scale_color() +
    geom_violin(aes(group = sequencer), alpha = 0, linewidth = 1) + 
    scale_color_manual(values = cols) +
  # stat_pvalue_manual(pwc, hide.ns = TRUE, size = 3.88, bracket.nudge.y = 0, label = "p = {p.adj}") +
  labs(
    subtitle = get_test_label(res.kruskal, detailed = TRUE),
    caption = get_pwc_label(pwc),
    color = "sequencer",
    y = "Shannon Diversity Index",
    x = "sequencer"
    )+
    ggtitle(paste("Shannon alpha diversity - rarefied"))+
    theme_bw()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        plot.title = element_text(size = 24, face = "bold"), 
        plot.subtitle = element_text(size = 20), 
        plot.caption = element_blank(), 
        legend.text =  element_text(size = 20, colour = "black"),
        legend.title =  element_text(face = "bold", size=23),
        legend.position = "none",
        axis.text = element_text(size = 20),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24, face = "bold")
        ) +
  guides(fill=guide_legend(override.aes = list(size=7)))
  
filename = paste(fig3_directory, 'Figure3f_Shannon_alpha_diversity_',proj_prefix,'_',marker,'.png', sep='')
print(filename)
ggsave(filename, width = 8, height = 10, units = "in")

filename = paste(fig3_directory, 'Figure3f_Shannon_alpha_diversity_',proj_prefix,'_',marker,'.svg', sep='')
print(filename)
ggsave(filename, width = 8, height = 10, units = "in")



```


## Shannon
##### sequencer - Shannon diversity inex with kruskal-wallis and post hoc dunn test 
```{r}

sample_data(oBiom_r_env)
# season "Shannon", , "Simpson", "Chao1"
  
  p_oBiom <- plot_richness(oBiom_r_env,x="sequencer", measures=c("Simpson"), nrow=1, color='sequencer') +
    geom_violin(aes(x=sequencer, y=value)) +
  # geom_boxplot() +
    geom_jitter() +
    theme(text = element_text(size=16), strip.text.x = element_text(size = 16), axis.text = element_text(size = rel(1), colour = "black")) +
    xlab("sequencer") +
    labs(subtitle = "Simpson Diversity") +
    ggtitle(paste("Community alpha diversity: ",marker,sep=""))+theme_bw() +scale_color_brewer(palette = "Dark2")

  p <- p_oBiom + 
    scale_fill_brewer(palette = "Spectral") + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
  print(p)
  
set.seed(1234)
p_oBiom$data %>% sample_n_by(sequencer, size = 1)

p_oBiom$data <- p_oBiom$data 

p_oBiom$data %>% 
  group_by(sequencer) %>%
  get_summary_stats(value, type = "common")

ggboxplot(p_oBiom$data, x = "sequencer", y = "value")

res.kruskal <- p_oBiom$data %>% kruskal_test(value ~ sequencer)
res.kruskal

p_oBiom$data %>% kruskal_effsize(value ~ sequencer)

# Pairwise comparisons
pwc <- p_oBiom$data %>% 
  dunn_test(value ~ sequencer, p.adjust.method = "BH") 
pwc

pwc2 <- p_oBiom$data %>% 
  wilcox_test(value ~ sequencer, p.adjust.method = "BH")
pwc2

cols6 <- c("navyblue", "goldenrod3", "darkmagenta", "firebrick", "firebrick2", "darkolivegreen4", "violetred3")

# 
# # Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "sequencer") 



pd <- position_dodge(0.2) # dodge position for aligning geom_line with geom_jitter


  ggplot(p_oBiom$data, aes(x = fct_rev(sequencer), y = value, group = original_name)) +
    geom_point(position = pd, aes(fill = depth_bin), shape = 21, size = 3) +
    scale_fill_manual(values = cols6) +
    geom_line(position = pd, aes(color = depth_bin), show.legend = "NA") +
    scale_color_manual(values = cols6) +
    new_scale_color() +
    geom_violin(aes(group = sequencer), alpha = 0, linewidth = 1) + 
    scale_color_manual(values = cols) +
  # stat_pvalue_manual(pwc, hide.ns = TRUE, size = 3.88, bracket.nudge.y = 0, label = "p = {p.adj}") +
  labs(
    subtitle = get_test_label(res.kruskal, detailed = TRUE),
    caption = get_pwc_label(pwc),
    color = "sequencer",
    y = "Simpson Diversity Index",
    x = "sequencer"
    )+
    ggtitle(paste("Simpson alpha diversity - rarefied"))+
    theme_bw()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        plot.title = element_text(size = 24, face = "bold"), 
        plot.subtitle = element_text(size = 20), 
        plot.caption = element_blank(), 
        legend.text =  element_text(size = 20, colour = "black"),
        legend.title =  element_text(face = "bold", size=23),
        legend.position = "none",
        axis.text = element_text(size = 20),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 24, face = "bold")
        ) +
  guides(fill=guide_legend(override.aes = list(size=7)))
  
filename = paste(fig3_directory, 'Figure3g_Simpson_alpha_diversity_',proj_prefix,'_',marker,'.png', sep='')
print(filename)
ggsave(filename, width = 8, height = 10, units = "in")

filename = paste(fig3_directory, 'Figure3g_Simpson_alpha_diversity_',proj_prefix,'_',marker,'.svg', sep='')
print(filename)
ggsave(filename, width = 8, height = 10, units = "in")



```






