---
title: "Supplementary Figure 4 Subsampling raw AVITI .fastq files"
author: "Jacoby Baker"
date: "2026-03-26"
output: html_document
---


# Set up directories and variables
```{r}
# setwd('scripts/')
proj_prefix <- "AVITI_MiSeq_comparison_subsetted"
proj_name <- "AVITI VS MiSeq Comparison Subsetted"

data_directory <- paste0('../data/AVITI_subsetted/')
deicode_directory <- paste0('../data/AVITI_subsetted/DEICODE/18S/')
sfig4_directory <- paste0("../figures/SFig4/")

getwd()

marker = "18S"
```



# Load libraries
```{r}

#load libraries
library(readr) #read csv files
# library(vegan3d)

library(dplyr)
library(ggplot2)
library(ggthemes)
library(magrittr)
library(RColorBrewer) #colors for plotting
library("viridis") # for color fill 


library(tidyverse)
library(phyloseq)

library(taxa)
# library(metacoder)
library(ggpubr) # for ggarrange
library(reshape2) # for presence/absence fig


library(ggnewscale) # multiple fill color scales


library(ggh4x) # for facet nested wrap
library(ComplexHeatmap) # for the upset plots

```


```{r}


#ASV table
otu.c <- read_csv(paste0(data_directory,"filtered_seq_data/", proj_prefix, "_", marker, "_asv_filtered.csv"))
otu.c

#taxa table
tax.c <- read_csv(paste0(data_directory,"filtered_seq_data/", proj_prefix, "_", marker, "_taxa_filtered.csv"))
tax.c

#metadata table
samp.c <- read_csv(paste0(data_directory,"filtered_seq_data/", proj_prefix, "_", marker, "_meta_filtered.csv")) %>%
  mutate(sample_source = fct_relevel(sample_source,  "CN23 MARS" , "MBTS 17024", "MBTS 20524", "MBTS 22624", "MBTS 25424", "CN24F 150m exp")) %>%
  mutate(short_name = "") %>%
  mutate(short_name = case_when(grepl("_FD", sample_name) ~ substr(sample_name,1,nchar(sample_name)-9), T ~ short_name)) %>%
  mutate(short_name = case_when(grepl("_FK", sample_name) ~ substr(sample_name,12,nchar(sample_name)-18), T ~ short_name)) %>%
  mutate(short_name = case_when(grepl("CN24F150", sample_name) ~ substr(sample_name,14,nchar(sample_name)-3), T ~ short_name)) %>%
  mutate(short_name = case_when(grepl("subsampled_CN24F150", sample_name) ~ substr(sample_name,25,nchar(sample_name)-13), T ~ short_name)) %>%
  mutate(original_name = case_when(grepl("subsampled_", original_name) ~ substr(original_name,12,nchar(original_name)-0), T ~ original_name)) #%>% select(sample_name, library, source_depth, depth_round, sample_source, seqID, original_name)
samp.c


data <- read_csv(paste0(data_directory,"18S_AVITI_subsetted_Dada2_read_track.csv"))  %>% mutate(MiSeq = MiSeq_perc_pass_filter, AVITI = AVITI_subsetted_perc_pass_filter) %>% select(MiSeq_plate, AVITI_subsetted_plate, Order, MiSeq, AVITI)  %>% filter(Order < 90) 

data

AVITI_subbed_and_non_data <- read_csv(paste0(data_directory,"AVITI_subsetted_and_non_subsetted_read_track.csv"))  %>% mutate(AVITI = AVITI_perc_pass_filter, AVITI_subsetted = AVITI_subsetted_perc_pass_filter) %>% select(AVITI_plate, AVITI_subsetted_plate, Order, AVITI_subsetted, AVITI)  %>% filter(Order < 90) 

# dat <- data %>%  pivot_wider( -sample_id, names_to ='sequencer',values_to = 'perc_pass_filter' ) #%>% pivot_wider(names_from = sequencer, values_from=perc_pass_filter)

# dat



data_long <- read_csv(paste0(data_directory,"18S_AVITI_subsetted_Dada2_read_track_long.csv")) %>% mutate(library = sample_id) %>%select(library, perc_pass_filter, Order, sequencer)  %>% filter(Order < 90) %>% left_join(samp.c, by = "library")

data_long

#dna concentrations
dna_dat <- read_csv(paste0("../data/DNA_extract_concentrations.csv"))
dna_dat


#OTU table long format with percent total reads
potu.c <- otu.c %>%
  tidyr::pivot_longer( -ASV, names_to ='sample_name',values_to = 'reads' ) %>%
  ungroup() %>%
  group_by(sample_name) %>%
  mutate(per_tot = reads / sum(reads) *100) %>% #proportions(reads)*100
  arrange(-reads)
head(potu.c)

```

# Supplemental Figure 4A - Dada2 Pass Filter reads

## paired t-test
```{r}

plot(MiSeq ~ AVITI, data = data)


t.test(x=data$MiSeq, y=data$AVITI, paired = T)


```


## DADA2 %PF reads for each sequencer
```{r}
cols6 <- c("navyblue", "goldenrod3", "darkmagenta", "firebrick", "firebrick2", "darkolivegreen4", "violetred3")

dat <- data_long %>%  mutate(depth_bin = case_when(depth_round >= 0 & depth_round <= 20 ~ "0-20"), 
                        depth_bin = case_when(depth_round == 60 ~ "60", T ~ depth_bin), 
                        depth_bin = case_when(depth_round == 100 ~ "100", T ~ depth_bin), 
                        depth_bin = case_when(depth_round == 150 ~ "150", T ~ depth_bin), 
                        depth_bin = case_when(depth_round >= 200 & depth_round <= 700 ~ "200-700", T ~ depth_bin),
                        depth_bin = case_when(sample_source == "CN24F 150m exp" ~ "150 - benchtop experiment", T ~ depth_bin),
                        depth_bin = fct_relevel(depth_bin, "0-20", "60", "100", "150", "150 - benchtop experiment", "200-700"),
                        source_depth_bin = paste(sample_source, depth_bin, sep = " - "),
                        source_depth_bin = reorder(source_depth_bin, depth_round)) %>%
  select(sequencer, sample_source, perc_pass_filter, depth_bin, original_name) %>% 
  left_join(dna_dat) %>%
  mutate(DNA_quant_ng_ul = round(DNA_quant_ng_ul, digits = 1)) %>%
  pivot_wider(names_from = sequencer, values_from = perc_pass_filter)

dat

ggplot(dat, aes(y=MiSeq, x=AVITI, color = sample_source)) +
  geom_point()


ggplot(dat, aes(y=MiSeq, x=AVITI, color = depth_bin)) +
  geom_point() + 
  scale_color_manual(values = cols6) +
  geom_abline(intercept = 0, slope = 1, color = "grey20", linetype = "dashed") +
  geom_text(x=58, y=59, label= "1:1 line", angle = 42, check_overlap = T, color = "grey20")

ggplot(dat, aes(y=MiSeq, x=AVITI, color = DNA_quant_ng_ul)) +
  geom_point() + 
  scale_color_gradient(high = "deepskyblue3", low =  "violetred3") +
  geom_abline(intercept = 0, slope = 1, color = "grey20", linetype = "dashed") +
  geom_text(x=58, y=59, label= "1:1 line", angle = 42, check_overlap = T, color = "grey20")

ggplot(dat, aes(x=MiSeq, y=AVITI, fill = depth_bin)) +
  geom_point(aes(size = DNA_quant_ng_ul), shape = 21) + 
  scale_fill_manual(values = cols6) +
  geom_abline(intercept = 0, slope = 1, color = "grey20", linetype = "dashed") +
  geom_text(x=58, y=59, label= "1:1 line", angle = 42, check_overlap = T, color = "grey20") +
  labs(
    title = "Percent DADA2 Pass Filter reads for each sample",
    fill = "Depth bin (m)",
    size = "DNA concentration (ng/ul)",
    y = "AVITI subsampled .fastq files"
  ) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),title = element_text(size = 16),
        axis.title = element_text(size = 22),
        axis.text = element_text(size = 20),
        legend.text = element_text(size = 18),
        legend.title = element_text(size = 20)
        )+ guides(fill = guide_legend(override.aes = list(size=5, shape=21)),
            shape = guide_legend(override.aes = list(size = 5, shape = c(19, 1))))

filename = paste(sfig4_directory, 'SFig4a_AVITI_subsampled_raw_fastq_DADA2_pass_filter_reads_',proj_prefix, '_', marker, '.png', sep='')
  print(marker)
  filename
  ggsave(filename,height = 8, width =10, units = 'in')


  filename = paste(sfig4_directory, 'SFig4a_AVITI_subsampled_raw_fastq_DADA2_pass_filter_reads_',proj_prefix, '_', marker, '.svg', sep='')
  print(marker)
  filename
  ggsave(filename,height = 8, width =10, units = 'in')


```


# Supplemental Figure 4B - RPCA
## PHYLOSEQ  ALL SAMPLES
define function
```{r}
vegan_otu <- function(physeq) {
  OTU <- otu_table(physeq)
  if (taxa_are_rows(OTU)) {
    OTU <- t(OTU)
  }
  return(as(OTU, "matrix"))
}
```


## ## all reads

### create Phyloseq object of all  samples
```{r}

  
  subset = c("all_reads")
  Sys.setenv(subset = subset)
  
  metadata_updated <- samp.c %>% mutate(depth_bin = case_when(depth >= 0 & depth <= 20 ~ "0-20"), 
                        depth_bin = case_when(depth == 60 ~ "60", T ~ depth_bin), 
                        depth_bin = case_when(depth == 100 ~ "100", T ~ depth_bin), 
                        depth_bin = case_when(depth == 150 ~ "150", T ~ depth_bin), 
                        depth_bin = case_when(depth >= 200 & depth <= 700 ~ "200-700", T ~ depth_bin),
                        depth_bin = case_when(sample_source == "CN24F 150m exp" ~ "150 - benchtop experiment", T ~ depth_bin),
                        depth_bin = fct_relevel(depth_bin, "0-20", "60", "100", "150", "150 - benchtop experiment", "200-700"))
  

  
species_label <- tax.c %>% column_to_rownames(var = "ASV")
  
  rownames(metadata_updated) <- NULL
  metadata_updated <- column_to_rownames(metadata_updated,"sample_name")
  
  
  phylo <- merge_phyloseq(otu_table(as.matrix(read.csv(paste0(data_directory,"filtered_seq_data/", proj_prefix, "_", marker, "_asv_filtered.csv"), row.names = 1,check.names = FALSE)),taxa_are_rows = TRUE),
                          tax_table(as.matrix(species_label)),
                          sample_data(metadata_updated))
  
  phylo

    sample_data(phylo)
  
  phylo_obj <- phylo

  phylo_obj_env <- prune_samples(sample_data(phylo)$sample_type %in% c("environmental"), phylo) 
  
  
  
  
  
  sample_data(phylo_obj)
  sample_data(phylo_obj_env)
  # sample_data(phylo_obj_cb)
  
  # Create Tables -----------------------------------------------------------
  
  #create vegan tables
  vegan_matrix_env=vegan_otu(phylo_obj_env)
  otu_tab_env=otu_table(phylo_obj_env)
  tax_tab_env= tax_table(phylo_obj_env)
  samp_dat_env <- sample_data(phylo_obj_env)
  
  # Rarefy ------------------------------------------------------------------
  #print(sample_sums(phylo_obj_env))
  print(min(sample_sums(phylo_obj_env)))
  print(max(sample_sums(phylo_obj_env)))
  print(mean(sample_sums(phylo_obj_env)))
  print(phylo_obj_env)
  # Remove samples with <500 reads (like DEICODE)
  oBiom_r_env = prune_samples(sample_sums(phylo_obj_env)>=5000, phylo_obj_env)
  #set random seed for reproducibility later
  print(min(sample_sums(oBiom_r_env)))
  oBiom_r_env = rarefy_even_depth(oBiom_r_env, sample.size = min(sample_sums(oBiom_r_env)), rngseed = 678, replace = FALSE, trimOTUs = TRUE, verbose = TRUE)
  print(oBiom_r_env)
  assign(paste0("oBiom_r_env_",marker),oBiom_r_env)
  
  #create matrixes
  rvegan_matrix_env=vegan_otu(oBiom_r_env)
  rotu_tab_env=otu_table(oBiom_r_env)
  rtax_tab_env= tax_table(oBiom_r_env)
  rsamp_dat_env <- sample_data(oBiom_r_env)
  
  #save rarefied data
  #save to csv file
  file = paste(rarefaction_data,marker,'_env_Non-Rarefied_asv_table_',subset,'.csv', sep="")
  write.csv(otu_tab_env, file)
  file = paste(rarefaction_data,marker,'_env_Non-Rarefied_taxa_table_',subset,'.csv', sep="")
  write.csv(tax_tab_env, file)
  # file = paste(rarefaction_data,marker,'_env_Non-Rarefied_meta_table_',subset,'.csv', sep="")
  # write.csv(samp_dat_env, file)
  
  file = paste(rarefaction_data,marker,'_env_Rarefied_asv_table_',subset,'.csv', sep="")
  write.csv(rotu_tab_env, file)
  file = paste(rarefaction_data,marker,'_env_Rarefied_taxa_table_',subset,'.csv', sep="")
  write.csv(rtax_tab_env, file)
  # file = paste(rarefaction_data,marker,'_env_Rarefied_meta_table_',subset,'.csv', sep="")
  # write.csv(rsamp_dat_env, file)
  
  
  ### DEICODE PCA 
  # environmental samples
  phylo_obj_env_asv_table <- as.data.frame(otu_table(phylo_obj_env))
  phylo_obj_env_asv_table <- rownames_to_column(phylo_obj_env_asv_table,"#OTUID")
  write.table(phylo_obj_env_asv_table,paste0(deicode_directory,"/phylo_obj_env_",marker,"_asv_table_for_biom_",subset,".txt"),sep = "\t",row.names = FALSE,quote = FALSE)
  phylo_obj_env_tax_table <- as.data.frame(as(tax_table(phylo_obj_env),"matrix"))
  phylo_obj_env_tax_table <- rownames_to_column(phylo_obj_env_tax_table,"#OTUID")
  write.table(phylo_obj_env_tax_table,paste0(deicode_directory,"/phylo_obj_env_",marker,"_tax_table_for_biom_",subset,".txt"),sep = "\t",row.names = FALSE,quote = FALSE)
  phylo_obj_env_metadata <- as.data.frame(as.matrix(sample_data(phylo_obj_env)))
  phylo_obj_env_metadata <- rownames_to_column(phylo_obj_env_metadata,"#SampleID")
  write.table(phylo_obj_env_metadata,paste0(deicode_directory,"/phylo_obj_env_",marker,"_metadata_for_biom_",subset,".txt"),sep = "\t",row.names = FALSE,quote = FALSE)
  

  
```


### DEICODE
#### deicode PCA use a loop to jump into each of the folders for the three markers for all samples
```{bash engine.opts='-l'}
# move into the directory where you are going to run the DEICODE script
cd ../data/AVITI_subsetted/DEICODE
pwd


# Activate the conda qiime2 enviroment - check version
conda activate qiime2-2022.11

for folder in 18S  #for folder in 12S 18S COI
do
  #for data in phylo_obj phylo_obj_env phylo_obj_cb
  for data in phylo_obj_env

  do
  cd $folder
  #!/usr/bin/env bash

  # This bash script can be run with bash deicode_template_072820.sh
  # All of the input files (asv_table_for_deicode.txt, tax_table_for_deicode.txt, and metadata_for_deicode.txt) must be in the   same directory as this script.

  ### Description ###
  #
  # You will need qiime2 and DEICODE installed on your machine in order for it to work.
  # You will have four outputs: 1) distance.qza, 2) ordination.qza, 3) biplot.qza, and 4_ qurro_plot_deicoderank.qzv
  # Of these outputs, the most import one is the "ordination.qza" file - this is what you will use in R
  # biplot.qza can be visualized at view.qiime2.org, and is an interactive way to view your PCA.
  # qurro_plot_deicoderank.qzv will show you the loadings of your taxa on the different PCs.

  #Make biom file
    biom convert -i ./${data}_${folder}_asv_table_for_biom_${subset}.txt -o table.from_txt_json_${data}_${subset}.biom --table-type="OTU table" --to-json
  #add metadata files to biom file - change the merged_tax_table_for_biomm.txt and _merged_for_biom.txt files
   biom add-metadata -i table.from_txt_json_${data}_${subset}.biom -o table.w_md_${data}_${subset}.biom --observation-metadata-fp ${data}_${folder}_tax_table_for_biom_${subset}.txt --sample-metadata-fp ${data}_${folder}_metadata_for_biom_${subset}.txt

  #import into Qiime2 - Change output file
    qiime tools import \
     --input-path table.w_md_${data}_${subset}.biom \
     --output-path ${data}_${folder}_master_${subset}.biom.qza \
     --type FeatureTable[Frequency]

  #run DEICODE - change input table
    qiime deicode rpca \
      --i-table ${data}_${folder}_master_${subset}.biom.qza \
      --p-n-components 3 \
      --p-min-feature-count 10 \
      --p-min-sample-count 5000 \
      --o-biplot ordination_${data}_${subset}.qza \
      --o-distance-matrix distance_${data}_${subset}.qza

  ## Create biplot - change metadata files
    qiime emperor biplot \
      --i-biplot ordination_${data}_${subset}.qza \
      --m-sample-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --m-feature-metadata-file ${data}_${folder}_tax_table_for_biom_${subset}.txt \
      --o-visualization biplot_${data}_${subset}.qzv \
      --p-number-of-features 8

  # Create Qurro Visualization - change metadata files
    qiime qurro loading-plot \
      --i-ranks ordination_${data}_${subset}.qza \
      --i-table ${data}_${folder}_master_${subset}.biom.qza \
      --m-sample-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --m-feature-metadata-file ${data}_${folder}_tax_table_for_biom_${subset}.txt \
      --o-visualization qurro_plot_deicoderank_${data}_${subset}.qzv

   # Run a permanova | permdisp | or anosim test between whatever metadata column you choose (argument --m-metadata-column [column of choosing])
     qiime diversity beta-group-significance \
      --i-distance-matrix distance_${data}_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --m-metadata-column sequencer \
      --p-method anosim \
      --o-visualization sequencer_significance_anosim_${data}_${subset}.qzv

    # Alpha diversity test
    #Shannon
     qiime diversity alpha \
       --i-table ${data}_${folder}_master_${subset}.biom.qza \
       --p-metric shannon \
       --o-alpha-diversity ${data}_${folder}_shannon_alpha_${subset}.qza

      qiime diversity alpha-group-significance \
      --i-alpha-diversity ${data}_${folder}_shannon_alpha_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --o-visualization ${data}_${folder}_shannon_alpha_${subset}.qzv
      
      #Chao1
      qiime diversity alpha \
       --i-table ${data}_${folder}_master_${subset}.biom.qza \
       --p-metric chao1 \
       --o-alpha-diversity ${data}_${folder}_chao1_alpha_${subset}.qza

      qiime diversity alpha-group-significance \
      --i-alpha-diversity ${data}_${folder}_chao1_alpha_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --o-visualization ${data}_${folder}_chao1_alpha_${subset}.qzv
      
      #Observed OTUs
      qiime diversity alpha \
       --i-table ${data}_${folder}_master_${subset}.biom.qza \
       --p-metric observed_features \
       --o-alpha-diversity ${data}_${folder}_observed_otus_alpha_${subset}.qza

      qiime diversity alpha-group-significance \
      --i-alpha-diversity ${data}_${folder}_observed_otus_alpha_${subset}.qza \
      --m-metadata-file ${data}_${folder}_metadata_for_biom_${subset}.txt \
      --o-visualization ${data}_${folder}_observed_otus_alpha_${subset}.qzv
   done

  cd ../
done

```




#### import Qiim2 results
```{r}

#Import Qiime2 Results
file = paste(deicode_directory,"/ordination_phylo_obj_env_",subset,".qza",sep="")
print(file)
pco<-read_qza(file)
pco$uuid
#look at data
head(pco$data$ProportionExplained)
pco$data$Vectors[1:5, 1:4]

#create proportion explained labels
label.PC1 <- paste("PC1: ", round(pco$data$ProportionExplained$PC1, 3)*100,"%")
label.PC1
label.PC2 <- paste("PC2: ", round(pco$data$ProportionExplained$PC2, 3)*100,"%")
label.PC2
label.PC3 <- paste("PC3: ", round(pco$data$ProportionExplained$PC3, 3)*100,"%")
label.PC3

#Join with sample data
pco$data$Vectors <- pco$data$Vectors %>% 
  rename(., "sample_name" = "SampleID" )

pcscores <- left_join(pco$data$Vectors, samp.c)


#format loading scores
loadings <- as.data.frame(pco$data$Species)
loadings$ASV <- loadings$FeatureID

#join on OTU, adding taxa info
loadings <- left_join(loadings, tax.c, by="ASV")

#export pcscores
file = paste(deicode_directory,"scores/pcscores_",marker,"_phylo_obj_env_Dada2_Qiime2_",subset,".csv",sep="")
print(file)
write.csv(pcscores, file)
file = paste(deicode_directory,"scores/loadings_",marker,"_phylo_obj_env_Dada2_Qiime2_",subset,".csv",sep="")
print(file)
write.csv(loadings, file)
pcscores[1:5, 1:9]  #long because of sample data
head(loadings)


library(ggalt)
# Plot RPCA ---------------------------------------------------------------


cols6 <- c("navyblue", "goldenrod3", "darkmagenta", "firebrick", "firebrick2", "darkolivegreen4", "violetred3")

pcscores %>% mutate(depth_bin = case_when(depth >= 0 & depth <= 20 ~ "0-20"), 
                        depth_bin = case_when(depth == 60 ~ "60", T ~ depth_bin), 
                        depth_bin = case_when(depth == 100 ~ "100", T ~ depth_bin), 
                        depth_bin = case_when(depth == 150 ~ "150", T ~ depth_bin), 
                        depth_bin = case_when(depth >= 200 & depth <= 700 ~ "200-700", T ~ depth_bin),
                        depth_bin = case_when(sample_source == "CN24F 150m exp" ~ "150 - benchtop experiment", T ~ depth_bin),
                        depth_bin = fct_relevel(depth_bin, "0-20", "60", "100", "150", "150 - benchtop experiment", "200-700"),
                        source_depth_bin = paste(sample_source, depth_bin, sep = " - "),
                        source_depth_bin = reorder(source_depth_bin, depth)) %>%  #merge( aggregate(cbind(mean.x=PC1, mean.y=PC2)~sequencer,pcscores,mean),by="sequencer") %>%
  ggplot(aes(PC1,PC2, fill=depth_bin))+
  geom_line(aes(group = original_name), linetype = 2, alpha = 0.7) +
  geom_point(size=4, alpha=0.8, aes(shape = sequencer, color = depth_bin)) + 
  # stat_ellipse(aes(group = depth_bin, color = depth_bin)) +
  # geom_point(aes(x=mean.x,y=mean.y),size=6)+
  # geom_segment(aes( x=mean.x, y=mean.y, xend=PC1, yend=PC2)) +
  # geom_segment(aes(x=mean.x, y=mean.y, xend=PC1, yend=PC2)) +
  scale_shape_manual(values = c(21, 1)) +
    scale_fill_manual(values = cols6) +
    scale_color_manual(values = cols6) +
  # scale_color_tableau(palette = "Tableau 20") +
  # scale_color_viridis(option = "G", direction = 1, discrete = T, begin=.1, end = .9) +
  labs(x=label.PC1 , 
       y=label.PC2, 
       colour = "Depth Bin",
       shape = "Sequencer",
       subtitle = "Comparison between sequencers\nLines connect sample replicates",
       title = "AVITI subsetted raw .fastq files RPCA")+ 
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.text = element_text(size = 15, colour = "black"),
        legend.title = element_text(face = "bold", size=12),
        legend.key = element_rect(fill = FALSE, colour = FALSE),
        legend.key.size = unit(0.1,"line"),
        plot.title = element_text(size = 22),
        plot.subtitle = element_text(size = 18),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15)
  )+ guides(fill = "none",
            shape = guide_legend(override.aes = list(size = 5, shape = c(19, 1))))


filename = paste(sfig4_directory, 'SFig4b_AVITI_subsampled_raw_fastq_RPCA_',proj_prefix, '_', marker, '.png', sep='')
print('Plot of PC1PC2 colored by Year:')
filename
ggsave(filename,height = 7, width =9, units = 'in')

filename = paste(sfig4_directory, 'SFig4b_AVITI_subsampled_raw_fastq_RPCA_',proj_prefix, '_', marker, '.svg', sep='')
print('Plot of PC1PC2 colored by Year:')
filename
ggsave(filename,height = 7, width =9, units = 'in')



# plot inverse of PC1 and PC2 values so the pattern matches more with the rarefied PCA (e.g., the 150m ESP samples in the same quadrant)

# pcscores %>% mutate(depth_bin = case_when(depth >= 0 & depth <= 20 ~ "0-20"), 
#                         depth_bin = case_when(depth == 60 ~ "60", T ~ depth_bin), 
#                         depth_bin = case_when(depth == 100 ~ "100", T ~ depth_bin), 
#                         depth_bin = case_when(depth == 150 ~ "150", T ~ depth_bin), 
#                         depth_bin = case_when(depth >= 200 & depth <= 700 ~ "200-700", T ~ depth_bin),
#                         depth_bin = case_when(sample_source == "CN24F 150m exp" ~ "150 - benchtop experiment", T ~ depth_bin),
#                         depth_bin = fct_relevel(depth_bin, "0-20", "60", "100", "150", "150 - benchtop experiment", "200-700"),
#                         source_depth_bin = paste(sample_source, depth_bin, sep = " - "),
#                         source_depth_bin = reorder(source_depth_bin, depth)) %>%  #merge( aggregate(cbind(mean.x=PC1, mean.y=PC2)~sequencer,pcscores,mean),by="sequencer") %>%
#   ggplot(aes(-PC1,-PC2, fill=depth_bin))+
#   geom_line(aes(group = original_name, color = depth_bin), linetype = 2, alpha = 0.7) +
#   geom_point(size=4, alpha=0.8, aes(shape = sequencer, color = depth_bin)) + 
#   # stat_ellipse(aes(group = depth_bin, color = depth_bin)) +
#   # geom_point(aes(x=mean.x,y=mean.y),size=6)+
#   # geom_segment(aes( x=mean.x, y=mean.y, xend=PC1, yend=PC2)) +
#   # geom_segment(aes(x=mean.x, y=mean.y, xend=PC1, yend=PC2)) +
#   scale_shape_manual(values = c(21, 1)) +
#     scale_fill_manual(values = cols6) +
#     scale_color_manual(values = cols6, guide="none") +
#   # scale_color_tableau(palette = "Tableau 20") +
#   # scale_color_viridis(option = "G", direction = 1, discrete = T, begin=.1, end = .9) +
#   labs(x=label.PC1 , 
#        y=label.PC2, 
#        fill = "Depth Bin (m)",
#        shape = "Sequencer",
#        subtitle = "Comparison between sequencers\nLines connect sample replicates" )+ 
#   ggtitle(paste(proj_name, marker, subset, "PCA showing beta diversity"))+ 
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#         panel.background = element_blank(), axis.line = element_line(colour = "black"),
#         legend.text = element_text(size = 12, colour = "black"),
#         legend.title = element_text(face = "bold", size=12),
#         legend.key = element_rect(fill = FALSE, colour = FALSE),
#         legend.key.size = unit(0.1,"line")
#   )+ guides(fill = guide_legend(override.aes = list(size=5, shape=c(21))),
#             shape = guide_legend(override.aes = list(size = 5, shape = c(19, 1))))
# 
# 
# filename = paste(pca_directory,subset,'/',marker,"_",subset,'_RPCA_PC1_PC2_by_depth_bin_inverse.png', sep='')
# print('Plot of PC1PC2 colored by Year:')
# filename
# ggsave(filename,height = 7, width =9, units = 'in')




```



# Supplemental Figure 4C - Total reads


## Compile data - 
```{r}

# environmental data
env_data <- inner_join(samp.c, potu.c, by = c("sample_name")) %>% 
  left_join(tax.c %>% 
  # assign  algae
  mutate(Phylum = case_when(Class == "Raphidophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Phaeophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Pelagophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Dictyochophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Chrysophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Synurophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Bolidophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Eustigmatophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  # assign oomycotes
  mutate(Phylum = case_when(Class == "Oomycota" ~ "Oomycota", T ~ Phylum))  %>%
  mutate(Phylum = case_when(Class == "Oomycetes" ~ "Oomycota", T ~ Phylum))%>%
  # assign microeuk
  mutate(Phylum = case_when(Class == "Bigyra" ~ "Bigyra", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Choanoflagellata" ~ "Choanoflagellata", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Cryptophyceae" ~ "Cryptista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Order == "Telonemida" ~ "Cryptista", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Haptophyta" ~ "Haptista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Centroplasthelida" ~ "Haptista", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Ichthyosporea" ~ "Choanozoa", T ~ Phylum))  %>% 
  
  mutate(Class = case_when(Phylum == "Imbricatea" ~ "Imbricatea", T ~ Class)) %>%
  mutate(Phylum = case_when(Class == "Imbricatea" ~ "Cercozoa", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Polycystinea" ~ "Retaria", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Acantharea" ~ "Retaria", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Breviatea" ~ "Sulcozoa", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Aphelidea" ~ "Aphelida", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Filasterea" ~ "Choanozoa", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Phylum %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Phylum)) %>% 
  mutate(Class = case_when(Class %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Class)) %>% 
  mutate(Order = case_when(Order %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Order)) %>% 
  mutate(Family = case_when(Family %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Family)) %>% 
  mutate(Genus = case_when(Genus %in% c("unknown", "unassigned", "no_hit", "g_") ~ "unknown", T ~ Genus)) %>% 
  mutate(Species = case_when(Species %in% c("unknown", "unassigned", "no_hit", "s_") ~ "unknown", T ~ Species)))

env_data 


```



## Total Reads
## by sequencer
#### all reads
```{r}

taxas <- c('Species') # , 'Kingdom', 'Class', 'Order', 'Family', 'Genus', 'Species'
  
  print(paste(marker,"results"))
  
 sample_taxa_comp <- env_data %>%
      # full_join(species_label) %>%
  filter(reads > 0) 
  
#get number of samples per year/month and sample_source
spp_counts <-  sample_taxa_comp %>% 
  filter(reads > 0) %>%
  select(sample_source, sequencer, Phylum, Class, Order, Family, Genus, Species) %>%
  # mutate(count = 1) %>%
  dplyr::ungroup() %>%
  distinct(sample_source, sequencer, Phylum, Class, Order, Family, Genus, Species, .keep_all = T) %>%
  mutate(count = 1) %>%
  # group_by(sample_name) %>%
  # mutate(count = 1) %>%
  # ungroup() %>%
  dplyr::group_by(sample_source, sequencer) %>%
  mutate(new_count = sum(count)) %>%
  dplyr::ungroup() %>%
  distinct(sample_source, sequencer, .keep_all = T)

  spp_counts
  
  
  
  
  # assign(paste0("samp.c_",marker),samp.c)
  for (val in taxas) {
    taxa_level = sym(val)
  
  
    top_20_spp <- sample_taxa_comp %>% 
      group_by(Phylum, Class, Order, Family, Genus, Species) %>%
      mutate(sum_per_tot = sum(per_tot)) %>%
      distinct(Phylum, Class, Order, Family, Genus, Species, .keep_all = TRUE ) %>%
      arrange(-sum_per_tot) %>%
      select( Phylum, Class, Order, Family, Genus, Species, sum_per_tot, sample_source) %>%
      # print(n = Inf) %>%
      ungroup() %>%
      select( Phylum, Class, Order, Family, Genus, Species, sum_per_tot)%>%
    top_n(20) %>%
    mutate(top_20_spp = "yes") 
  
    
  large_color_palette_cal <- c("black", "slategray3", "deepskyblue4", "darkgoldenrod4", "lightgoldenrod3",  "indianred1", "violetred3", "#55ad89", "darkseagreen2", "chocolate2", "orange1", 
                              "sienna4", "hotpink2", "darkolivegreen4", "honeydew2", "goldenrod3", "gold1", "darkmagenta",  "plum3", "slateblue3", "darkslategray3"
                               )
  
  select_sample <- sample_taxa_comp %>%
    left_join(spp_counts) %>%
    full_join(top_20_spp) %>%
    mutate(top_20_spp = case_when(top_20_spp == "yes" ~ "yes",
                                  TRUE ~ "no")) %>%
    mutate(Species = case_when(top_20_spp == "no" ~ paste0("(not top 20 Taxa)"),
                                     TRUE ~ !!taxa_level)) %>%
      ungroup() %>%
      select(sample_name, ASV, Phylum, Class, Order, Family, Genus, Species, !!taxa_level,reads, per_tot, sample_source, sequencer, SC, source_depth, new_count, top_20_spp) %>%
      group_by(sample_source, sequencer) %>%
      mutate(pcr_reads = sum(reads)) %>%
      # mutate(sum_per_tot = sum(per_tot)) %>%
      ungroup() %>%
    group_by(sample_source, sequencer, Phylum, Class, Order, Family, Genus, Species) %>%
      mutate(pcr_taxa_reads = sum(reads)) %>%
      mutate(pcr_per_tot_taxa_reads = (pcr_taxa_reads/pcr_reads)*100) %>%
      ungroup() %>%
      distinct(sample_source, sequencer, Phylum, Class, Order, Family, Genus, Species, .keep_all = TRUE) 
    
    select_sample
    
    
    textcol <- "grey40"
    print(textcol)
    
    sample_line_plot <-  select_sample %>% 
      mutate(ord_fam_gen = "") %>%
      mutate(ord_fam_gen = paste(Phylum, Class, Order, Family, Genus, Species, sep = ", ")) %>%
      mutate(ord_fam_gen = case_when(Species ==  "(not top 20 Taxa)" ~ "(not top 20 Taxa)", T ~ ord_fam_gen)) %>%
      ungroup() %>% 
      # mutate(sample_name_fct = fct_reorder(as_factor(sample_name), source_depth)) %>%
      ggplot(aes(x=sequencer, y= pcr_taxa_reads)) +
    geom_bar(stat = 'identity', aes(fill = ord_fam_gen), size = .5, color = "black") +
    geom_text(aes(label=new_count, y= -30000), size = 5) +
      scale_y_continuous(breaks = seq(from = 0, to = 1200000, by = 200000), labels = scales::label_comma()) +
    labs(title = paste("Total reads per sequencer"), 
         subtitle = "Raw AVITI .fastq files subsampled",
         x="# of Species\nSequencer",
         y="Total reads by sample set",
         fill = "Phylum, Class, Order, Family, Genus, Species") +
      scale_fill_manual(values = large_color_palette_cal) +
      facet_nested_wrap(sample_source  ~ ., scales = "free_x", nrow = 1) +
    theme(
      #legend
      legend.position="bottom",legend.direction="vertical",
      legend.text=element_text(colour=textcol,size=12,face="bold"),
      legend.title=element_text(colour=textcol,size=15,face="bold"),
      axis.text=element_text(size=13,colour=textcol),
      axis.title = element_text(size=15),
      plot.title = element_text(size = 25),
      plot.subtitle = element_text(size = 22),
      # axis.ticks.x = element_blank(),
      plot.background=element_blank(),
      panel.border=element_blank(),
      panel.background = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_blank(),
      # plot.margin=margin(0.1,0.1,0.1,0.1,"cm"),
      strip.text = element_text(size=12, face = "bold", angle = 0),
      # strip.background = element_rect(fill = "white", color = "black"),
      # strip.background = element_rect(),
      # strip.text.y.left = element_blank(),
      # strip.text.y.right = element_blank()
      ) +
      guides(fill=guide_legend(ncol=1, title.position="top"))
  print(sample_line_plot) 
  
  filename = paste(sfig4_directory, 'SFig4c_AVITI_subsampled_raw_fastq_total_reads_by_sample_set_',proj_prefix, '_', marker, '.png', sep='')
print('Plot of PC1PC2 colored by Year:')
filename
ggsave(filename,height = 13, width =10, units = 'in')


filename = paste(sfig4_directory, 'SFig4c_AVITI_subsampled_raw_fastq_total_reads_by_sample_set_',proj_prefix, '_', marker, '.svg', sep='')
print('Plot of PC1PC2 colored by Year:')
filename
ggsave(filename,height = 13, width =10, units = 'in')
   
   
  }

  
  
  
```



# Supplemental Figure 4D - Upset plot

## setup dataset variable
```{r}
dataset <- "all_samples"
dataset1 <- "All samples"
# grep_search <- "Kelp"
comparisons <- c("MiSeq", "AVITI")

```

## Add bar chart with % reads of taxa in the entire dataset

```{r}


taxa_data <- env_data %>%
  filter(reads > 0) %>%
  select(sample_name, -ASV, reads, per_tot, sequencer, Phylum, Class, Order, Family, Genus, Species) %>%
  group_by(sample_name, Phylum, Class, Order, Family, Genus, Species) %>%
  mutate(reads = sum(reads), per_tot = sum(per_tot)) %>%
  ungroup() %>%
  distinct(sample_name, Phylum, Class, Order, Family, Genus, Species, .keep_all = T) %>%
  group_by(sample_name) %>%
  mutate(sample_tot_reads = sum(reads)) %>%
  ungroup() %>%
  group_by(sequencer) %>%
  mutate(seq_tot_reads = sum(reads))  %>%
  ungroup() %>%
  group_by(sequencer, Phylum, Class, Order, Family, Genus, Species) %>%
  mutate(seq_taxa_reads = sum(reads), seq_taxa_per_tot = seq_taxa_reads/seq_tot_reads*100) %>%
  ungroup() %>%
  distinct(sequencer, Phylum, Class, Order, Family, Genus, Species, .keep_all = T) %>%
  select(sequencer, Phylum, Class, Order, Family, Genus, Species, seq_taxa_per_tot) %>% 
  mutate(present = case_when(seq_taxa_per_tot > 0 ~ 1, T ~ 0)) %>%
  select(-seq_taxa_per_tot) %>%
  arrange(sequencer) %>%
  pivot_wider(names_from = "sequencer", values_from = "present", values_fill = 0) 


taxa_data

comb_data <- make_comb_mat(taxa_data)
comb_data
set_name(comb_data)
comb_name(comb_data)
set_size(comb_data)
comb_size(comb_data)
comb_degree(comb_data)
t(comb_data)



 long_dat_pres_taxa <- env_data %>%
  filter(reads > 0) %>%
  select(sample_name, -ASV, reads, per_tot, sequencer, Phylum, Class, Order, Family, Genus, Species) %>%
  group_by(sample_name, Phylum, Class, Order, Family, Genus, Species) %>%
  mutate(reads = sum(reads), per_tot = sum(per_tot)) %>%
  ungroup() %>%
  distinct(sample_name, Phylum, Class, Order, Family, Genus, Species, .keep_all = T) %>%
  group_by(sample_name) %>%
  mutate(sample_tot_reads = sum(reads)) %>%
  ungroup() %>%
  group_by(sequencer) %>%
  mutate(seq_tot_reads = sum(reads))  %>%
  ungroup() %>%
  group_by(sequencer, Phylum, Class, Order, Family, Genus, Species) %>%
  mutate(seq_taxa_reads = sum(reads), seq_taxa_per_tot = seq_taxa_reads/seq_tot_reads*100) %>%
  ungroup() %>%
  distinct(sequencer, Phylum, Class, Order, Family, Genus, Species, .keep_all = T) %>%
  select(sequencer, Phylum, Class, Order, Family, Genus, Species, seq_tot_reads, seq_taxa_reads, seq_taxa_per_tot)
  

 long_dat_pres_taxa
 
sequencers = colnames(taxa_data)[7:8]
sequencers

taxa_data[sequencers] = taxa_data[sequencers] == 1
t(head(taxa_data[sequencers], 100))


top_20_spp <- env_data %>%
  filter(reads > 0) %>%
  select(sample_name, -ASV, reads, per_tot, sequencer, Phylum, Class, Order, Family, Genus, Species) %>%
  group_by(sample_name, Phylum, Class, Order, Family, Genus, Species) %>%
  mutate(reads = sum(reads), per_tot = sum(per_tot)) %>%
  ungroup() %>%
  distinct(sample_name, Phylum, Class, Order, Family, Genus, Species, .keep_all = T) %>%
      group_by(Phylum) %>% 
      mutate(sum_per_tot = sum(per_tot)) %>%
      distinct(Phylum,.keep_all = TRUE ) %>% 
      arrange(-sum_per_tot) %>%
      select( Phylum, sum_per_tot) %>%
      ungroup()  %>%
    top_n(20) %>%
    mutate(top_20_spp = "yes") %>%
      select( Phylum, top_20_spp)
top_20_spp

     
 large_color_palette_cal <- c("black", "slategray3", "deepskyblue4", "darkgoldenrod4", "lightgoldenrod3",  "indianred1", "violetred3", "#55ad89", "darkseagreen2", "chocolate2", "orange1", 
                              "sienna4", "hotpink2", "darkolivegreen4", "honeydew2", "goldenrod3", "gold1", "darkmagenta",  "plum3", "slateblue3", "darkslategray3",
                              "red", "green", "blue", "yellow", "purple", "brown", "darkgreen", "skyblue1"
                               )
 

 updated3 <- long_dat_pres_taxa %>%
  mutate(full_tax = paste(Phylum)) %>%
    full_join(top_20_spp) %>%
    mutate(top_20_spp = case_when(top_20_spp == "yes" ~ "yes",
                                  TRUE ~ "no")) %>%
    mutate(full_tax = case_when(top_20_spp == "no" ~ paste0("(not top 20 Taxa)"),
                                     TRUE ~ full_tax)) 

test <- updated3 %>% full_join(taxa_data, by=c("Phylum", "Class", "Order", "Family", "Genus", "Species")) %>% 
  group_by_at(vars(13:14, 2:7)) %>%
  mutate(seq_taxa_reads_2 = sum(seq_taxa_reads), seq_tot_reads_2 = sum(seq_tot_reads), seq_taxa_per_tot_2 = seq_taxa_reads_2/seq_tot_reads_2*100) %>% 
  ungroup() %>%
  distinct_at(vars(13:14, 2:7), .keep_all = T) 

test %>% filter(AVITI == "FALSE")
  

 
 p <- upset(
    test,
    sequencers,
    mode = "distinct",
    annotations = list(
        'seq_taxa_per_tot_2'=(
            ggplot(mapping=aes(fill=reorder(full_tax, seq_taxa_per_tot_2)))
            + geom_bar(stat='identity', aes(y = seq_taxa_per_tot_2)) # , color = "grey20"
            # + geom_bar(stat='identity', position='fill', aes(y = pore_taxa_reads))
            # + scale_y_continuous(labels=scales::percent_format())
            # + scale_fill_manual(values=c(
            #     'R'='#E41A1C', 'PG'='#377EB8',
            #     'PG-13'='#4DAF4A', 'NC-17'='#FF7F00'
            # ))
            + scale_fill_manual(values = large_color_palette_cal, name = "Top 20 Phylum")
            + ylab('Percent taxa reads in the dataset') 
            + labs(title = paste("Upset plot - number of taxa and % reads of taxa in each combination", sep = " "))
            + theme(legend.position = "right", 
              legend.direction = "vertical",
              legend.text = element_text(size = 14),
              title = element_text(size = 22),
              axis.text.y = element_text(size = 16),
              axis.title = element_text(size = 20, face = "bold"))
      + guides(fill=guide_legend(ncol=1))
        )
    ),
    width_ratio=0.1,
    themes=upset_modify_themes(
        list(
            'intersections_matrix'=theme(text=element_text(size=20))))
  ,
    base_annotations=list(
    'Number of unique taxa'=intersection_size(
      counts=TRUE, 
      mapping=aes(fill=full_tax),
      text=list(check_overlap=TRUE, size = 8),
    ) + scale_fill_manual(values = large_color_palette_cal, guide='none') +
      theme(
        axis.text.y = element_text(size = 16),
        axis.title = element_text(size = 20, face = "bold")
      )
  ),
    set_sizes=(
        upset_set_size(
            geom=geom_bar(
                aes(fill=full_tax, x=group), # , color = "grey20"
                width=0.8
            ),
            position='right',
        ) + scale_fill_manual(values = large_color_palette_cal, guide='none')
        + labs(title = "Number of taxa")
        + geom_text(aes(label=..count..), hjust=.5, vjust = 1, stat='count', angle = 90)
        + expand_limits(y=1600)
        + theme(axis.ticks.x=element_line(),
            title = element_text(size = 12))
    ), 
guides = 'collect', 
sort_intersections_by='degree',
sort_sets=FALSE
)
 
 p
 
filename = paste(sfig4_directory, 'SFig4d_AVITI_subsampled_raw_fastq_upset_plot_',proj_prefix, '_', marker, '.png', sep='')
print('Plot of PC1PC2 colored by Year:')
filename
ggsave(filename,height = 12, width =14, units = 'in')


filename = paste(sfig4_directory, 'SFig4d_AVITI_subsampled_raw_fastq_upset_plot_',proj_prefix, '_', marker, '.svg', sep='')
print('Plot of PC1PC2 colored by Year:')
filename
ggsave(filename,height = 12, width =14, units = 'in')


```










