---
title: "ASV taxa comparisons"
author: "Katherine Silliman"
date: "2026-03-27"
output: html_document
---




Adapted from J. Baker's ASV_seq_comparisions.Rmd.

# Set up directories and variables
```{r}
# setwd('scripts/')
proj_prefix <- "AVITI_MiSeq_comparison"
proj_name <- "AVITI VS MiSeq Comparison"

subset_data_directory <- paste0('../data/AVITI_subsetted/filtered_seq_data/')
subset_rare_data <- paste0('../data/AVITI_subsetted/rarefaction_data/')
data_directory <- paste0('../data/filtered_seq_data/')
rarefaction_data <- paste0('../data/rarefaction_data/')

getwd()
enrichment_directory <- paste0("../data/enrichment_analysis/")
sfig5_directory <- paste0("../figures/SFig5/")

marker = "18S"
```

# Load libraries
```{r}
library(tidyverse)
library(ggplot2)
```

# UnRarefied data
```{r}
#ASV table
otu.c <- read_csv(paste0(data_directory, proj_prefix, "_", marker, "_asv_filtered.csv"))
head(otu.c)

#taxa table
tax.c <- read_csv(paste0(data_directory, proj_prefix, "_", marker, "_taxa_filtered.csv"))
head(tax.c)

# add corrected phyla
#class2phyla <- read_csv(paste0(data_directory,"phyla_unknown_aviti_class2phyla.csv"))

# Perform the update using left_join and mutate
#tax.c_updated <- tax.c %>%
#  left_join(class2phyla, by = c("Class" = "orig_class")) %>%
#  mutate(Phylum = ifelse(!is.na(corr_phylum), corr_phylum, Phylum)) %>%
#  select(-corr_phylum) # Remove the temporary join column

#metadata table
samp.c <- read_csv(paste0(data_directory, proj_prefix, "_", marker, "_meta_filtered.csv")) %>%
  mutate(sample_source = fct_relevel(sample_source,  "CN23 MARS" , "MBTS 17024", "MBTS 20524", "MBTS 22624", "MBTS 25424", "CN24F 150m exp")) %>%
  mutate(short_name = "") %>%
  mutate(short_name = case_when(grepl("eDNA", sample_name) ~ substr(sample_name,1,nchar(sample_name)-9), T ~ short_name)) %>%
  mutate(short_name = case_when(grepl("CN24F150", sample_name) ~ substr(sample_name,14,nchar(sample_name)-3), T ~ short_name))
  
head(samp.c)

seq.c <- read_csv(paste0(data_directory, proj_prefix, "_", marker, "_seq_filtered.csv"))
head(seq.c)

#OTU table long format with percent total reads
potu.c <- otu.c %>%
  tidyr::pivot_longer( -ASV, names_to ='sample_name',values_to = 'reads' ) %>%
  ungroup() %>%
  group_by(sample_name) %>%
  mutate(per_tot = reads / sum(reads) *100) %>% #proportions(reads)*100
  arrange(-reads)
head(potu.c)
```


## set lowest taxonomic annotation level for ASVs
```{r}

species_label <- tax.c  %>%
  # assign  algae
  mutate(Phylum = case_when(Class == "Raphidophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Phaeophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Pelagophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Dictyochophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Chrysophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Synurophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Bolidophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Eustigmatophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  # assign oomycotes
  mutate(Phylum = case_when(Class == "Oomycota" ~ "Oomycota", T ~ Phylum))  %>%
  mutate(Phylum = case_when(Class == "Oomycetes" ~ "Oomycota", T ~ Phylum))%>%
  # assign microeuk
  mutate(Phylum = case_when(Class == "Bigyra" ~ "Bigyra", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Choanoflagellata" ~ "Choanoflagellata", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Cryptophyceae" ~ "Cryptista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Order == "Telonemida" ~ "Cryptista", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Haptophyta" ~ "Haptista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Centroplasthelida" ~ "Haptista", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Ichthyosporea" ~ "Choanozoa", T ~ Phylum))  %>% 
  
  mutate(Class = case_when(Phylum == "Imbricatea" ~ "Imbricatea", T ~ Class)) %>%
  mutate(Phylum = case_when(Class == "Imbricatea" ~ "Cercozoa", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Polycystinea" ~ "Retaria", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Acantharea" ~ "Retaria", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Breviatea" ~ "Sulcozoa", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Aphelidea" ~ "Aphelida", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Filasterea" ~ "Choanozoa", T ~ Phylum)) %>%
  
  mutate(Kingdom = case_when(Kingdom %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Kingdom)) %>%
  mutate(Phylum = case_when(Phylum %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Phylum)) %>% 
  mutate(Class = case_when(Class %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Class)) %>% 
  mutate(Order = case_when(Order %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Order)) %>% 
  mutate(Family = case_when(Family %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Family)) %>% 
  mutate(Genus = case_when(Genus %in% c("unknown", "unassigned", "no_hit", "g_") ~ "unknown", T ~ Genus)) %>% 
  mutate(Species = case_when(Species %in% c("unknown", "unassigned", "no_hit", "s_") ~ "unknown", T ~ Species))

```




```{r}
# Compile data - 
# environmental data
env_data <- inner_join(samp.c, potu.c, by = c("sample_name")) %>% 
  left_join(species_label)

head(env_data)


```


## SHARED/UNIQUE Taxa

```{r}
length(unique(env_data$Species))
dim(env_data %>% select(sample_name, ASV, reads, sequencer, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  # filter(grepl(paste(grep_search), sample_name)) %>%
  filter(reads > 0))
```

## Unique taxa strings in Phyla
```{r}
taxa_data_unique <- env_data %>% select(sample_name, ASV, reads, sequencer, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  filter(reads > 0) %>%
  group_by(sequencer, Phylum, Class, Order, Family, Genus,Species) %>%
  mutate(sum_reads_unique= sum(reads)) %>%
  ungroup() %>%
  distinct(sequencer, Phylum, Class, Order, Family, Genus, Species, .keep_all = T) %>%
  mutate(present = case_when(sum_reads_unique> 0 ~ 1, T ~ 0)) %>% 
  select(-ASV,-sample_name, -reads, -sum_reads_unique) %>%
  pivot_wider(names_from = "sequencer", values_from = "present", values_fill = 0)

shared_uniq <- taxa_data_unique %>%
  filter(MiSeq == 1 & AVITI == 1)


MiSeq_uniq <- taxa_data_unique %>%
  filter(MiSeq == 1 & AVITI == 0)


AVITI_uniq <- taxa_data_unique %>%
  filter(MiSeq == 0 & AVITI == 1)

  
  
#comb_data <- make_comb_mat(data)

# Display the number of unique species found by each sequencer
cat("Number of unique taxa detected by MiSeq:", nrow((MiSeq_uniq)), "\n")
cat("Number of unique taxa detected by AVITI:", nrow(AVITI_uniq), "\n")
cat("Number of shared taxa detected:", nrow(shared_uniq), "\n")

```

```{r}
# Create a combined dataset of taxas unique to each sequencer
# Add a column 'DetectionSource' to indicate which sequencer detected it uniquely
all_unique_detections_uniq <- bind_rows(
  MiSeq_uniq %>% mutate(DetectionSource = "MiSeq_Unique"),
  AVITI_uniq%>% mutate(DetectionSource = "AVITI_Unique")
)

# only keep phyla with at least 5 ASVs in at least 1 sequencer
phylum_contingency_table_uniq <- all_unique_detections_uniq %>%
  count(Phylum, DetectionSource) %>%
  pivot_wider(names_from = DetectionSource, values_from = n, values_fill = 0) 

cat("\nContingency Table (Phylum vs. Unique Detection Source):\n")
print(phylum_contingency_table_uniq)

```

### Phylum-specific Enrichment Analysis (Unique vs. Common Taxa) 
This section compares the phylum distribution of taxas unique to each sequencer against the phylum distribution of taxas detected by *both* sequencers.

```{r}
phyla_combined_uniq <- unique(c(all_unique_detections_uniq$Phylum, shared_uniq$Phylum))
enrichment_results_unique_vs_common_uniq <- data.frame(
  Phylum = character(),
  Comparison = character(), # e.g., "SequencerA_Unique_vs_Both", "SequencerB_Unique_vs_Both"
  OddsRatio = numeric(),
  P_Value = numeric(),
  Adjusted_P_Value = numeric(),
  stringsAsFactors = FALSE
)

```

```{r}
for (phy in phyla_combined_uniq) {
  # --- Comparison: SequencerA_Unique vs. ASV_Both ---
  count_phy_A_unique <- sum(MiSeq_uniq$Phylum == phy)
  count_not_phy_A_unique <- sum(MiSeq_uniq$Phylum != phy)
  
  count_phy_both <- sum(shared_uniq$Phylum == phy)
  count_not_phy_both <- sum(shared_uniq$Phylum != phy)
  
  contingency_matrix_A_vs_Both <- matrix(c(count_phy_A_unique, count_not_phy_A_unique,
                                           count_phy_both, count_not_phy_both),
                                         nrow = 2, byrow = FALSE,
                                         dimnames = list(c(phy, paste0("Not_", phy)),
                                                         c("MiSeq_Unique", "Taxa_Both")))
  
  if (sum(contingency_matrix_A_vs_Both) > 0) {
    fisher_test_A_vs_Both <- fisher.test(contingency_matrix_A_vs_Both, alternative = "two.sided")
    odds_ratio_A_vs_Both <- fisher_test_A_vs_Both$estimate
    p_value_A_vs_Both <- fisher_test_A_vs_Both$p.value
    
    enrichment_results_unique_vs_common_uniq <- bind_rows(enrichment_results_unique_vs_common_uniq, data.frame(
      Phylum = phy,
      Comparison = "MiSeq_Unique_vs_Both",
      OddsRatio = odds_ratio_A_vs_Both,
      P_Value = p_value_A_vs_Both
    ))
  } else {
    cat(paste0("Skipping Fisher's test for Phylum '", phy, "' (A_Unique vs Both) due to no data.\n"))
  }

# --- Comparison: SequencerB_Unique vs. Species_Both ---
  count_phy_B_unique <- sum(AVITI_uniq$Phylum == phy)
  count_not_phy_B_unique <- sum(AVITI_uniq$Phylum != phy)
  
  contingency_matrix_B_vs_Both <- matrix(c(count_phy_B_unique, count_not_phy_B_unique,
                                           count_phy_both, count_not_phy_both),
                                         nrow = 2, byrow = FALSE,
                                         dimnames = list(c(phy, paste0("Not_", phy)),
                                                         c("AVITI_Unique", "Taxa_Both")))
  
  if (sum(contingency_matrix_B_vs_Both) > 0) {
    fisher_test_B_vs_Both <- fisher.test(contingency_matrix_B_vs_Both, alternative = "two.sided")
    odds_ratio_B_vs_Both <- fisher_test_B_vs_Both$estimate
    p_value_B_vs_Both <- fisher_test_B_vs_Both$p.value
    
    enrichment_results_unique_vs_common_uniq <- bind_rows(enrichment_results_unique_vs_common_uniq, data.frame(
      Phylum = phy,
      Comparison = "AVITI_Unique_vs_Both",
      OddsRatio = odds_ratio_B_vs_Both,
      P_Value = p_value_B_vs_Both
    ))
  } else {
    cat(paste0("Skipping Fisher's test for Phylum '", phy, "' (B_Unique vs Both) due to no data.\n"))
  }
}

# Adjust p-values for multiple comparisons across all unique vs. common tests
if (nrow(enrichment_results_unique_vs_common_uniq) > 0) {
  enrichment_results_unique_vs_common_uniq$Adjusted_P_Value <- p.adjust(enrichment_results_unique_vs_common_uniq$P_Value, method = "BH")
  
  cat("\nIndividual Phylum Enrichment Analysis Results (Unique vs. Common - Fisher's Exact Test):\n")
  print(enrichment_results_unique_vs_common_uniq %>% arrange(Adjusted_P_Value))
  
  cat("\nInterpretation of Individual Phylum Enrichment Analysis (Unique vs. Common):\n")
  cat("- 'OddsRatio' > 1 for 'SequencerA_Unique_vs_Both' suggests enrichment in SequencerA_Unique relative to Taxa detected by both.\n")
  cat("- 'OddsRatio' < 1 for 'SequencerA_Unique_vs_Both' suggests depletion in SequencerA_Unique relative to Taxa detected by both.\n")
  cat("- Similar interpretation applies for 'SequencerB_Unique_vs_Both'.\n")
  cat("- 'P_Value' indicates significance for that specific phylum and comparison.\n")
  cat("- 'Adjusted_P_Value' accounts for multiple comparisons.\n")
} else {
  cat("\nNo phyla found for individual enrichment analysis (Unique vs. Common).\n")
}
```
```{r}
filename <- paste(enrichment_directory, proj_prefix, '_', marker, '_enrichment_results_uniqueTaxa_in_Phyla.csv', sep='')
write.csv(enrichment_results_unique_vs_common_uniq %>% arrange(Adjusted_P_Value),filename,row.names = FALSE,quote = FALSE )
```

# ## S.FIG. 5A
```{r}
plot_data_categorized_uniq <- bind_rows(
  MiSeq_uniq %>% mutate(DetectionGroup = "MiSeq_Unique"),
  AVITI_uniq %>% mutate(DetectionGroup = "AVITI_Unique"),
  shared_uniq %>% mutate(DetectionGroup = "Detected_by_Both")
)

# Calculate percentages for plotting across all three groups
phylum_percentages_overall_uniq <- plot_data_categorized_uniq %>%
  group_by(DetectionGroup, Phylum) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(DetectionGroup) %>%
  mutate(Percentage = Count / sum(Count))

# Determine which phyla are significant based on enrichment_results_unique_vs_common
# A phylum is marked with '*' if either SequencerA_Unique_vs_Both or SequencerB_Unique_vs_Both is significant
significant_phyla_for_plot <- enrichment_results_unique_vs_common_uniq %>%
  filter(Adjusted_P_Value < 0.05) %>%
  distinct(Phylum) %>%
  mutate(Label = "*")

# Merge significance labels into the plot data for correct positioning
plot_data_with_significance <- phylum_percentages_overall_uniq %>%
  left_join(significant_phyla_for_plot, by = "Phylum") %>%
  filter(!is.na(Label))

# Find the maximum percentage for each phylum to position the asterisk
max_percentage_per_phylum <- plot_data_with_significance %>%
  group_by(Phylum) %>%
  summarise(Max_Percentage = max(Percentage), .groups = 'drop')

# Join max percentage back to the significance data for plotting
significant_phyla_labels <- significant_phyla_for_plot %>%
  left_join(max_percentage_per_phylum, by = "Phylum") %>%
  mutate(Y_Position = Max_Percentage + 0.02) # Add a small offset above the bar


# This bar plot helps visualize the phylum distribution for species
# unique to each sequencer and those detected by both, as percentages.
ggplot(plot_data_with_significance, aes(x = Phylum, y = Percentage, fill = DetectionGroup)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Distribution of Unique/Shared Taxa by Phylum and Sequencer, Unrarefied",
       x = "Phylum",
       y = "Percentage of Total Taxa in Group",
       fill = "Group") +
  scale_y_continuous(labels = scales::percent, limits = c(0, max(plot_data_with_significance$Percentage) * 1.1)) + # Adjust y-axis limit
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=15),
        axis.text.y = element_text(size=15),
        axis.title = element_text(size = 18),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 15),
        plot.title = element_text(size = 18)) +
  # Add asterisks for significant phyla
  geom_text(data = significant_phyla_labels, aes(x = Phylum, y = Y_Position, label = Label),
            inherit.aes = FALSE, size = 6, color = "black") +
  # Add a small legend for the asterisk
  annotate("text", x = Inf, y = Inf, label = "* Adjusted P-value < 0.05 (Unique vs. Common)",
           hjust = 1.1, vjust = 1.1, size = 5, color = "black", fontface = "italic")


filename = paste(sfig5_directory, 'SFig5a_enrichment_analysis_Taxa_nonrarefied_',proj_prefix, '_', marker, '.png', sep='')
filename
ggsave(filename,height = 7, width =10, units = 'in')

filename = paste(sfig5_directory, 'SFig5a_enrichment_analysis_Taxa_nonrarefied_',proj_prefix, '_', marker, '.svg', sep='')
filename
ggsave(filename,height = 7, width =10, units = 'in')

```
```{r}
phylum_percentages_overall_uniq %>%
  filter(Phylum %in% significant_phyla_for_plot$Phylum) %>%
  select(-Percentage) %>%
  pivot_wider(names_from = DetectionGroup, values_from = Count)

phylum_percentages_overall_uniq %>%
  filter(Phylum %in% significant_phyla_for_plot$Phylum) %>%
  select(-Count) %>%
  pivot_wider(names_from = DetectionGroup, values_from = Percentage)

```

## Species in Phyla. 
```{r}
taxa_data_spp <- env_data %>% select(sample_name, ASV, reads, sequencer, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  # filter(grepl(paste(grep_search), sample_name)) %>%
  filter(reads > 0) %>%
  group_by(sequencer, Species) %>%
  mutate(sum_reads_spp = sum(reads)) %>%
  ungroup() %>%
  distinct(sequencer, Species, .keep_all = T) %>%
  mutate(present = case_when(sum_reads_spp > 0 ~ 1, T ~ 0)) %>% 
  select(-ASV,-sample_name, -reads, -sum_reads_spp) %>%
  pivot_wider(names_from = "sequencer", values_from = "present", values_fill = 0) %>% arrange(Species)

shared_spp <- taxa_data_spp %>%
  filter(MiSeq == 1 & AVITI == 1)


MiSeq_spp <- taxa_data_spp %>%
  filter(MiSeq == 1 & AVITI == 0)


AVITI_spp <- taxa_data_spp %>%
  filter(MiSeq == 0 & AVITI == 1)

  
  
#comb_data <- make_comb_mat(data)

# Display the number of unique species found by each sequencer
cat("Number of unique species detected by MiSeq:", nrow((MiSeq_spp)), "\n")
cat("Number of unique species detected by AVITI:", nrow(AVITI_spp), "\n")
cat("Number of shared species detected:", nrow(shared_spp), "\n")

```

```{r}
# Create a combined dataset of species unique to each sequencer
# Add a column 'DetectionSource' to indicate which sequencer detected it uniquely
all_unique_detections <- bind_rows(
  MiSeq_spp %>% mutate(DetectionSource = "MiSeq_Unique"),
  AVITI_spp%>% mutate(DetectionSource = "AVITI_Unique")
)

phylum_contingency_table <- all_unique_detections %>%
  count(Phylum, DetectionSource) %>%
  pivot_wider(names_from = DetectionSource, values_from = n, values_fill = 0)

cat("\nContingency Table (Phylum vs. Unique Detection Source):\n")
print(phylum_contingency_table)

```
### Test unique species against each other
```{r}
# Convert the contingency table to a matrix for statistical testing
# Remove the Phylum column before converting to matrix
phylum_matrix <- as.matrix(phylum_contingency_table[, -1])
rownames(phylum_matrix) <- phylum_contingency_table$Phylum

# Perform a Chi-squared test of independence
# This test determines if there is a significant association between
# the sequencer (as the source of unique species) and the phylum distribution.
# A small p-value (typically < 0.05) suggests a significant association.
# Note: Chi-squared test assumptions include sufficient expected cell counts.
# If expected counts are low (<5), Fisher's exact test is more appropriate.
# We'll check for this.
# Check expected counts for Chi-squared test
chi_sq_test_result <- chisq.test(phylum_matrix, simulate.p.value = FALSE)
cat("\nChi-squared Test Results:\n")
print(chi_sq_test_result)

# Get expected values
expected_counts <- chi_sq_test_result$expected
cat("\nExpected Counts for Chi-squared Test:\n")
print(expected_counts)
```

```{r}
# Check if any expected counts are too low for Chi-squared test
if (any(expected_counts < 5)) {
  cat("\nWarning: Some expected counts are less than 5. Fisher's Exact Test might be more appropriate.\n")
  # Perform Fisher's Exact Test if Chi-squared assumptions are violated
  # Fisher's exact test is suitable for small sample sizes and sparse tables.
  fisher_test_result <- fisher.test(phylum_matrix,workspace=2e8)
  cat("\nFisher's Exact Test Results:\n")
  print(fisher_test_result)
} else {
  cat("\nExpected counts are sufficient for Chi-squared test.\n")
}
```
#### Enrichment test
```{r}
enrichment_results <- data.frame(
  Phylum = character(),
  Sequencer_Enriched = character(),
  OddsRatio = numeric(),
  P_Value = numeric(),
  Adjusted_P_Value = numeric(),
  stringsAsFactors = FALSE
)

# Get all unique phyla present in the unique detections
all_phyla <- unique(all_unique_detections$Phylum)

cat("\n--- Phylum-specific Enrichment Analysis ---\n")

for (phy in all_phyla) {
  # Create a 2x2 contingency table for the current phylum
  # Row 1: Species in 'phy' vs. Species NOT in 'phy'
  # Column 1: Unique to SequencerA vs. Unique to SequencerB

  # Count of unique species in current phylum for SequencerA
  count_phy_A <- sum(all_unique_detections$Phylum == phy & all_unique_detections$DetectionSource == "MiSeq_Unique")
  # Count of unique species NOT in current phylum for SequencerA
  count_not_phy_A <- sum(all_unique_detections$Phylum != phy & all_unique_detections$DetectionSource == "MiSeq_Unique")

  # Count of unique species in current phylum for SequencerB
  count_phy_B <- sum(all_unique_detections$Phylum == phy & all_unique_detections$DetectionSource == "AVITI_Unique")
  # Count of unique species NOT in current phylum for SequencerB
  count_not_phy_B <- sum(all_unique_detections$Phylum != phy & all_unique_detections$DetectionSource == "AVITI_Unique")

  # Construct the 2x2 matrix for Fisher's exact test
  # Matrix structure:
  #         SequencerA_Unique   SequencerB_Unique
  # Phylum_X      a                   b
  # Not_Phylum_X  c                   d
  # where a = count_phy_A, b = count_phy_B, c = count_not_phy_A, d = count_not_phy_B
  
  contingency_matrix <- matrix(c(count_phy_A, count_not_phy_A,
                                 count_phy_B, count_not_phy_B),
                               nrow = 2, byrow = FALSE,
                               dimnames = list(c(phy, paste0("Not_", phy)),
                                               c("MiSeq_Unique", "AVITI_Unique")))

  # Perform Fisher's Exact Test
  # alternative = "two.sided" tests for enrichment or depletion
  # alternative = "greater" tests for enrichment in the first column (SequencerA_Unique)
  # alternative = "less" tests for enrichment in the second column (SequencerB_Unique)
  
  # For a general enrichment analysis, we often look for enrichment in either direction.
  # However, to specify which sequencer it's enriched in, we can look at the odds ratio.
  # If OR > 1, enriched in SequencerA_Unique relative to SequencerB_Unique
  # If OR < 1, enriched in SequencerB_Unique relative to SequencerA_Unique
  
  if (sum(contingency_matrix) > 0) { # Ensure there's data to test
    fisher_test <- fisher.test(contingency_matrix, alternative = "two.sided")
    
    odds_ratio <- fisher_test$estimate
    p_value <- fisher_test$p.value
    
    sequencer_enriched <- "None"
    if (odds_ratio > 1 && p_value < 0.05) { # Arbitrary p-value threshold for initial indication
      sequencer_enriched <- "MiSeq_Unique"
    } else if (odds_ratio < 1 && p_value < 0.05) {
      sequencer_enriched <- "AVITI_Unique"
    }
    
    enrichment_results <- bind_rows(enrichment_results, data.frame(
      Phylum = phy,
      Sequencer_Enriched = sequencer_enriched,
      OddsRatio = odds_ratio,
      P_Value = p_value
    ))
  } else {
      cat(paste0("Skipping Fisher's test for Phylum '", phy, "' due to no data.\n"))
  }
}

# Adjust p-values for multiple comparisons (e.g., using Benjamini-Hochberg method)
if (nrow(enrichment_results) > 0) {
  enrichment_results$Adjusted_P_Value <- p.adjust(enrichment_results$P_Value, method = "BH")
  
  cat("\nIndividual Phylum Enrichment Analysis Results (Fisher's Exact Test):\n")
  print(enrichment_results %>% arrange(Adjusted_P_Value))
  
  cat("\nInterpretation of Individual Phylum Enrichment Analysis:\n")
  cat("- 'OddsRatio' > 1 suggests enrichment in SequencerA_Unique relative to SequencerB_Unique.\n")
  cat("- 'OddsRatio' < 1 suggests enrichment in SequencerB_Unique relative to SequencerA_Unique.\n")
  cat("- 'P_Value' indicates significance for that specific phylum.\n")
  cat("- 'Adjusted_P_Value' accounts for multiple comparisons (important for controlling False Discovery Rate).\n")
  cat("- A 'Sequencer_Enriched' value indicates a significant enrichment (p < 0.05) in that sequencer.\n")
} else {
  cat("\nNo phyla found for individual enrichment analysis.\n")
}
```

```{r}
# Calculate percentages for plotting
phylum_percentages <- all_unique_detections %>%
  group_by(DetectionSource, Phylum) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(DetectionSource) %>%
  mutate(Percentage = Count / sum(Count))

# This bar plot helps visualize the phylum distribution for species
# uniquely detected by each sequencer, as percentages.
ggplot(phylum_percentages, aes(x = Phylum, y = Percentage, fill = DetectionSource)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Phylum Distribution of Species Unique to Each Sequencer (Percentages)",
       x = "Phylum",
       y = "Percentage of Unique Species",
       fill = "Sequencer") +
  scale_y_continuous(labels = scales::percent) + # Format y-axis as percentages
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Phylum-specific Enrichment Analysis (Unique vs. Common Species) 
This section compares the phylum distribution of species unique to each sequencer against the phylum distribution of species detected by *both* sequencers.

```{r}
all_phyla_combined <- unique(c(all_unique_detections$Phylum, shared_spp$Phylum))
enrichment_results_unique_vs_common <- data.frame(
  Phylum = character(),
  Comparison = character(), # e.g., "SequencerA_Unique_vs_Both", "SequencerB_Unique_vs_Both"
  OddsRatio = numeric(),
  P_Value = numeric(),
  Adjusted_P_Value = numeric(),
  stringsAsFactors = FALSE
)

```

```{r}
for (phy in all_phyla_combined) {
  # --- Comparison: SequencerA_Unique vs. Species_Both ---
  count_phy_A_unique <- sum(MiSeq_spp$Phylum == phy)
  count_not_phy_A_unique <- sum(MiSeq_spp$Phylum != phy)
  
  count_phy_both <- sum(shared_spp$Phylum == phy)
  count_not_phy_both <- sum(shared_spp$Phylum != phy)
  
  contingency_matrix_A_vs_Both <- matrix(c(count_phy_A_unique, count_not_phy_A_unique,
                                           count_phy_both, count_not_phy_both),
                                         nrow = 2, byrow = FALSE,
                                         dimnames = list(c(phy, paste0("Not_", phy)),
                                                         c("MiSeq_Unique", "Species_Both")))
  
  if (sum(contingency_matrix_A_vs_Both) > 0) {
    fisher_test_A_vs_Both <- fisher.test(contingency_matrix_A_vs_Both, alternative = "two.sided")
    odds_ratio_A_vs_Both <- fisher_test_A_vs_Both$estimate
    p_value_A_vs_Both <- fisher_test_A_vs_Both$p.value
    
    enrichment_results_unique_vs_common <- bind_rows(enrichment_results_unique_vs_common, data.frame(
      Phylum = phy,
      Comparison = "MiSeq_Unique_vs_Both",
      OddsRatio = odds_ratio_A_vs_Both,
      P_Value = p_value_A_vs_Both
    ))
  } else {
    cat(paste0("Skipping Fisher's test for Phylum '", phy, "' (A_Unique vs Both) due to no data.\n"))
  }

# --- Comparison: SequencerB_Unique vs. Species_Both ---
  count_phy_B_unique <- sum(AVITI_spp$Phylum == phy)
  count_not_phy_B_unique <- sum(AVITI_spp$Phylum != phy)
  
  contingency_matrix_B_vs_Both <- matrix(c(count_phy_B_unique, count_not_phy_B_unique,
                                           count_phy_both, count_not_phy_both),
                                         nrow = 2, byrow = FALSE,
                                         dimnames = list(c(phy, paste0("Not_", phy)),
                                                         c("AVITI_Unique", "Species_Both")))
  
  if (sum(contingency_matrix_B_vs_Both) > 0) {
    fisher_test_B_vs_Both <- fisher.test(contingency_matrix_B_vs_Both, alternative = "two.sided")
    odds_ratio_B_vs_Both <- fisher_test_B_vs_Both$estimate
    p_value_B_vs_Both <- fisher_test_B_vs_Both$p.value
    
    enrichment_results_unique_vs_common <- bind_rows(enrichment_results_unique_vs_common, data.frame(
      Phylum = phy,
      Comparison = "AVITI_Unique_vs_Both",
      OddsRatio = odds_ratio_B_vs_Both,
      P_Value = p_value_B_vs_Both
    ))
  } else {
    cat(paste0("Skipping Fisher's test for Phylum '", phy, "' (B_Unique vs Both) due to no data.\n"))
  }
}

# Adjust p-values for multiple comparisons across all unique vs. common tests
if (nrow(enrichment_results_unique_vs_common) > 0) {
  enrichment_results_unique_vs_common$Adjusted_P_Value <- p.adjust(enrichment_results_unique_vs_common$P_Value, method = "BH")
  
  cat("\nIndividual Phylum Enrichment Analysis Results (Unique vs. Common - Fisher's Exact Test):\n")
  print(enrichment_results_unique_vs_common %>% arrange(Adjusted_P_Value))
  
  cat("\nInterpretation of Individual Phylum Enrichment Analysis (Unique vs. Common):\n")
  cat("- 'OddsRatio' > 1 for 'SequencerA_Unique_vs_Both' suggests enrichment in SequencerA_Unique relative to species detected by both.\n")
  cat("- 'OddsRatio' < 1 for 'SequencerA_Unique_vs_Both' suggests depletion in SequencerA_Unique relative to species detected by both.\n")
  cat("- Similar interpretation applies for 'SequencerB_Unique_vs_Both'.\n")
  cat("- 'P_Value' indicates significance for that specific phylum and comparison.\n")
  cat("- 'Adjusted_P_Value' accounts for multiple comparisons.\n")
} else {
  cat("\nNo phyla found for individual enrichment analysis (Unique vs. Common).\n")
}
```
```{r}
filename <- paste(enrichment_directory, proj_prefix, '_', marker, '_enrichment_results_uniqueSpp_in_Phyla.csv', sep='')
write.csv(enrichment_results_unique_vs_common %>% arrange(Adjusted_P_Value),filename,row.names = FALSE,quote = FALSE )
```

# ## S.FIG. 5B
```{r}
plot_data_categorized <- bind_rows(
  MiSeq_spp %>% mutate(DetectionGroup = "MiSeq_Unique"),
  AVITI_spp %>% mutate(DetectionGroup = "AVITI_Unique"),
  shared_spp %>% mutate(DetectionGroup = "Detected_by_Both")
)

# Calculate percentages for plotting across all three groups
phylum_percentages_overall <- plot_data_categorized %>%
  group_by(DetectionGroup, Phylum) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(DetectionGroup) %>%
  mutate(Percentage = Count / sum(Count))

# Determine which phyla are significant based on enrichment_results_unique_vs_common
# A phylum is marked with '*' if either SequencerA_Unique_vs_Both or SequencerB_Unique_vs_Both is significant
significant_phyla_for_plot <- enrichment_results_unique_vs_common %>%
  filter(Adjusted_P_Value < 0.05) %>%
  distinct(Phylum) %>%
  mutate(Label = "*")

# Merge significance labels into the plot data for correct positioning
plot_data_with_significance <- phylum_percentages_overall %>%
  left_join(significant_phyla_for_plot, by = "Phylum") %>% 
  filter(!is.na(Label))

# Find the maximum percentage for each phylum to position the asterisk
max_percentage_per_phylum <- plot_data_with_significance %>%
  group_by(Phylum) %>%
  summarise(Max_Percentage = max(Percentage), .groups = 'drop')

# Join max percentage back to the significance data for plotting
significant_phyla_labels <- significant_phyla_for_plot %>%
  left_join(max_percentage_per_phylum, by = "Phylum") %>%
  mutate(Y_Position = Max_Percentage + 0.02) # Add a small offset above the bar


# This bar plot helps visualize the phylum distribution for species
# unique to each sequencer and those detected by both, as percentages.
ggplot(plot_data_with_significance, aes(x = Phylum, y = Percentage, fill = DetectionGroup)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Distribution of Unique/Shared Species by Phylum and Sequencer, Unrarefied",
       x = "Phylum",
       y = "Percentage of Total Species in Group",
       fill = "Group") +
  scale_y_continuous(labels = scales::percent, limits = c(0, max(plot_data_with_significance$Percentage) * 1.15)) + # Adjust y-axis limit
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=15),
        axis.text.y = element_text(size=15),
        axis.title = element_text(size = 18),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 15),
        plot.title = element_text(size = 18)) +
  # Add asterisks for significant phyla
  geom_text(data = significant_phyla_labels, aes(x = Phylum, y = Y_Position, label = Label),
            inherit.aes = FALSE, size = 6, color = "black") +
  # Add a small legend for the asterisk
  annotate("text", x = Inf, y = Inf, label = "* Adjusted P-value < 0.05 (Unique vs. Common)",
           hjust = 1.1, vjust = 1.1, size = 5, color = "black", fontface = "italic")


filename = paste(sfig5_directory, 'SFig5b_enrichment_analysis_Species_nonrarefied_',proj_prefix, '_', marker, '.png', sep='')
filename
ggsave(filename,height = 7, width =10, units = 'in')

filename = paste(sfig5_directory, 'SFig5b_enrichment_analysis_Species_nonrarefied_',proj_prefix, '_', marker, '.svg', sep='')
filename
ggsave(filename,height = 7, width =10, units = 'in')

```
```{r}
phylum_percentages_overall %>%
  filter(Phylum == "Foraminifera")
```

Plot raw spp numbers (not done)
```{r}
plot_data_categorized <- bind_rows(
  MiSeq_spp %>% mutate(DetectionGroup = "MiSeq_Unique"),
  AVITI_spp %>% mutate(DetectionGroup = "AVITI_Unique"),
  shared_spp %>% mutate(DetectionGroup = "Detected_by_Both")
)

# Calculate percentages for plotting across all three groups
phylum_percentages_overall <- plot_data_categorized %>%
  group_by(DetectionGroup, Phylum) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(DetectionGroup) %>%
  mutate(Percentage = Count / sum(Count))

# Determine which phyla are significant based on enrichment_results_unique_vs_common
# A phylum is marked with '*' if either SequencerA_Unique_vs_Both or SequencerB_Unique_vs_Both is significant
significant_phyla_for_plot <- enrichment_results_unique_vs_common %>%
  filter(Adjusted_P_Value < 0.05) %>%
  distinct(Phylum) %>%
  mutate(Label = "*")

# Merge significance labels into the plot data for correct positioning
plot_data_with_significance <- phylum_percentages_overall %>%
  left_join(significant_phyla_for_plot, by = "Phylum")

# Find the maximum percentage for each phylum to position the asterisk
max_percentage_per_phylum <- plot_data_with_significance %>%
  group_by(Phylum) %>%
  summarise(Max_Percentage = max(Percentage), .groups = 'drop')

# Join max percentage back to the significance data for plotting
significant_phyla_labels <- significant_phyla_for_plot %>%
  left_join(max_percentage_per_phylum, by = "Phylum") %>%
  mutate(Y_Position = Max_Percentage + 0.02) # Add a small offset above the bar


# This bar plot helps visualize the phylum distribution for species
# unique to each sequencer and those detected by both, as percentages.
ggplot(plot_data_with_significance, aes(x = Phylum, y = Percentage, fill = DetectionGroup)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Phylum Distribution by Detection Group (Percentages)",
       x = "Phylum",
       y = "Percentage of Species",
       fill = "Detection Group") +
  scale_y_continuous(labels = scales::percent, limits = c(0, max(plot_data_with_significance$Percentage) * 1.15)) + # Adjust y-axis limit
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  # Add asterisks for significant phyla
  geom_text(data = significant_phyla_labels, aes(x = Phylum, y = Y_Position, label = Label),
            inherit.aes = FALSE, size = 6, color = "black") +
  # Add a small legend for the asterisk
  annotate("text", x = Inf, y = Inf, label = "* Adjusted P-value < 0.05 (Unique vs. Common)",
           hjust = 1.1, vjust = 1.1, size = 3, color = "black", fontface = "italic")
```

## ASVs in Phyla
```{r}
taxa_data_asv <- env_data %>% select(sample_name, ASV, reads, sequencer, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  # filter(grepl(paste(grep_search), sample_name)) %>%
  filter(reads > 0) %>%
  group_by(sequencer, ASV) %>%
  mutate(sum_reads_asv = sum(reads)) %>%
  ungroup() %>%
  distinct(sequencer, ASV, .keep_all = T) %>%
  mutate(present = case_when(sum_reads_asv > 0 ~ 1, T ~ 0)) %>% 
  select(-sample_name, -reads, -sum_reads_asv) %>%
  pivot_wider(names_from = "sequencer", values_from = "present", values_fill = 0) %>% arrange(ASV)

shared_asv <- taxa_data_asv %>%
  filter(MiSeq == 1 & AVITI == 1)


MiSeq_asv <- taxa_data_asv %>%
  filter(MiSeq == 1 & AVITI == 0)


AVITI_asv <- taxa_data_asv %>%
  filter(MiSeq == 0 & AVITI == 1)

  
  
#comb_data <- make_comb_mat(data)

# Display the number of unique species found by each sequencer
cat("Number of unique ASVs detected by MiSeq:", nrow((MiSeq_asv)), "\n")
cat("Number of unique ASVs detected by AVITI:", nrow(AVITI_asv), "\n")
cat("Number of shared ASVs detected:", nrow(shared_asv), "\n")

```

```{r}
# Create a combined dataset of ASVs unique to each sequencer
# Add a column 'DetectionSource' to indicate which sequencer detected it uniquely
all_unique_detections_asv <- bind_rows(
  MiSeq_asv %>% mutate(DetectionSource = "MiSeq_Unique"),
  AVITI_asv%>% mutate(DetectionSource = "AVITI_Unique")
)

# only keep phyla with at least 5 ASVs in at least 1 sequencer
phylum_contingency_table_asv <- all_unique_detections_asv %>%
  count(Phylum, DetectionSource) %>%
  pivot_wider(names_from = DetectionSource, values_from = n, values_fill = 0) %>%
  filter(if_any(c(MiSeq_Unique, AVITI_Unique), ~ . > 4))

cat("\nContingency Table (Phylum vs. Unique Detection Source):\n")
print(phylum_contingency_table_asv)

```

### Phylum-specific Enrichment Analysis (Unique vs. Common ASVs) 
This section compares the phylum distribution of ASVs unique to each sequencer against the phylum distribution of ASVs detected by *both* sequencers.

```{r}
phyla_combined_asv <- unique(c(all_unique_detections_asv$Phylum, shared_asv$Phylum))
filt_phyla_combined_asv <- unique(phylum_contingency_table_asv$Phylum)
enrichment_results_unique_vs_common_asv <- data.frame(
  Phylum = character(),
  Comparison = character(), # e.g., "SequencerA_Unique_vs_Both", "SequencerB_Unique_vs_Both"
  OddsRatio = numeric(),
  P_Value = numeric(),
  Adjusted_P_Value = numeric(),
  stringsAsFactors = FALSE
)

```

```{r}
for (phy in filt_phyla_combined_asv) {
  # --- Comparison: SequencerA_Unique vs. ASV_Both ---
  count_phy_A_unique <- sum(MiSeq_asv$Phylum == phy)
  count_not_phy_A_unique <- sum(MiSeq_asv$Phylum != phy)
  
  count_phy_both <- sum(shared_asv$Phylum == phy)
  count_not_phy_both <- sum(shared_asv$Phylum != phy)
  
  contingency_matrix_A_vs_Both <- matrix(c(count_phy_A_unique, count_not_phy_A_unique,
                                           count_phy_both, count_not_phy_both),
                                         nrow = 2, byrow = FALSE,
                                         dimnames = list(c(phy, paste0("Not_", phy)),
                                                         c("MiSeq_Unique", "ASV_Both")))
  
  if (sum(contingency_matrix_A_vs_Both) > 0) {
    fisher_test_A_vs_Both <- fisher.test(contingency_matrix_A_vs_Both, alternative = "two.sided")
    odds_ratio_A_vs_Both <- fisher_test_A_vs_Both$estimate
    p_value_A_vs_Both <- fisher_test_A_vs_Both$p.value
    
    enrichment_results_unique_vs_common_asv <- bind_rows(enrichment_results_unique_vs_common_asv, data.frame(
      Phylum = phy,
      Comparison = "MiSeq_Unique_vs_Both",
      OddsRatio = odds_ratio_A_vs_Both,
      P_Value = p_value_A_vs_Both
    ))
  } else {
    cat(paste0("Skipping Fisher's test for Phylum '", phy, "' (A_Unique vs Both) due to no data.\n"))
  }

# --- Comparison: SequencerB_Unique vs. Species_Both ---
  count_phy_B_unique <- sum(AVITI_asv$Phylum == phy)
  count_not_phy_B_unique <- sum(AVITI_asv$Phylum != phy)
  
  contingency_matrix_B_vs_Both <- matrix(c(count_phy_B_unique, count_not_phy_B_unique,
                                           count_phy_both, count_not_phy_both),
                                         nrow = 2, byrow = FALSE,
                                         dimnames = list(c(phy, paste0("Not_", phy)),
                                                         c("AVITI_Unique", "ASV_Both")))
  
  if (sum(contingency_matrix_B_vs_Both) > 0) {
    fisher_test_B_vs_Both <- fisher.test(contingency_matrix_B_vs_Both, alternative = "two.sided")
    odds_ratio_B_vs_Both <- fisher_test_B_vs_Both$estimate
    p_value_B_vs_Both <- fisher_test_B_vs_Both$p.value
    
    enrichment_results_unique_vs_common_asv <- bind_rows(enrichment_results_unique_vs_common_asv, data.frame(
      Phylum = phy,
      Comparison = "AVITI_Unique_vs_Both",
      OddsRatio = odds_ratio_B_vs_Both,
      P_Value = p_value_B_vs_Both
    ))
  } else {
    cat(paste0("Skipping Fisher's test for Phylum '", phy, "' (B_Unique vs Both) due to no data.\n"))
  }
}

# Adjust p-values for multiple comparisons across all unique vs. common tests
if (nrow(enrichment_results_unique_vs_common_asv) > 0) {
  enrichment_results_unique_vs_common_asv$Adjusted_P_Value <- p.adjust(enrichment_results_unique_vs_common_asv$P_Value, method = "BH")
  
  cat("\nIndividual Phylum Enrichment Analysis Results (Unique vs. Common - Fisher's Exact Test):\n")
  print(enrichment_results_unique_vs_common_asv %>% arrange(Adjusted_P_Value))
  
  cat("\nInterpretation of Individual Phylum Enrichment Analysis (Unique vs. Common):\n")
  cat("- 'OddsRatio' > 1 for 'SequencerA_Unique_vs_Both' suggests enrichment in SequencerA_Unique relative to ASVs detected by both.\n")
  cat("- 'OddsRatio' < 1 for 'SequencerA_Unique_vs_Both' suggests depletion in SequencerA_Unique relative to ASVs detected by both.\n")
  cat("- Similar interpretation applies for 'SequencerB_Unique_vs_Both'.\n")
  cat("- 'P_Value' indicates significance for that specific phylum and comparison.\n")
  cat("- 'Adjusted_P_Value' accounts for multiple comparisons.\n")
} else {
  cat("\nNo phyla found for individual enrichment analysis (Unique vs. Common).\n")
}
```
```{r}
filename <- paste(enrichment_directory, proj_prefix, '_', marker, '_enrichment_results_uniqueASVs_in_Phyla.csv', sep='')
write.csv(enrichment_results_unique_vs_common_asv %>% arrange(Adjusted_P_Value),filename,row.names = FALSE,quote = FALSE )
```


# ## S.FIG. 5C
```{r}
plot_data_categorized_asv <- bind_rows(
  MiSeq_asv %>% mutate(DetectionGroup = "MiSeq_Unique"),
  AVITI_asv %>% mutate(DetectionGroup = "AVITI_Unique"),
  shared_asv %>% mutate(DetectionGroup = "Detected_by_Both")
) %>% filter(Phylum %in% filt_phyla_combined_asv)

# Calculate percentages for plotting across all three groups
phylum_percentages_overall_asv <- plot_data_categorized_asv %>%
  group_by(DetectionGroup, Phylum) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(DetectionGroup) %>%
  mutate(Percentage = Count / sum(Count))

# Determine which phyla are significant based on enrichment_results_unique_vs_common
# A phylum is marked with '*' if either SequencerA_Unique_vs_Both or SequencerB_Unique_vs_Both is significant
significant_phyla_for_plot <- enrichment_results_unique_vs_common_asv %>%
  filter(Adjusted_P_Value < 0.05) %>%
  distinct(Phylum) %>%
  mutate(Label = "*")

# Merge significance labels into the plot data for correct positioning
plot_data_with_significance <- phylum_percentages_overall_asv %>%
  left_join(significant_phyla_for_plot, by = "Phylum") %>%
  filter(!is.na(Label))

# Find the maximum percentage for each phylum to position the asterisk
max_percentage_per_phylum <- plot_data_with_significance %>%
  group_by(Phylum) %>%
  summarise(Max_Percentage = max(Percentage), .groups = 'drop')

# Join max percentage back to the significance data for plotting
significant_phyla_labels <- significant_phyla_for_plot %>%
  left_join(max_percentage_per_phylum, by = "Phylum") %>%
  mutate(Y_Position = Max_Percentage + 0.02) # Add a small offset above the bar


# This bar plot helps visualize the phylum distribution for species
# unique to each sequencer and those detected by both, as percentages.
ggplot(plot_data_with_significance, aes(x = Phylum, y = Percentage, fill = DetectionGroup)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Distribution of unique/shated ASVs by Phyl and Sequencer, Unrarefied",
       x = "Phylum",
       y = "Percentage of total ASVs in Group",
       fill = "Group") +
  scale_y_continuous(labels = scales::percent, limits = c(0, max(plot_data_with_significance$Percentage) * 1.1)) + # Adjust y-axis limit
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=15),
        axis.text.y = element_text(size=15),
        axis.title = element_text(size = 18),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 15),
        plot.title = element_text(size = 18)) +
  # Add asterisks for significant phyla
  geom_text(data = significant_phyla_labels, aes(x = Phylum, y = Y_Position, label = Label),
            inherit.aes = FALSE, size = 6, color = "black") +
  # Add a small legend for the asterisk
  annotate("text", x = Inf, y = Inf, label = "* Adjusted P-value < 0.05 (Unique vs. Common)",
           hjust = 1.1, vjust = 1.1, size = 5, color = "black", fontface = "italic")


filename = paste(sfig5_directory, 'SFig5c_enrichment_analysis_ASVs_nonrarefied_',proj_prefix, '_', marker, '.png', sep='')
filename
ggsave(filename,height = 7, width =10, units = 'in')

filename = paste(sfig5_directory, 'SFig5c_enrichment_analysis_ASVs_nonrarefied_',proj_prefix, '_', marker, '.svg', sep='')
filename
ggsave(filename,height = 7, width =10, units = 'in')


```

```{r}
phylum_percentages_overall_asv %>%
  filter(Phylum %in% significant_phyla_for_plot$Phylum) %>%
  select(-Percentage) %>%
  pivot_wider(names_from = DetectionGroup, values_from = Count)

phylum_percentages_overall_asv%>%
  filter(Phylum %in% significant_phyla_for_plot$Phylum) %>%
  select(-Count) %>%
  pivot_wider(names_from = DetectionGroup, values_from = Percentage)

```

## Unique taxa strings in Kingdom

```{r}
# Create a combined dataset of taxas unique to each sequencer

king_contingency_table_uniq <- all_unique_detections_uniq %>%
  count(Kingdom, DetectionSource) %>%
  pivot_wider(names_from = DetectionSource, values_from = n, values_fill = 0) 

cat("\nContingency Table (Kingdom vs. Unique Detection Source):\n")
print(king_contingency_table_uniq)

```

### Kingdom-specific Enrichment Analysis (Unique vs. Common Taxa) 
This section compares the kingdom distribution of taxas unique to each sequencer against the kingdom distribution of taxas detected by *both* sequencers.

```{r}
king_combined_uniq <- unique(c(all_unique_detections_uniq$Kingdom, shared_uniq$Kingdom))
enrichment_results_unique_vs_common_uniq <- data.frame(
  Kingdom = character(),
  Comparison = character(), # e.g., "SequencerA_Unique_vs_Both", "SequencerB_Unique_vs_Both"
  OddsRatio = numeric(),
  P_Value = numeric(),
  Adjusted_P_Value = numeric(),
  stringsAsFactors = FALSE
)

```

```{r}
for (phy in king_combined_uniq) {
  # --- Comparison: SequencerA_Unique vs. ASV_Both ---
  count_phy_A_unique <- sum(MiSeq_uniq$Kingdom == phy)
  count_not_phy_A_unique <- sum(MiSeq_uniq$Kingdom != phy)
  
  count_phy_both <- sum(shared_uniq$Kingdom == phy)
  count_not_phy_both <- sum(shared_uniq$Kingdom != phy)
  
  contingency_matrix_A_vs_Both <- matrix(c(count_phy_A_unique, count_not_phy_A_unique,
                                           count_phy_both, count_not_phy_both),
                                         nrow = 2, byrow = FALSE,
                                         dimnames = list(c(phy, paste0("Not_", phy)),
                                                         c("MiSeq_Unique", "Taxa_Both")))
  
  if (sum(contingency_matrix_A_vs_Both) > 0) {
    fisher_test_A_vs_Both <- fisher.test(contingency_matrix_A_vs_Both, alternative = "two.sided")
    odds_ratio_A_vs_Both <- fisher_test_A_vs_Both$estimate
    p_value_A_vs_Both <- fisher_test_A_vs_Both$p.value
    
    enrichment_results_unique_vs_common_uniq <- bind_rows(enrichment_results_unique_vs_common_uniq, data.frame(
      Kingdom = phy,
      Comparison = "MiSeq_Unique_vs_Both",
      OddsRatio = odds_ratio_A_vs_Both,
      P_Value = p_value_A_vs_Both
    ))
  } else {
    cat(paste0("Skipping Fisher's test for Kingdom '", phy, "' (A_Unique vs Both) due to no data.\n"))
  }

# --- Comparison: SequencerB_Unique vs. Species_Both ---
  count_phy_B_unique <- sum(AVITI_uniq$Kingdom == phy)
  count_not_phy_B_unique <- sum(AVITI_uniq$Kingdom != phy)
  
  contingency_matrix_B_vs_Both <- matrix(c(count_phy_B_unique, count_not_phy_B_unique,
                                           count_phy_both, count_not_phy_both),
                                         nrow = 2, byrow = FALSE,
                                         dimnames = list(c(phy, paste0("Not_", phy)),
                                                         c("AVITI_Unique", "Taxa_Both")))
  
  if (sum(contingency_matrix_B_vs_Both) > 0) {
    fisher_test_B_vs_Both <- fisher.test(contingency_matrix_B_vs_Both, alternative = "two.sided")
    odds_ratio_B_vs_Both <- fisher_test_B_vs_Both$estimate
    p_value_B_vs_Both <- fisher_test_B_vs_Both$p.value
    
    enrichment_results_unique_vs_common_uniq <- bind_rows(enrichment_results_unique_vs_common_uniq, data.frame(
      Kingdom = phy,
      Comparison = "AVITI_Unique_vs_Both",
      OddsRatio = odds_ratio_B_vs_Both,
      P_Value = p_value_B_vs_Both
    ))
  } else {
    cat(paste0("Skipping Fisher's test for Kingdom '", phy, "' (B_Unique vs Both) due to no data.\n"))
  }
}

# Adjust p-values for multiple comparisons across all unique vs. common tests
if (nrow(enrichment_results_unique_vs_common_uniq) > 0) {
  enrichment_results_unique_vs_common_uniq$Adjusted_P_Value <- p.adjust(enrichment_results_unique_vs_common_uniq$P_Value, method = "BH")
  
  cat("\nIndividual Phylum Enrichment Analysis Results (Unique vs. Common - Fisher's Exact Test):\n")
  print(enrichment_results_unique_vs_common_uniq %>% arrange(Adjusted_P_Value))
  
  cat("\nInterpretation of Individual Kingdom Enrichment Analysis (Unique vs. Common):\n")
  cat("- 'OddsRatio' > 1 for 'SequencerA_Unique_vs_Both' suggests enrichment in SequencerA_Unique relative to Taxa detected by both.\n")
  cat("- 'OddsRatio' < 1 for 'SequencerA_Unique_vs_Both' suggests depletion in SequencerA_Unique relative to Taxa detected by both.\n")
  cat("- Similar interpretation applies for 'SequencerB_Unique_vs_Both'.\n")
  cat("- 'P_Value' indicates significance for that specific phylum and comparison.\n")
  cat("- 'Adjusted_P_Value' accounts for multiple comparisons.\n")
} else {
  cat("\nNo Kingdom found for individual enrichment analysis (Unique vs. Common).\n")
}
```
```{r}
plot_data_categorized_uniq <- bind_rows(
  MiSeq_uniq %>% mutate(DetectionGroup = "MiSeq_Unique"),
  AVITI_uniq %>% mutate(DetectionGroup = "AVITI_Unique"),
  shared_uniq %>% mutate(DetectionGroup = "Detected_by_Both")
)

# Calculate percentages for plotting across all three groups
king_percentages_overall_uniq <- plot_data_categorized_uniq %>%
  group_by(DetectionGroup, Kingdom) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(DetectionGroup) %>%
  mutate(Percentage = Count / sum(Count))

# Determine which phyla are significant based on enrichment_results_unique_vs_common
# A phylum is marked with '*' if either SequencerA_Unique_vs_Both or SequencerB_Unique_vs_Both is significant
significant_king_for_plot <- enrichment_results_unique_vs_common_uniq %>%
  filter(Adjusted_P_Value < 0.05) %>%
  distinct(Kingdom) %>%
  mutate(Label = "*")

# Merge significance labels into the plot data for correct positioning
plot_data_with_significance <- king_percentages_overall_uniq %>%
  left_join(significant_king_for_plot, by = "Kingdom")

# Find the maximum percentage for each phylum to position the asterisk
max_percentage_per_king <- plot_data_with_significance %>%
  group_by(Kingdom) %>%
  summarise(Max_Percentage = max(Percentage), .groups = 'drop')

# Join max percentage back to the significance data for plotting
significant_king_labels <- significant_king_for_plot %>%
  left_join(max_percentage_per_king, by = "Kingdom") %>%
  mutate(Y_Position = Max_Percentage + 0.02) # Add a small offset above the bar


# This bar plot helps visualize the phylum distribution for species
# unique to each sequencer and those detected by both, as percentages.
ggplot(plot_data_with_significance, aes(x = Kingdom, y = Percentage, fill = DetectionGroup)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Kingdom Distribution by Detection Group (Percentages)",
       x = "Kingdom",
       y = "Percentage of Taxa",
       fill = "Detection Group") +
  scale_y_continuous(labels = scales::percent, limits = c(0, max(plot_data_with_significance$Percentage) * 1.1)) + # Adjust y-axis limit
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  # Add asterisks for significant phyla
  geom_text(data = significant_king_labels, aes(x = Kingdom, y = Y_Position, label = Label),
            inherit.aes = FALSE, size = 6, color = "black") +
  # Add a small legend for the asterisk
  annotate("text", x = Inf, y = Inf, label = "* Adjusted P-value < 0.05 (Unique vs. Common)",
           hjust = 1.1, vjust = 1.1, size = 3, color = "black", fontface = "italic")
```


# Rarefied

```{r}
#ASV table
#ASV table
otu.c.rarefied <- read_csv(paste0(rarefaction_data, marker, "_Rarefied_data_OTU_table.csv"))
head(otu.c.rarefied)


#taxa table
#taxa table
tax.c.rarefied <- read_csv(paste0(rarefaction_data, marker, "_Rarefied_data_taxa_table.csv")) 
head(tax.c.rarefied)


#metadata table
samp.c.rarefied <- read_csv(paste0(rarefaction_data, marker, "_Rarefied_data_metadata_table.csv"))  %>%
  mutate(sample_source = fct_relevel(sample_source,  "CN23 MARS" , "MBTS 17024", "MBTS 20524", "MBTS 22624", "MBTS 25424", "CN24F 150m exp")) %>%
  mutate(short_name = "") %>%
  mutate(short_name = case_when(grepl("eDNA", sample_name) ~ substr(sample_name,1,nchar(sample_name)-9), T ~ short_name)) %>%
  mutate(short_name = case_when(grepl("CN24F150", sample_name) ~ substr(sample_name,14,nchar(sample_name)-3), T ~ short_name))
samp.c.rarefied
  
head(samp.c.rarefied)

#OTU table long format with percent total reads
potu.c.rarefied <- otu.c.rarefied %>%
  tidyr::pivot_longer( -ASV, names_to ='sample_name',values_to = 'reads' ) %>%
  ungroup() %>%
  group_by(sample_name) %>%
  mutate(per_tot = reads / sum(reads) *100) %>% #proportions(reads)*100
  arrange(-reads)
head(potu.c.rarefied)
```


## set lowest taxonomic annotation level for ASVs
```{r}

species_label <- tax.c.rarefied  %>%
  # assign  algae
  mutate(Phylum = case_when(Class == "Raphidophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Phaeophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Pelagophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Dictyochophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Chrysophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Synurophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Bolidophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Eustigmatophyceae" ~ "Ochrophyta", T ~ Phylum)) %>%
  # assign oomycotes
  mutate(Phylum = case_when(Class == "Oomycota" ~ "Oomycota", T ~ Phylum))  %>%
  mutate(Phylum = case_when(Class == "Oomycetes" ~ "Oomycota", T ~ Phylum))%>%
  # assign microeuk
  mutate(Phylum = case_when(Class == "Bigyra" ~ "Bigyra", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Choanoflagellata" ~ "Choanoflagellata", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Cryptophyceae" ~ "Cryptista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Order == "Telonemida" ~ "Cryptista", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Haptophyta" ~ "Haptista", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Centroplasthelida" ~ "Haptista", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Ichthyosporea" ~ "Choanozoa", T ~ Phylum))  %>% 
  
  mutate(Class = case_when(Phylum == "Imbricatea" ~ "Imbricatea", T ~ Class)) %>%
  mutate(Phylum = case_when(Class == "Imbricatea" ~ "Cercozoa", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Polycystinea" ~ "Retaria", T ~ Phylum)) %>%
  mutate(Phylum = case_when(Class == "Acantharea" ~ "Retaria", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Breviatea" ~ "Sulcozoa", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Aphelidea" ~ "Aphelida", T ~ Phylum)) %>%
  
  mutate(Phylum = case_when(Class == "Filasterea" ~ "Choanozoa", T ~ Phylum)) %>%
  
  mutate(Kingdom = case_when(Kingdom %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Kingdom)) %>%
  mutate(Phylum = case_when(Phylum %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Phylum)) %>% 
  mutate(Class = case_when(Class %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Class)) %>% 
  mutate(Order = case_when(Order %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Order)) %>% 
  mutate(Family = case_when(Family %in% c("unknown", "unassigned", "no_hit") ~ "unknown", T ~ Family)) %>% 
  mutate(Genus = case_when(Genus %in% c("unknown", "unassigned", "no_hit", "g_") ~ "unknown", T ~ Genus)) %>% 
  mutate(Species = case_when(Species %in% c("unknown", "unassigned", "no_hit", "s_") ~ "unknown", T ~ Species))

```


```{r}
# Compile data - 
# environmental data
env_data_rarefied <- inner_join(samp.c.rarefied, potu.c.rarefied, by = c("sample_name")) %>% 
  left_join(species_label)
head(env_data_rarefied)


```


## SHARED/UNIQUE Taxa

```{r}
length(unique(env_data_rarefied$Species))
dim(env_data_rarefied %>% select(sample_name, ASV, reads, sequencer, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  # filter(grepl(paste(grep_search), sample_name)) %>%
  filter(reads > 0))
```

## Unique taxa strings in Phyla
```{r}
taxa_data_unique <- env_data_rarefied %>% select(sample_name, ASV, reads, sequencer, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  filter(reads > 0) %>%
  group_by(sequencer, Phylum, Class, Order, Family, Genus,Species) %>%
  mutate(sum_reads_unique= sum(reads)) %>%
  ungroup() %>%
  distinct(sequencer, Phylum, Class, Order, Family, Genus, Species, .keep_all = T) %>%
  mutate(present = case_when(sum_reads_unique> 0 ~ 1, T ~ 0)) %>% 
  select(-ASV,-sample_name, -reads, -sum_reads_unique) %>%
  pivot_wider(names_from = "sequencer", values_from = "present", values_fill = 0)

shared_uniq_rare <- taxa_data_unique %>%
  filter(MiSeq == 1 & AVITI == 1)


MiSeq_uniq_rare <- taxa_data_unique %>%
  filter(MiSeq == 1 & AVITI == 0)


AVITI_uniq_rare <- taxa_data_unique %>%
  filter(MiSeq == 0 & AVITI == 1)

  
  
#comb_data <- make_comb_mat(data)

# Display the number of unique species found by each sequencer
cat("Number of unique taxa detected by MiSeq:", nrow((MiSeq_uniq_rare)), "\n")
cat("Number of unique taxa detected by AVITI:", nrow(AVITI_uniq_rare), "\n")
cat("Number of shared taxa detected:", nrow(shared_uniq_rare), "\n")

```

```{r}
# Create a combined dataset of taxas unique to each sequencer
# Add a column 'DetectionSource' to indicate which sequencer detected it uniquely
all_unique_detections_uniq <- bind_rows(
  MiSeq_uniq_rare %>% mutate(DetectionSource = "MiSeq_Unique"),
  AVITI_uniq_rare%>% mutate(DetectionSource = "AVITI_Unique")
)

# only keep phyla with at least 5 ASVs in at least 1 sequencer
phylum_contingency_table_uniq <- all_unique_detections_uniq %>%
  count(Phylum, DetectionSource) %>%
  pivot_wider(names_from = DetectionSource, values_from = n, values_fill = 0) 

cat("\nContingency Table (Phylum vs. Unique Detection Source):\n")
print(phylum_contingency_table_uniq)

```

### Phylum-specific Enrichment Analysis (Unique vs. Common Taxa), Rarefied 
This section compares the phylum distribution of taxas unique to each sequencer against the phylum distribution of taxas detected by *both* sequencers.

```{r}
phyla_combined_uniq <- unique(c(all_unique_detections_uniq$Phylum, shared_uniq_rare$Phylum))
enrichment_results_unique_vs_common_uniq <- data.frame(
  Phylum = character(),
  Comparison = character(), # e.g., "SequencerA_Unique_vs_Both", "SequencerB_Unique_vs_Both"
  OddsRatio = numeric(),
  P_Value = numeric(),
  Adjusted_P_Value = numeric(),
  stringsAsFactors = FALSE
)

```

```{r}
for (phy in phyla_combined_uniq) {
  # --- Comparison: SequencerA_Unique vs. ASV_Both ---
  count_phy_A_unique <- sum(MiSeq_uniq_rare$Phylum == phy)
  count_not_phy_A_unique <- sum(MiSeq_uniq_rare$Phylum != phy)
  
  count_phy_both <- sum(shared_uniq_rare$Phylum == phy)
  count_not_phy_both <- sum(shared_uniq_rare$Phylum != phy)
  
  contingency_matrix_A_vs_Both <- matrix(c(count_phy_A_unique, count_not_phy_A_unique,
                                           count_phy_both, count_not_phy_both),
                                         nrow = 2, byrow = FALSE,
                                         dimnames = list(c(phy, paste0("Not_", phy)),
                                                         c("MiSeq_Unique", "Taxa_Both")))
  
  if (sum(contingency_matrix_A_vs_Both) > 0) {
    fisher_test_A_vs_Both <- fisher.test(contingency_matrix_A_vs_Both, alternative = "two.sided")
    odds_ratio_A_vs_Both <- fisher_test_A_vs_Both$estimate
    p_value_A_vs_Both <- fisher_test_A_vs_Both$p.value
    
    enrichment_results_unique_vs_common_uniq <- bind_rows(enrichment_results_unique_vs_common_uniq, data.frame(
      Phylum = phy,
      Comparison = "MiSeq_Unique_vs_Both",
      OddsRatio = odds_ratio_A_vs_Both,
      P_Value = p_value_A_vs_Both
    ))
  } else {
    cat(paste0("Skipping Fisher's test for Phylum '", phy, "' (A_Unique vs Both) due to no data.\n"))
  }

# --- Comparison: SequencerB_Unique vs. Species_Both ---
  count_phy_B_unique <- sum(AVITI_uniq_rare$Phylum == phy)
  count_not_phy_B_unique <- sum(AVITI_uniq_rare$Phylum != phy)
  
  contingency_matrix_B_vs_Both <- matrix(c(count_phy_B_unique, count_not_phy_B_unique,
                                           count_phy_both, count_not_phy_both),
                                         nrow = 2, byrow = FALSE,
                                         dimnames = list(c(phy, paste0("Not_", phy)),
                                                         c("AVITI_Unique", "Taxa_Both")))
  
  if (sum(contingency_matrix_B_vs_Both) > 0) {
    fisher_test_B_vs_Both <- fisher.test(contingency_matrix_B_vs_Both, alternative = "two.sided")
    odds_ratio_B_vs_Both <- fisher_test_B_vs_Both$estimate
    p_value_B_vs_Both <- fisher_test_B_vs_Both$p.value
    
    enrichment_results_unique_vs_common_uniq <- bind_rows(enrichment_results_unique_vs_common_uniq, data.frame(
      Phylum = phy,
      Comparison = "AVITI_Unique_vs_Both",
      OddsRatio = odds_ratio_B_vs_Both,
      P_Value = p_value_B_vs_Both
    ))
  } else {
    cat(paste0("Skipping Fisher's test for Phylum '", phy, "' (B_Unique vs Both) due to no data.\n"))
  }
}

# Adjust p-values for multiple comparisons across all unique vs. common tests
if (nrow(enrichment_results_unique_vs_common_uniq) > 0) {
  enrichment_results_unique_vs_common_uniq$Adjusted_P_Value <- p.adjust(enrichment_results_unique_vs_common_uniq$P_Value, method = "BH")
  
  cat("\nIndividual Phylum Enrichment Analysis Results (Unique vs. Common - Fisher's Exact Test):\n")
  print(enrichment_results_unique_vs_common_uniq %>% arrange(Adjusted_P_Value))
  
  cat("\nInterpretation of Individual Phylum Enrichment Analysis (Unique vs. Common):\n")
  cat("- 'OddsRatio' > 1 for 'SequencerA_Unique_vs_Both' suggests enrichment in SequencerA_Unique relative to Taxa detected by both.\n")
  cat("- 'OddsRatio' < 1 for 'SequencerA_Unique_vs_Both' suggests depletion in SequencerA_Unique relative to Taxa detected by both.\n")
  cat("- Similar interpretation applies for 'SequencerB_Unique_vs_Both'.\n")
  cat("- 'P_Value' indicates significance for that specific phylum and comparison.\n")
  cat("- 'Adjusted_P_Value' accounts for multiple comparisons.\n")
} else {
  cat("\nNo phyla found for individual enrichment analysis (Unique vs. Common).\n")
}
```
```{r}
filename <- paste(enrichment_directory, proj_prefix, '_', marker, '_Rarefied_enrichment_results_uniqueTaxa_in_Phyla.csv', sep='')
write.csv(enrichment_results_unique_vs_common_uniq %>% arrange(Adjusted_P_Value),filename,row.names = FALSE,quote = FALSE )
```

# ## S.Fig. 5D
```{r}
plot_data_categorized_uniq <- bind_rows(
  MiSeq_uniq_rare %>% mutate(DetectionGroup = "MiSeq_Unique"),
  AVITI_uniq_rare %>% mutate(DetectionGroup = "AVITI_Unique"),
  shared_uniq_rare %>% mutate(DetectionGroup = "Detected_by_Both")
)

# Calculate percentages for plotting across all three groups
phylum_percentages_overall_uniq <- plot_data_categorized_uniq %>%
  group_by(DetectionGroup, Phylum) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(DetectionGroup) %>%
  mutate(Percentage = Count / sum(Count))

# Determine which phyla are significant based on enrichment_results_unique_vs_common
# A phylum is marked with '*' if either SequencerA_Unique_vs_Both or SequencerB_Unique_vs_Both is significant
significant_phyla_for_plot <- enrichment_results_unique_vs_common_uniq %>%
  filter(Adjusted_P_Value < 0.05) %>%
  distinct(Phylum) %>%
  mutate(Label = "*")

# Merge significance labels into the plot data for correct positioning
plot_data_with_significance <- phylum_percentages_overall_uniq %>%
  left_join(significant_phyla_for_plot, by = "Phylum") %>%
  filter(!is.na(Label))

# Find the maximum percentage for each phylum to position the asterisk
max_percentage_per_phylum <- plot_data_with_significance %>%
  group_by(Phylum) %>%
  summarise(Max_Percentage = max(Percentage), .groups = 'drop')

# Join max percentage back to the significance data for plotting
significant_phyla_labels <- significant_phyla_for_plot %>%
  left_join(max_percentage_per_phylum, by = "Phylum") %>%
  mutate(Y_Position = Max_Percentage + 0.02) # Add a small offset above the bar


# This bar plot helps visualize the phylum distribution for species
# unique to each sequencer and those detected by both, as percentages.
ggplot(plot_data_with_significance, aes(x = Phylum, y = Percentage, fill = DetectionGroup)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Distribution of Uniques/Shared Taxa by Phyla and Sequencer, Rarefied",
       x = "Phylum",
       y = "Percentage of Total Taxa in Group",
       fill = "Group") +
  scale_y_continuous(labels = scales::percent, limits = c(0, max(plot_data_with_significance$Percentage) * 1.1)) + # Adjust y-axis limit
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=15),
        axis.text.y = element_text(size=15),
        axis.title = element_text(size = 18),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 15),
        plot.title = element_text(size = 18)) +
  # Add asterisks for significant phyla
  geom_text(data = significant_phyla_labels, aes(x = Phylum, y = Y_Position, label = Label),
            inherit.aes = FALSE, size = 6, color = "black") +
  # Add a small legend for the asterisk
  annotate("text", x = Inf, y = Inf, label = "* Adjusted P-value < 0.05 (Unique vs. Common)",
           hjust = 1.1, vjust = 1.1, size = 5, color = "black", fontface = "italic")



filename = paste(sfig5_directory, 'SFig5d_enrichment_analysis_taxa_rarefied_',proj_prefix, '_', marker, '.png', sep='')
filename
ggsave(filename,height = 7, width =10, units = 'in')

filename = paste(sfig5_directory, 'SFig5d_enrichment_analysis_taxa_rarefied_',proj_prefix, '_', marker, '.svg', sep='')
filename
ggsave(filename,height = 7, width =10, units = 'in')



```
```{r}
phylum_percentages_overall_uniq %>%
  filter(Phylum %in% significant_phyla_for_plot$Phylum) %>%
  select(-Percentage) %>%
  pivot_wider(names_from = DetectionGroup, values_from = Count)

phylum_percentages_overall_uniq %>%
  filter(Phylum %in% significant_phyla_for_plot$Phylum) %>%
  select(-Count) %>%
  pivot_wider(names_from = DetectionGroup, values_from = Percentage)

```

## Species in Phyla, Rarefied. 
```{r}
taxa_data_spp <- env_data_rarefied %>% select(sample_name, ASV, reads, sequencer, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  # filter(grepl(paste(grep_search), sample_name)) %>%
  filter(reads > 0) %>%
  group_by(sequencer, Species) %>%
  mutate(sum_reads_spp = sum(reads)) %>%
  ungroup() %>%
  distinct(sequencer, Species, .keep_all = T) %>%
  mutate(present = case_when(sum_reads_spp > 0 ~ 1, T ~ 0)) %>% 
  select(-ASV,-sample_name, -reads, -sum_reads_spp) %>%
  pivot_wider(names_from = "sequencer", values_from = "present", values_fill = 0) %>% arrange(Species)

shared_spp_rare <- taxa_data_spp %>%
  filter(MiSeq == 1 & AVITI == 1)


MiSeq_spp_rare <- taxa_data_spp %>%
  filter(MiSeq == 1 & AVITI == 0)


AVITI_spp_rare <- taxa_data_spp %>%
  filter(MiSeq == 0 & AVITI == 1)

  
  
#comb_data <- make_comb_mat(data)

# Display the number of unique species found by each sequencer
cat("Number of unique species detected by MiSeq, Rarefied:", nrow((MiSeq_spp_rare)), "\n")
cat("Number of unique species detected by AVITI, Rarefied:", nrow(AVITI_spp_rare), "\n")
cat("Number of shared species detected, Rarefied:", nrow(shared_spp_rare), "\n")

```

```{r}
# Create a combined dataset of species unique to each sequencer
# Add a column 'DetectionSource' to indicate which sequencer detected it uniquely
all_unique_detections <- bind_rows(
  MiSeq_spp_rare %>% mutate(DetectionSource = "MiSeq_Unique"),
  AVITI_spp_rare%>% mutate(DetectionSource = "AVITI_Unique")
)

phylum_contingency_table <- all_unique_detections %>%
  count(Phylum, DetectionSource) %>%
  pivot_wider(names_from = DetectionSource, values_from = n, values_fill = 0)

cat("\nContingency Table (Phylum vs. Unique Detection Source):\n")
print(phylum_contingency_table)

```

### Phylum-specific Enrichment Analysis (Unique vs. Common Species), Rarefied 
This section compares the phylum distribution of species unique to each sequencer against the phylum distribution of species detected by *both* sequencers.

```{r}
all_phyla_combined <- unique(c(all_unique_detections$Phylum, shared_spp_rare$Phylum))
enrichment_results_unique_vs_common <- data.frame(
  Phylum = character(),
  Comparison = character(), # e.g., "SequencerA_Unique_vs_Both", "SequencerB_Unique_vs_Both"
  OddsRatio = numeric(),
  P_Value = numeric(),
  Adjusted_P_Value = numeric(),
  stringsAsFactors = FALSE
)

```

```{r}
for (phy in all_phyla_combined) {
  # --- Comparison: SequencerA_Unique vs. Species_Both ---
  count_phy_A_unique <- sum(MiSeq_spp_rare$Phylum == phy)
  count_not_phy_A_unique <- sum(MiSeq_spp_rare$Phylum != phy)
  
  count_phy_both <- sum(shared_spp_rare$Phylum == phy)
  count_not_phy_both <- sum(shared_spp_rare$Phylum != phy)
  
  contingency_matrix_A_vs_Both <- matrix(c(count_phy_A_unique, count_not_phy_A_unique,
                                           count_phy_both, count_not_phy_both),
                                         nrow = 2, byrow = FALSE,
                                         dimnames = list(c(phy, paste0("Not_", phy)),
                                                         c("MiSeq_Unique", "Species_Both")))
  
  if (sum(contingency_matrix_A_vs_Both) > 0) {
    fisher_test_A_vs_Both <- fisher.test(contingency_matrix_A_vs_Both, alternative = "two.sided")
    odds_ratio_A_vs_Both <- fisher_test_A_vs_Both$estimate
    p_value_A_vs_Both <- fisher_test_A_vs_Both$p.value
    
    enrichment_results_unique_vs_common <- bind_rows(enrichment_results_unique_vs_common, data.frame(
      Phylum = phy,
      Comparison = "MiSeq_Unique_vs_Both",
      OddsRatio = odds_ratio_A_vs_Both,
      P_Value = p_value_A_vs_Both
    ))
  } else {
    cat(paste0("Skipping Fisher's test for Phylum '", phy, "' (A_Unique vs Both) due to no data.\n"))
  }

# --- Comparison: SequencerB_Unique vs. Species_Both ---
  count_phy_B_unique <- sum(AVITI_spp_rare$Phylum == phy)
  count_not_phy_B_unique <- sum(AVITI_spp_rare$Phylum != phy)
  
  contingency_matrix_B_vs_Both <- matrix(c(count_phy_B_unique, count_not_phy_B_unique,
                                           count_phy_both, count_not_phy_both),
                                         nrow = 2, byrow = FALSE,
                                         dimnames = list(c(phy, paste0("Not_", phy)),
                                                         c("AVITI_Unique", "Species_Both")))
  
  if (sum(contingency_matrix_B_vs_Both) > 0) {
    fisher_test_B_vs_Both <- fisher.test(contingency_matrix_B_vs_Both, alternative = "two.sided")
    odds_ratio_B_vs_Both <- fisher_test_B_vs_Both$estimate
    p_value_B_vs_Both <- fisher_test_B_vs_Both$p.value
    
    enrichment_results_unique_vs_common <- bind_rows(enrichment_results_unique_vs_common, data.frame(
      Phylum = phy,
      Comparison = "AVITI_Unique_vs_Both",
      OddsRatio = odds_ratio_B_vs_Both,
      P_Value = p_value_B_vs_Both
    ))
  } else {
    cat(paste0("Skipping Fisher's test for Phylum '", phy, "' (B_Unique vs Both) due to no data.\n"))
  }
}

# Adjust p-values for multiple comparisons across all unique vs. common tests
if (nrow(enrichment_results_unique_vs_common) > 0) {
  enrichment_results_unique_vs_common$Adjusted_P_Value <- p.adjust(enrichment_results_unique_vs_common$P_Value, method = "BH")
  
  cat("\nIndividual Phylum Enrichment Analysis Results (Unique vs. Common - Fisher's Exact Test):\n")
  print(enrichment_results_unique_vs_common %>% arrange(Adjusted_P_Value))
  
  cat("\nInterpretation of Individual Phylum Enrichment Analysis (Unique vs. Common):\n")
  cat("- 'OddsRatio' > 1 for 'SequencerA_Unique_vs_Both' suggests enrichment in SequencerA_Unique relative to species detected by both.\n")
  cat("- 'OddsRatio' < 1 for 'SequencerA_Unique_vs_Both' suggests depletion in SequencerA_Unique relative to species detected by both.\n")
  cat("- Similar interpretation applies for 'SequencerB_Unique_vs_Both'.\n")
  cat("- 'P_Value' indicates significance for that specific phylum and comparison.\n")
  cat("- 'Adjusted_P_Value' accounts for multiple comparisons.\n")
} else {
  cat("\nNo phyla found for individual enrichment analysis (Unique vs. Common).\n")
}
```
```{r}
filename <- paste(enrichment_directory, proj_prefix, '_', marker, '_Rarefied_enrichment_results_uniqueSpp_in_Phyla.csv', sep='')
write.csv(enrichment_results_unique_vs_common %>% arrange(Adjusted_P_Value),filename,row.names = FALSE,quote = FALSE )
```

# ## S.FIG 5E
```{r}
plot_data_categorized <- bind_rows(
  MiSeq_spp_rare %>% mutate(DetectionGroup = "MiSeq_Unique"),
  AVITI_spp_rare %>% mutate(DetectionGroup = "AVITI_Unique"),
  shared_spp_rare %>% mutate(DetectionGroup = "Detected_by_Both")
)

# Calculate percentages for plotting across all three groups
phylum_percentages_overall <- plot_data_categorized %>%
  group_by(DetectionGroup, Phylum) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(DetectionGroup) %>%
  mutate(Percentage = Count / sum(Count))

# Determine which phyla are significant based on enrichment_results_unique_vs_common
# A phylum is marked with '*' if either SequencerA_Unique_vs_Both or SequencerB_Unique_vs_Both is significant
significant_phyla_for_plot <- enrichment_results_unique_vs_common %>%
  filter(Adjusted_P_Value < 0.05) %>%
  distinct(Phylum) %>%
  mutate(Label = "*")

# Merge significance labels into the plot data for correct positioning
plot_data_with_significance <- phylum_percentages_overall %>%
  left_join(significant_phyla_for_plot, by = "Phylum") %>%
  filter(!is.na(Label))

# Find the maximum percentage for each phylum to position the asterisk
max_percentage_per_phylum <- plot_data_with_significance %>%
  group_by(Phylum) %>%
  summarise(Max_Percentage = max(Percentage), .groups = 'drop')

# Join max percentage back to the significance data for plotting
significant_phyla_labels <- significant_phyla_for_plot %>%
  left_join(max_percentage_per_phylum, by = "Phylum") %>%
  mutate(Y_Position = Max_Percentage + 0.02) # Add a small offset above the bar


# This bar plot helps visualize the phylum distribution for species
# unique to each sequencer and those detected by both, as percentages.
ggplot(plot_data_with_significance, aes(x = Phylum, y = Percentage, fill = DetectionGroup)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Distribution of unique/shared species by Phyla and Sequencer, Rarefied",
       x = "Phylum",
       y = "Percentage of Total Species in Group",
       fill = "Group") +
  scale_y_continuous(labels = scales::percent, limits = c(0, max(plot_data_with_significance$Percentage) * 1.15)) + # Adjust y-axis limit
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=15),
        axis.text.y = element_text(size=15),
        axis.title = element_text(size = 18),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 15),
        plot.title = element_text(size = 18)) +
  # Add asterisks for significant phyla
  geom_text(data = significant_phyla_labels, aes(x = Phylum, y = Y_Position, label = Label),
            inherit.aes = FALSE, size = 6, color = "black") +
  # Add a small legend for the asterisk
  annotate("text", x = Inf, y = Inf, label = "* Adjusted P-value < 0.05 (Unique vs. Common)",
           hjust = 1.1, vjust = 1.1, size = 5, color = "black", fontface = "italic")



filename = paste(sfig5_directory, 'SFig5e_enrichment_analysis_species_rarefied_',proj_prefix, '_', marker, '.png', sep='')
filename
ggsave(filename,height = 7, width =10, units = 'in')

filename = paste(sfig5_directory, 'SFig5e_enrichment_analysis_species_rarefied_',proj_prefix, '_', marker, '.svg', sep='')
filename
ggsave(filename,height = 7, width =10, units = 'in')


```
```{r}
phylum_percentages_overall %>%
  filter(Phylum == "Foraminifera")
```

Plot raw spp numbers (not done)
```{r}
plot_data_categorized <- bind_rows(
  MiSeq_spp %>% mutate(DetectionGroup = "MiSeq_Unique"),
  AVITI_spp %>% mutate(DetectionGroup = "AVITI_Unique"),
  shared_spp %>% mutate(DetectionGroup = "Detected_by_Both")
)

# Calculate percentages for plotting across all three groups
phylum_percentages_overall <- plot_data_categorized %>%
  group_by(DetectionGroup, Phylum) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(DetectionGroup) %>%
  mutate(Percentage = Count / sum(Count))

# Determine which phyla are significant based on enrichment_results_unique_vs_common
# A phylum is marked with '*' if either SequencerA_Unique_vs_Both or SequencerB_Unique_vs_Both is significant
significant_phyla_for_plot <- enrichment_results_unique_vs_common %>%
  filter(Adjusted_P_Value < 0.05) %>%
  distinct(Phylum) %>%
  mutate(Label = "*")

# Merge significance labels into the plot data for correct positioning
plot_data_with_significance <- phylum_percentages_overall %>%
  left_join(significant_phyla_for_plot, by = "Phylum")

# Find the maximum percentage for each phylum to position the asterisk
max_percentage_per_phylum <- plot_data_with_significance %>%
  group_by(Phylum) %>%
  summarise(Max_Percentage = max(Percentage), .groups = 'drop')

# Join max percentage back to the significance data for plotting
significant_phyla_labels <- significant_phyla_for_plot %>%
  left_join(max_percentage_per_phylum, by = "Phylum") %>%
  mutate(Y_Position = Max_Percentage + 0.02) # Add a small offset above the bar


# This bar plot helps visualize the phylum distribution for species
# unique to each sequencer and those detected by both, as percentages.
ggplot(plot_data_with_significance, aes(x = Phylum, y = Percentage, fill = DetectionGroup)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Phylum Distribution by Detection Group (Percentages)",
       x = "Phylum",
       y = "Percentage of Species",
       fill = "Detection Group") +
  scale_y_continuous(labels = scales::percent, limits = c(0, max(plot_data_with_significance$Percentage) * 1.15)) + # Adjust y-axis limit
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  # Add asterisks for significant phyla
  geom_text(data = significant_phyla_labels, aes(x = Phylum, y = Y_Position, label = Label),
            inherit.aes = FALSE, size = 6, color = "black") +
  # Add a small legend for the asterisk
  annotate("text", x = Inf, y = Inf, label = "* Adjusted P-value < 0.05 (Unique vs. Common)",
           hjust = 1.1, vjust = 1.1, size = 3, color = "black", fontface = "italic")
```

## ASVs in Phyla
```{r}
taxa_data_asv <- env_data_rarefied %>% select(sample_name, ASV, reads, sequencer, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  # filter(grepl(paste(grep_search), sample_name)) %>%
  filter(reads > 0) %>%
  group_by(sequencer, ASV) %>%
  mutate(sum_reads_asv = sum(reads)) %>%
  ungroup() %>%
  distinct(sequencer, ASV, .keep_all = T) %>%
  mutate(present = case_when(sum_reads_asv > 0 ~ 1, T ~ 0)) %>% 
  select(-sample_name, -reads, -sum_reads_asv) %>%
  pivot_wider(names_from = "sequencer", values_from = "present", values_fill = 0) %>% arrange(ASV)

shared_asv_rare <- taxa_data_asv %>%
  filter(MiSeq == 1 & AVITI == 1)


MiSeq_asv_rare <- taxa_data_asv %>%
  filter(MiSeq == 1 & AVITI == 0)


AVITI_asv_rare <- taxa_data_asv %>%
  filter(MiSeq == 0 & AVITI == 1)

  
  
#comb_data <- make_comb_mat(data)

# Display the number of unique species found by each sequencer
cat("Number of unique ASVs detected by MiSeq, Rarefied:", nrow((MiSeq_asv_rare)), "\n")
cat("Number of unique ASVs detected by AVITI, Rarefied:", nrow(AVITI_asv_rare), "\n")
cat("Number of shared ASVs detected, Rarefied:", nrow(shared_asv_rare), "\n")

```

```{r}
# Create a combined dataset of ASVs unique to each sequencer
# Add a column 'DetectionSource' to indicate which sequencer detected it uniquely
all_unique_detections_asv <- bind_rows(
  MiSeq_asv_rare %>% mutate(DetectionSource = "MiSeq_Unique"),
  AVITI_asv_rare%>% mutate(DetectionSource = "AVITI_Unique")
)

# only keep phyla with at least 5 ASVs in at least 1 sequencer
phylum_contingency_table_asv <- all_unique_detections_asv %>%
  count(Phylum, DetectionSource) %>%
  pivot_wider(names_from = DetectionSource, values_from = n, values_fill = 0) %>%
  filter(if_any(c(MiSeq_Unique, AVITI_Unique), ~ . > 4))

cat("\nContingency Table (Phylum vs. Unique Detection Source):\n")
print(phylum_contingency_table_asv)

```

### Phylum-specific Enrichment Analysis (Unique vs. Common ASVs) 
This section compares the phylum distribution of ASVs unique to each sequencer against the phylum distribution of ASVs detected by *both* sequencers.

```{r}
phyla_combined_asv <- unique(c(all_unique_detections_asv$Phylum, shared_asv_rare$Phylum))
filt_phyla_combined_asv <- unique(phylum_contingency_table_asv$Phylum)
enrichment_results_unique_vs_common_asv <- data.frame(
  Phylum = character(),
  Comparison = character(), # e.g., "SequencerA_Unique_vs_Both", "SequencerB_Unique_vs_Both"
  OddsRatio = numeric(),
  P_Value = numeric(),
  Adjusted_P_Value = numeric(),
  stringsAsFactors = FALSE
)

```

```{r}
for (phy in filt_phyla_combined_asv) {
  # --- Comparison: SequencerA_Unique vs. ASV_Both ---
  count_phy_A_unique <- sum(MiSeq_asv_rare$Phylum == phy)
  count_not_phy_A_unique <- sum(MiSeq_asv_rare$Phylum != phy)
  
  count_phy_both <- sum(shared_asv_rare$Phylum == phy)
  count_not_phy_both <- sum(shared_asv_rare$Phylum != phy)
  
  contingency_matrix_A_vs_Both <- matrix(c(count_phy_A_unique, count_not_phy_A_unique,
                                           count_phy_both, count_not_phy_both),
                                         nrow = 2, byrow = FALSE,
                                         dimnames = list(c(phy, paste0("Not_", phy)),
                                                         c("MiSeq_Unique", "ASV_Both")))
  
  if (sum(contingency_matrix_A_vs_Both) > 0) {
    fisher_test_A_vs_Both <- fisher.test(contingency_matrix_A_vs_Both, alternative = "two.sided")
    odds_ratio_A_vs_Both <- fisher_test_A_vs_Both$estimate
    p_value_A_vs_Both <- fisher_test_A_vs_Both$p.value
    
    enrichment_results_unique_vs_common_asv <- bind_rows(enrichment_results_unique_vs_common_asv, data.frame(
      Phylum = phy,
      Comparison = "MiSeq_Unique_vs_Both",
      OddsRatio = odds_ratio_A_vs_Both,
      P_Value = p_value_A_vs_Both
    ))
  } else {
    cat(paste0("Skipping Fisher's test for Phylum '", phy, "' (A_Unique vs Both) due to no data.\n"))
  }

# --- Comparison: SequencerB_Unique vs. Species_Both ---
  count_phy_B_unique <- sum(AVITI_asv_rare$Phylum == phy)
  count_not_phy_B_unique <- sum(AVITI_asv_rare$Phylum != phy)
  
  contingency_matrix_B_vs_Both <- matrix(c(count_phy_B_unique, count_not_phy_B_unique,
                                           count_phy_both, count_not_phy_both),
                                         nrow = 2, byrow = FALSE,
                                         dimnames = list(c(phy, paste0("Not_", phy)),
                                                         c("AVITI_Unique", "ASV_Both")))
  
  if (sum(contingency_matrix_B_vs_Both) > 0) {
    fisher_test_B_vs_Both <- fisher.test(contingency_matrix_B_vs_Both, alternative = "two.sided")
    odds_ratio_B_vs_Both <- fisher_test_B_vs_Both$estimate
    p_value_B_vs_Both <- fisher_test_B_vs_Both$p.value
    
    enrichment_results_unique_vs_common_asv <- bind_rows(enrichment_results_unique_vs_common_asv, data.frame(
      Phylum = phy,
      Comparison = "AVITI_Unique_vs_Both",
      OddsRatio = odds_ratio_B_vs_Both,
      P_Value = p_value_B_vs_Both
    ))
  } else {
    cat(paste0("Skipping Fisher's test for Phylum '", phy, "' (B_Unique vs Both) due to no data.\n"))
  }
}

# Adjust p-values for multiple comparisons across all unique vs. common tests
if (nrow(enrichment_results_unique_vs_common_asv) > 0) {
  enrichment_results_unique_vs_common_asv$Adjusted_P_Value <- p.adjust(enrichment_results_unique_vs_common_asv$P_Value, method = "BH")
  
  cat("\nIndividual Phylum Enrichment Analysis Results (Unique vs. Common - Fisher's Exact Test):\n")
  print(enrichment_results_unique_vs_common_asv %>% arrange(Adjusted_P_Value))
  
  cat("\nInterpretation of Individual Phylum Enrichment Analysis (Unique vs. Common):\n")
  cat("- 'OddsRatio' > 1 for 'SequencerA_Unique_vs_Both' suggests enrichment in SequencerA_Unique relative to ASVs detected by both.\n")
  cat("- 'OddsRatio' < 1 for 'SequencerA_Unique_vs_Both' suggests depletion in SequencerA_Unique relative to ASVs detected by both.\n")
  cat("- Similar interpretation applies for 'SequencerB_Unique_vs_Both'.\n")
  cat("- 'P_Value' indicates significance for that specific phylum and comparison.\n")
  cat("- 'Adjusted_P_Value' accounts for multiple comparisons.\n")
} else {
  cat("\nNo phyla found for individual enrichment analysis (Unique vs. Common).\n")
}
```
```{r}
filename <- paste(enrichment_directory,  proj_prefix, '_', marker, '_Rarefied_enrichment_results_uniqueASVs_in_Phyla.csv', sep='')
write.csv(enrichment_results_unique_vs_common_asv %>% arrange(Adjusted_P_Value),filename,row.names = FALSE,quote = FALSE )
```

# ## S.FIG 5F
```{r}
plot_data_categorized_asv <- bind_rows(
  MiSeq_asv_rare %>% mutate(DetectionGroup = "MiSeq_Unique"),
  AVITI_asv_rare %>% mutate(DetectionGroup = "AVITI_Unique"),
  shared_asv_rare %>% mutate(DetectionGroup = "Detected_by_Both")
) %>% filter(Phylum %in% filt_phyla_combined_asv)

# Calculate percentages for plotting across all three groups
phylum_percentages_overall_asv <- plot_data_categorized_asv %>%
  group_by(DetectionGroup, Phylum) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(DetectionGroup) %>%
  mutate(Percentage = Count / sum(Count))

# Determine which phyla are significant based on enrichment_results_unique_vs_common
# A phylum is marked with '*' if either SequencerA_Unique_vs_Both or SequencerB_Unique_vs_Both is significant
significant_phyla_for_plot <- enrichment_results_unique_vs_common_asv %>%
  filter(Adjusted_P_Value < 0.05) %>%
  distinct(Phylum) %>%
  mutate(Label = "*")

# Merge significance labels into the plot data for correct positioning
plot_data_with_significance <- phylum_percentages_overall_asv %>%
  left_join(significant_phyla_for_plot, by = "Phylum") %>% 
  filter(!is.na(Label))

# Find the maximum percentage for each phylum to position the asterisk
max_percentage_per_phylum <- plot_data_with_significance %>%
  group_by(Phylum) %>%
  summarise(Max_Percentage = max(Percentage), .groups = 'drop')

# Join max percentage back to the significance data for plotting
significant_phyla_labels <- significant_phyla_for_plot %>%
  left_join(max_percentage_per_phylum, by = "Phylum") %>%
  mutate(Y_Position = Max_Percentage + 0.02) # Add a small offset above the bar


# This bar plot helps visualize the phylum distribution for species
# unique to each sequencer and those detected by both, as percentages.
ggplot(plot_data_with_significance, aes(x = Phylum, y = Percentage, fill = DetectionGroup)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Distribution of unique/shared ASVs by Sequencer and Phyla",
       x = "Phylum",
       y = "Percentage of Total ASVs in Group",
       fill = "Group") +
  scale_y_continuous(labels = scales::percent, limits = c(0, max(plot_data_with_significance$Percentage) * 1.1)) + # Adjust y-axis limit
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=15),
        axis.text.y = element_text(size=15),
        axis.title = element_text(size = 18),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 15),
        plot.title = element_text(size = 18)) +
  # Add asterisks for significant phyla
  geom_text(data = significant_phyla_labels, aes(x = Phylum, y = Y_Position, label = Label),
            inherit.aes = FALSE, size = 6, color = "black") +
  # Add a small legend for the asterisk
  annotate("text", x = Inf, y = Inf, label = "* Adjusted P-value < 0.05 (Unique vs. Common)",
           hjust = 1.1, vjust = 1.1, size = 5, color = "black", fontface = "italic")



filename = paste(sfig5_directory, 'SFig5f_enrichment_analysis_ASVs_rarefied_',proj_prefix, '_', marker, '.png', sep='')
filename
ggsave(filename,height = 7, width =10, units = 'in')

filename = paste(sfig5_directory, 'SFig5f_enrichment_analysis_ASVs_rarefied_',proj_prefix, '_', marker, '.svg', sep='')
filename
ggsave(filename,height = 7, width =10, units = 'in')



```

```{r}
phylum_percentages_overall_asv %>%
  filter(Phylum %in% significant_phyla_for_plot$Phylum) %>%
  select(-Percentage) %>%
  pivot_wider(names_from = DetectionGroup, values_from = Count)

phylum_percentages_overall_asv%>%
  filter(Phylum %in% significant_phyla_for_plot$Phylum) %>%
  select(-Count) %>%
  pivot_wider(names_from = DetectionGroup, values_from = Percentage)

```













